### 2011-06-14 Article.tex
Forming the base

\documentclass{article}

\usepackage {amsmath}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}
\usepackage{lineno}

% some formatting tags
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textwidth      6.5in

\title{Temperature Effects on Surface Sulfur Dioxide}
\author{Eric S. Shamay \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

%\maketitle

%\onehalfspacing
\linenumbers 
\doublespacing

%\begin{abstract}
	%...Abstract goes here...
%\end{abstract}

%\include{Introduction}
%\include{Methods}
%\include{Results}
%\include{Conclusions}

\bibliography{bib/so2temperature}

\end{document}

### 2011-06-14 clean.sh

@@ -1,0 +1,5 @@
+rm *.aux
+rm *.log
+rm *.pdf
+rm *.blg
+rm *.bbl

### 2011-06-14 ComputationalMethods.tex
New computational methods section.

\section{Computational Methods}

On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).~\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions. The electronic charge density was desccribed using an auxilliary basis set of plane waves. Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1.0 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the cold and hot simulations, respectively. It was verified that these computational parameters yield a reasonable description of bulk room temperature water when simulating a neat-water system.

Initially, 10 equilibrated boxes of side-lengths 10.0\angs~, with 36 randomly packed water molecules were used. 5 of the boxes were used for each of the cold and hot simulations. A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis. A copy of the initial system cubes were then expanded along one axis (z-axis) to 25\angs. The system energy was minimized. Subsequently, the system was equilibrated for 1 ns in canonical ensemble (NVT) conditions. Periodic boundaries were set on the two short axes to form an infinite slab. The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1.0 fs. The initial 1 ns equilibration trajectory was not included in the final analysis. This resulted in 20,000 time steps of system trajectory for analysis in each replica of the system.

### 2011-06-15 Introduction.tex
Working on coordination images and typed up some results.

\section {Introduction}

The chemistry of our atmosphere is vastly complex and poorly understood. Reactions taking place on particle surfaces and liquid boundaries dominate many atmospheric chemical pathways, and guide terrestrial and industrial byproducts to their ultimate forms and destinations. How do atmospheric conditions affect the chemistry of environmental constituents?

### 2011-06-15 coordination.tex
Working on coordination images and typed up some results.

\section {Sulfur Dioxide Bonding Coordinations}

To further assess the effect of temperature on \suldiox~behavior we turn to analysis of the bonds formed between \suldiox~and the water surface. The bonding coordinations were determined for the \suldiox~based on the bondlength criteria noted in the introduction. Bonds formed between \suldiox-sulfur and \wat-oxygen are labeled 'S' coordinated, and bonds formed between \suldiox-oxygen and \wat-hydrogen are labeled 'O' coordinated. For each bond an additional letter label is appended to the coordination. For example, an 'SOO' coordinated \suldiox~forms three bonds to waters: two bonds are formed through the \suldiox-oxygens, and one through the \suldiox~sulfur. This naming scheme is taken from the work of Baer et al. where the \suldiox~bonding distribution was determined for a surface molecule at a single temperature simulated under similar conditions.\cite{Baer2010}

Figure \ref{fig:so2-bonding-coordinations} shows the distribution of bonding coordinations throughout the set of simulated trajectories for both the cold and hot temperatures.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/coordination/so2-bonding-coordinations.png}
		\caption{}
		\label{fig:so2-bonding-coordinations
	\end{center}
\end{figure}

### 2011-06-30 images/scripts/PlotCyclicStructureDistributions.py
Added lots and LOTS of images for the cyclic structures, and starting a scripts directory for plotting things out.

@@ -1,0 +1,100 @@
+import sys
+import glob
+from ColumnDataFile import ColumnDataFile as CDF
+import matplotlib.pyplot as plt
+from pylab import *
+import numpy
+import operator
+import PlotUtility
+
+
+# translate a numeric value of an SO2 coordination to the equivalent text representation
+def CoordinationNumberToName (number):
+	name = ''
+	for i in range(int(number) / 10):
+		name = name + 'S'
+	for i in range(int(number) % 10):
+		name = name + 'O'
+
+	if number == 0:
+		name = 'Unbound'
+
+	return name
+
+def ConvertNumbers(numbers):
+	#j = 261
+	#print j, 10*(int(j)/100) + (int(j)%10) + (int(j%100)/10)
+	return [10*(int(n)/100) + (int(n)%10) + (int(n%100)/10) for n in numbers]
+
+def ExtractCoords(files):
+	cdfs = [CDF(f) for f in files]
+	data = [c[0] for c in cdfs]
+
+	# convert the abc number to ab
+	# a = number of S bonds
+	# b,c = numbers of O bonds
+	# this converts abc and returns a(b+c)
+	data = [ConvertNumbers(d) for d in data]
+	data = reduce(operator.add, data)
+	# this next line sums the total coordination
+	#data = [(int(i) % 10) + (int(i)/10) for i in data] 
+
+	return data
+	
+
+def PrintStats (name, data):
+
+	print name
+	data = numpy.array(data)
+
+	total = float(data.sum())
+	print total
+	# Un O OO OOO S SO SOO SOOO SS SSO SSOO SSOOO
+	print 'Unbound = ', data[0]/total*100, '%'
+	print 'O = ', data[1]/total*100, '%'
+	print 'OO = ', data[2]/total*100, '%'
+	print 'OOO = ', data[3]/total*100, '%'
+	print 'All O = ', (data[1] + data[2] + data[3])/total*100, '%'
+	print 'S = ', data[4]/total*100, '%'
+	print 'SO = ', data[5]/total*100, '%'
+	print 'SOO = ', data[6]/total*100, '%'
+	print 'SOOO = ', data[7]/total*100, '%'
+	print 'All S = ', (data[4] + data[5] + data[6] + data[7])/total*100, '%'
+
+new_bins = []
+num_bins = 4
+for i in range(num_bins):
+  for j in range(num_bins):
[…]

### 2011-06-30 images/scripts/PlotCyclicStructureDistributions.py
New image and script for the % breakdown by type of cyclic structure for at least SO-coordinated so2s.

@@ -1,65 +1,10 @@
 import sys
-import glob
-from ColumnDataFile import ColumnDataFile as CDF
 import matplotlib.pyplot as plt
 from pylab import *
 import numpy
 import operator
 import PlotUtility
 
-
-# translate a numeric value of an SO2 coordination to the equivalent text representation
-def CoordinationNumberToName (number):
-	name = ''
-	for i in range(int(number) / 10):
-		name = name + 'S'
-	for i in range(int(number) % 10):
-		name = name + 'O'
-
-	if number == 0:
-		name = 'Unbound'
-
-	return name
-
-def ConvertNumbers(numbers):
-	#j = 261
-	#print j, 10*(int(j)/100) + (int(j)%10) + (int(j%100)/10)
-	return [10*(int(n)/100) + (int(n)%10) + (int(n%100)/10) for n in numbers]
-
-def ExtractCoords(files):
-	cdfs = [CDF(f) for f in files]
-	data = [c[0] for c in cdfs]
-
-	# convert the abc number to ab
-	# a = number of S bonds
-	# b,c = numbers of O bonds
-	# this converts abc and returns a(b+c)
-	data = [ConvertNumbers(d) for d in data]
-	data = reduce(operator.add, data)
-	# this next line sums the total coordination
-	#data = [(int(i) % 10) + (int(i)/10) for i in data] 
-
-	return data
-	
-
-def PrintStats (name, data):
-
-	print name
-	data = numpy.array(data)
-
-	total = float(data.sum())
-	print total
-	# Un O OO OOO S SO SOO SOOO SS SSO SSOO SSOOO
-	print 'Unbound = ', data[0]/total*100, '%'
-	print 'O = ', data[1]/total*100, '%'
-	print 'OO = ', data[2]/total*100, '%'
-	print 'OOO = ', data[3]/total*100, '%'
-	print 'All O = ', (data[1] + data[2] + data[3])/total*100, '%'
-	print 'S = ', data[4]/total*100, '%'
-	print 'SO = ', data[5]/total*100, '%'
-	print 'SOO = ', data[6]/total*100, '%'
-	print 'SOOO = ', data[7]/total*100, '%'
-	print 'All S = ', (data[4] + data[5] + data[6] + data[7])/total*100, '%'
 
 new_bins = []
 num_bins = 4
[…]

### 2011-06-30 coordination.tex
Added more coordination images

@@
 For example, an 'SOO' coordinated \suldiox~forms three bonds to waters: two bonds are formed through the \suldiox-oxygens, and one through the \suldiox~sulfur.
-This naming scheme is taken from the work of Baer et al.
-where the \suldiox~bonding distribution was determined for a surface molecule at a single temperature simulated under similar conditions.\cite{Baer2010}
+This naming scheme is taken from a similar convention used for describing water bonding interactions, but applied now to the \suldiox~molecule.\cite{Buch2005}
 
 Figure \ref{fig:so2-bonding-coordinations} shows the distribution of bonding coordinations throughout the set of simulated trajectories for both the cold and hot temperatures.
+The two most common coordination types are the S and SO, followed by a completely unbound \suldiox~molecule.
+Bonding is primarily found through the \suldiox-sulfur atom, with higher coordinationsThe distribution shows that the \suldiox~mostly forms one or two bonding interactions with the waters.
+The bonding of the \suldiox~at both low and high temperatures follows similar trends, however the higher temperature shifts the distribution of coordinations slightly to forming more bonds than the cold temperature \suldiox.
+Figure \ref{fig:coordination-breakdown} breaks down the coordination distribution to show more details of the \suldiox~interactions.
+
+The total number of bonds formed to the \suldiox~are shown in the top plot of figure \ref{fig:coordination-breakdown}.
+The \suldiox~spends more than 70\% of the trajectories forming single or double interactions with surface waters.
+The center and bottom plots of figure \ref{fig:coordination-breakdown} show the distribution of bonds formed through the \suldiox~sulfur and oxygen atoms, respectively.
+The sulfur interaction distribution (center) shows the \suldiox~clearly spending most of the simulations forming a single bond through the sulfur atom.
+This supports the overall distributions heavily weighted by the S and SO coordination types.
+Looking at the oxygen interaction distribution plot (bottom) of figure \ref{fig:coordination-breakdown}, the oxygen remains mostly unbound or singly bound.
+By increasing the temperature the amount of unbound oxygen coordinations decreases 10\%, but the increase in the singly bound state is only by 4\%, while the double oxygen interactions increase by 6\% over the colder temperature distribution.
+Higher temperatures shift the distribution of interactions with \suldiox~to higher coordination types.
+
+The sfg experiment performed by our group on \suldiox~surface adsorption suggested an SO coordination type based on...\cite{Tarbuck2005,Tarkbuch2006} This was further supported by DFT simulations of single surface \suldiox~molecules bound to water at room temperature.\cite{Baer2010} Our simulations show that the SO interaction type is nearly as prevalent as the single S-type coordination at both temperatures.
+The colder temperature favors the single bond to the \suldiox~through the sulfur atom more than the high temperature, but the increase in coordination number of the \suldiox~is subtle and does not drastically change with temperature.
+
+% Figure showing all the different types of so2 coordinations
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/coordination/so2-bonding-coordinations.png}
+		\caption{Distribution of bonding coordinations of the surface \suldiox~molecule.
+Each set of bars represents the amount of time spent in the particular coordination type, including the unbound coordination.
+The results for the cold (blue) and hot (red) temperatures are shown together.
+At the high temperature the \suldiox~bonding becomes more complex with more bonds to the sulfur and oxygen atoms.
+The lower coordination states are more dominated in the cold temperature.
+The naming scheme is similar to that of a previous scheme for water coordinations,\cite{Buch2005} applied to the \suldiox~molecule.}
+		\label{fig:so2-bonding-coordinations}
+	\end{center}
+\end{figure}
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/coordination/so2-bonding-coordinations.png}
-		\caption{}
-		\label{fig:so2-bonding-coordinations
+		\includegraphics[scale=1.0]{images/coordination/coordination-breakdown.png}
+		\caption{Bonding coordinations of the \suldiox~broken down into specific interactions.
+The total number of bonds to the molecule (top) are shown regardless of coordination type.
+The total number of bonds through either the sulfur atom (center) or through both of the oxygens (bottom).
+The \suldiox~primarily forms one or two bonds to the surface waters.
+The sulfur atom mostly bonds to a single water oxygen.
+Between the two \suldiox~oxygen atoms, the molecule mostly does not bond to the water hydrogens, or forms a single hydrogen-bond.}
+		\label{fig:coordination-breakdown}
 	\end{center}

### 2011-07-21 Background.tex
partly finished the new coordination and temp change writeup.

\section {Background}

\subsection {Bonding Coordination}

A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water hydrogens, or interacts via the sulfur atoms with water oxygens. To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters, we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters. This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in other computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters. Thus a bonding coordination of ``OOH'' designates two proton-accepting bondind interactions through the water oxygen, and a single proton-donor bonding interaction through a hydrogen. More recently, Baer et al devised a nomenclature that explicitly enumerates the bonding to \suldiox~via the sulfur or oxygen atoms.\cite{Baer2010}

In this work we port Buch et al's nomenclature to \suldiox~in order to show the amount of hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions form from the \suldiox-sulfur to water oxygens. Thus, an ``SOO'' coordinated \suldiox~molecule forms a single iteraction through the sulfur atoom to a neighboring water oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox~oxygens. Analysis of the distribution of \suldiox~coordinations will give insight to how it binds to the water surface.

During the course of the analysis we use the definition of bonds of Baer et al. in determining the \suldiox~bonding coordinations.\cite{Baer2010} The definition is formed from a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox~oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.

\subsection {Cyclic Bonding Structures}

Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations congealing to a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayash1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of the surface hydrated \suldiox~by Baer et al. showed that the bonding coordination distributions of \suldiox~at the surface are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''. In the same work they then focused on the moste dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.

The cluster geometries found in Baer's study of the surface \suldiox~hydrates imply a cyclic bonding structure through the two or three waters involved. This is manifested as a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved. Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the \suldiox-sulfur and returning through the \suldiox-oxygen in a two-water cluster.

The optimized geometry of the \suldiox-hydrates suggest cyclic bonding structures, but it remains a different entirely when a \suldiox~is placed in a dynamics environment such as in the course of molecular dynamics simulations of an aqueous surface. Although geometry optimization shows the formation of these cyclic complexes, do they form during a dynamic bonding process on a simulated water surface? Graph theoretical techniques can be employed to determine if cyclic structures form during the simulations. Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998}

Here we will only briefly introduce graph theoretical concepts as they have been described well by others, and with varied application to cyclic structures.\cite{Tutte1984, Balakrishnan2000, Harary1973, Huber2007, Garcia2004, Dury2001} A graph consists of a set of vertexes, and edges that connect the vertexes. A molecule can be represented with atoms as vertexes, and edges for each intramolecular covalent bond connecting the atoms. The set of edges is then further expanded to include intermolecular interactions such as hydrogen bonds and other bonding bonding interactions. Edges may be assigned weights (i.e. bondlengths), types, and can be directional pointing towards a specific target from a source. A molecular system including all atoms, bonds, and interactions is thus fully described by a graph.

To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queueing vertexes and all adjacent neighboring vertexes while performing a specified procedure on each visited vertex. This is easily performed on adjaceny list or connectivity matrix data structures, iterating through atms of interest in the graph as starting points of the search. In BFS terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (grey), and visited nodes (black). Using a BFS on a graph, cyclic structures are detected any time a ``grey target'' is encountered when queueing adjacent neighbors of a vertex. A benefit of BFS on a graph is the ability to determine the smallest cyclic structure in which a given vertex is a member. In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox~sulfur as the root vertex will discover cyclic bonding structures in order of size. Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells aroun the \suldiox. Furthermore, it is possible to reconstruct a cycle's structure to find the size of the cycle (number of atoms involved), the number of unique waters in the cycle, and the type of cyclic structure encountered.

The various types of cyclic bonding arrangements are shown in Figure \ref{fig:cyclic-structures} for a \suldiox~molecule bonding to waters. Although three waters are shown in the molecular cartoon, any number of waters are allowed. Also, further bonding may occur to any of the participating molecules extraneous to the cyclic structure. Cycle types \Rmnum{2}, \Rmnum{3}, \Rmnum{4} in Figure \ref{fig:cyclic-structures} are cyclic structures in which the \suldiox~is a member of the cycle. Types \Rmnum{1} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle formed near the \suldiox. Type \Rmnum{3} is of particular interest because it is one of the most dominant \suldiox~bonding coordinations (``SO'', as shown later) configured as a cyclic structure.

Baer et al. presented a detailed geometric and spectroscopic breakdown of types \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varietes, shown in Figure \ref{fig:type-3-varieties}, that differ in the arrangement of the waters involved. Type \Rmnum{3}A is arranged with each water contributing an OH bond to the structure of the cycle. Type \Rmnum{3}B involves a single water contributing an OH, whereas the other two waters contribute only the oxygen atom or the entire water molecule to the structure, respectively. This nuance of the three-water type \Rmnum{3} structures, and the overall distribution are presented in more detail later. We also show the distribution of cyclic structures encountered during MD simulations to further understand the behaviors of \suldiox-hydrates at the water surface.

### 2011-07-21 Coordination-Results.tex
partly finished the new coordination and temp change writeup.

\section {Sulfur Dioxide Bonding Coordinations}

The bonding coordination of the surface \suldiox~was determined at each timestep of the simulations. Figure \ref{fig:bonding-coordinations} shows the distribution of coordinations of the surface \suldiox, by percent of simulation time, across all the simulations for both the cold (blue) and hot (red) trajectories. A first visual inspection clearly indicates the two most common coordinations: ``S'' and ``SO'', followed by the unbound coordination, and ``SOO''. This trend holds at both temperatures. Also, a \suldiox~configuration with bonds forming solely through the oxygens without any sulfur interactions, or any configuration with more than a single interaction the the sulfur atom are rare in comparison to the single sulfur coordinations. These less common coordinations likely form as transitory configurations when the waters briefly move closer or further away from the surface \suldiox.

The distribution paints a likely picture of the surface adsorption process of a \suldiox~first coming in contact with surface waters. Starting in an unbound state, two methods of initial contact are possible: through the \suldiox-sulfur of through one of the oxygens. Because the population of non-sulfur interactions (i.e. ``O'', ``OO'', ``OOO'', etc.) only account for 3\% of all the configurations sampled, either there is a very brief transition state where the \suldiox~first binds through an oxygen and then quickly adds, or switches to a sulfur interaction with the surface waters, or the first contact is made through the sulfur atom. Because the ``S'' coordination has such a large population, accounting for 35\% of the trajectories in the cold systems, and 32\% in the hot systems, it is most likely that the first binding contact to the waters is through the \suldiox~sulfur to a water oxygen. Furthermore, once in the ``S'' coordination, the \suldiox~can either unbind, add a second sulfur interaction, or bind through an oxygen. The histogram of Figure \ref{fig:bonding-coordinations} suggests that the preferred route is to further coordinate through an oxygen to achieve the ``SO'' coordination. ``SO'' accounts for 35\% of the cold coordinations, and 36\% of the hot trajectories. Binding to a second oxygen to the ``SOO'' coordination (7\% of the cold, 12\% of the hot systems) is less likely than losing an interaction through an oxygen back to the ``S'' coordination.

We thus have a likely route for adsorption and binding of an unbound \suldiox~to a water surface. The unbound \suldiox~initially binds mostly to a water through the sulfur atom, and then will either unbind, or add an oxygen interaction to oscillate between unbound, ``S'', and ``SO'' coordinations. It may then undergo brief sojourns to other coordinations because of the motions and rearrangements of the surface water molecules. These brief transitory coordinations are overwhelmed by the dominant ``S'' and ``SO'' states. The most probable deviation from those two coordinations is to the ``SOO'', where a second oxygen interaction is formed. Otherwsie the \suldiox~unbinds from the surface to repeat the adsorption process.

Baer et al performed this coordination analysis for their single-temperature study, but discriminating between \suldiox~binding through the two different oxygens. They concluded that there is assymetric hydrogen bonding through the \suldiox~oxygens, with one binding more often than the other. This is supported by the findings here where all the double oxygen coordinations (i.e. ``OO'', ``SOO'', etc.) represent a much lower percentage of the coordinations than the single oxygen counterparts (i.e. ``O'', ``SO'', etc.).

%To further assess the effect of temperature on \suldiox~behavior we turn to analysis of the bonds formed between \suldiox~and the water surface. The bonding coordinations were determined for the \suldiox~based on the bondlength criteria noted in the introduction. Bonds formed between \suldiox-sulfur and \wat-oxygen are labeled 'S' coordinated, and bonds formed between \suldiox-oxygen and \wat-hydrogen are labeled 'O' coordinated. For each bond an additional letter label is appended to the coordination. For example, an 'SOO' coordinated \suldiox~forms three bonds to waters: two bonds are formed through the \suldiox-oxygens, and one through the \suldiox~sulfur. This naming scheme is taken from a similar convention used for describing water bonding interactions, but applied now to the \suldiox~molecule.\cite{Buch2005}

%Figure \ref{fig:so2-bonding-coordinations} shows the distribution of bonding coordinations throughout the set of simulated trajectories for both the cold and hot temperatures. The two most common coordination types are the S and SO, followed by a completely unbound \suldiox~molecule. Bonding is primarily found through the \suldiox-sulfur atom, with higher coordinationsThe distribution shows that the \suldiox~mostly forms one or two bonding interactions with the waters. The bonding of the \suldiox~at both low and high temperatures follows similar trends, however the higher temperature shifts the distribution of coordinations slightly to forming more bonds than the cold temperature \suldiox. Figure \ref{fig:coordination-breakdown} breaks down the coordination distribution to show more details of the \suldiox~interactions.

%The total number of bonds formed to the \suldiox~are shown in the top plot of figure \ref{fig:coordination-breakdown}. The \suldiox~spends more than 70\% of the trajectories forming single or double interactions with surface waters. The center and bottom plots of figure \ref{fig:coordination-breakdown} show the distribution of bonds formed through the \suldiox~sulfur and oxygen atoms, respectively. The sulfur interaction distribution (center) shows the \suldiox~clearly spending most of the simulations forming a single bond through the sulfur atom. This supports the overall distributions heavily weighted by the S and SO coordination types. Looking at the oxygen interaction distribution plot (bottom) of figure \ref{fig:coordination-breakdown}, the oxygen remains mostly unbound or singly bound. By increasing the temperature the amount of unbound oxygen coordinations decreases 10\%, but the increase in the singly bound state is only by 4\%, while the double oxygen interactions increase by 6\% over the colder temperature distribution. Higher temperatures shift the distribution of interactions with \suldiox~to higher coordination types.

%The sfg experiment performed by our group on \suldiox~surface adsorption suggested an SO coordination type based on...\cite{Tarbuck2005,Tarkbuch2006} This was further supported by DFT simulations of single surface \suldiox~molecules bound to water at room temperature.\cite{Baer2010} Our simulations show that the SO interaction type is nearly as prevalent as the single S-type coordination at both temperatures. The colder temperature favors the single bond to the \suldiox~through the sulfur atom more than the high temperature, but the increase in coordination number of the \suldiox~is subtle and does not drastically change with temperature.

% Figure showing all the different types of so2 coordinations
\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/coordination/so2-bonding-coordinations.png}
		\caption{Distribution of bonding coordinations of the surface \suldiox~molecule. Each set of bars represents the amount of time spent in the particular coordination type, including the unbound coordination. The results for the cold (blue) and hot (red) temperatures are shown together. At the high temperature the \suldiox~bonding becomes more complex with more bonds to the sulfur and oxygen atoms. The lower coordination states are more dominated in the cold temperature. The naming scheme is similar to that of a previous scheme for water coordinations,\cite{Buch2005} applied to the \suldiox~molecule.}
		\label{fig:so2-bonding-coordinations}
	\end{center}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/coordination/coordination-breakdown.png}
		\caption{Bonding coordinations of the \suldiox~broken down into specific interactions. The total number of bonds to the molecule (top) are shown regardless of coordination type. The total number of bonds through either the sulfur atom (center) or through both of the oxygens (bottom). The \suldiox~primarily forms one or two bonds to the surface waters. The sulfur atom mostly bonds to a single water oxygen. Between the two \suldiox~oxygen atoms, the molecule mostly does not bond to the water hydrogens, or forms a single hydrogen-bond.}
		\label{fig:coordination-breakdown}
	\end{center}
\end{figure}

\subsection {Temperature Effects on Bonding Coordinations}

The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot. In the cold temperature, the unbound, single oxygen, and ``S'' coordinations are more populated than in the hot systems. The increased temperature causes all the coordinations with a single sulfur interaction or more (with the exception of ``S'') to increase over the equivalent cold temperature coordination populations. We see that the cold \suldiox~spends nearly twice as much time unbound, and only slightly more time in the ``S'' coordination. Most of the unbound population in the hot simulations appears to have shifted to the states with double-oxygen and double-sulfur coordinations. Figure \ref{fig:coordination-breakdown} shows the distribution of the number of bonds through the \suldiox~oxygens (left) and through the \suldiox~sulfur (right). When comparing from cold to hot distributions, the greatest shift occurs in the number of \suldiox~oxygen bonds as the hot \suldiox~has more multiply-bound oxygens (double and triple interactions). The increased temperature allows the \suldiox~oxygens to bind to more water hydrogens than in the cold systems.

These increased

### 2011-07-21 Background.tex
Finished writing out the hand-notes for coordinations + rdfs, and the background for coordinations and cycles

@@
 This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in other computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters.
-Thus a bonding coordination of ``OOH'' designates two proton-accepting bondind interactions through the water oxygen, and a single proton-donor bonding interaction through a hydrogen.
+Thus a bonding coordination of ``OOH'' designates two proton-accepting bonding interactions through the water oxygen, and a single proton-donor bonding interaction through a hydrogen.
 More recently, Baer et al devised a nomenclature that explicitly enumerates the bonding to \suldiox~via the sulfur or oxygen atoms.\cite{Baer2010}
@@
 In this work we port Buch et al's nomenclature to \suldiox~in order to show the amount of hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions form from the \suldiox-sulfur to water oxygens.
-Thus, an ``SOO'' coordinated \suldiox~molecule forms a single iteraction through the sulfur atoom to a neighboring water oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox~oxygens.
+Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox~oxygens.
 Analysis of the distribution of \suldiox~coordinations will give insight to how it binds to the water surface.
@@
 
-Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations congealing to a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayash1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of the surface hydrated \suldiox~by Baer et al.
+Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations congealing to a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of the surface hydrated \suldiox~by Baer et al.
 showed that the bonding coordination distributions of \suldiox~at the surface are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''.
-In the same work they then focused on the moste dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.
+In the same work they then focused on the most dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.
 
@@
 Edges may be assigned weights (i.e.
-bondlengths), types, and can be directional pointing towards a specific target from a source.
+bond lengths), types, and can be directional pointing towards a specific target from a source.
 A molecular system including all atoms, bonds, and interactions is thus fully described by a graph.
 
-To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queueing vertexes and all adjacent neighboring vertexes while performing a specified procedure on each visited vertex.
-This is easily performed on adjaceny list or connectivity matrix data structures, iterating through atms of interest in the graph as starting points of the search.
-In BFS terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (grey), and visited nodes (black).
-Using a BFS on a graph, cyclic structures are detected any time a ``grey target'' is encountered when queueing adjacent neighbors of a vertex.
+To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queuing vertexes and all adjacent neighboring vertexes while performing a specified procedure on each visited vertex.
+This is easily performed on adjacency list or connectivity matrix data structures, iterating through atoms of interest in the graph as starting points of the search.
+In BFS terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black).
+Using a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a vertex.
 A benefit of BFS on a graph is the ability to determine the smallest cyclic structure in which a given vertex is a member.
 In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox~sulfur as the root vertex will discover cyclic bonding structures in order of size.
-Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells aroun the \suldiox.
+Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox.
 Furthermore, it is possible to reconstruct a cycle's structure to find the size of the cycle (number of atoms involved), the number of unique waters in the cycle, and the type of cyclic structure encountered.
@@
 Baer et al.
-presented a detailed geometric and spectroscopic breakdown of types \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varietes, shown in Figure \ref{fig:type-3-varieties}, that differ in the arrangement of the waters involved.
+presented a detailed geometric and spectroscopic breakdown of types \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the arrangement of the waters involved.
 Type \Rmnum{3}A is arranged with each water contributing an OH bond to the structure of the cycle.

### 2011-07-21 Coordination-Results.tex
Finished writing out the hand-notes for coordinations + rdfs, and the background for coordinations and cycles

@@
 The most probable deviation from those two coordinations is to the ``SOO'', where a second oxygen interaction is formed.
-Otherwsie the \suldiox~unbinds from the surface to repeat the adsorption process.
+Otherwise the \suldiox~unbinds from the surface to repeat the adsorption process.
 
 Baer et al performed this coordination analysis for their single-temperature study, but discriminating between \suldiox~binding through the two different oxygens.
-They concluded that there is assymetric hydrogen bonding through the \suldiox~oxygens, with one binding more often than the other.
+They concluded that there is asymmetric hydrogen bonding through the \suldiox~oxygens, with one binding more often than the other.
 This is supported by the findings here where all the double oxygen coordinations (i.e.
@@
 % Figure showing all the different types of so2 coordinations
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{images/coordination/so2-bonding-coordinations.png}
-		\caption{Distribution of bonding coordinations of the surface \suldiox~molecule.
+%\begin{figure}[h!]
+	%\begin{center}
+		%\includegraphics[scale=1.0]{images/coordination/so2-bonding-coordinations.png}
+		%\caption{Distribution of bonding coordinations of the surface \suldiox~molecule.
 Each set of bars represents the amount of time spent in the particular coordination type, including the unbound coordination.
@@
 The naming scheme is similar to that of a previous scheme for water coordinations,\cite{Buch2005} applied to the \suldiox~molecule.}
-		\label{fig:so2-bonding-coordinations}
-	\end{center}
-\end{figure}
+		%\label{fig:so2-bonding-coordinations}
+	%\end{center}
+%\end{figure}
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{images/coordination/coordination-breakdown.png}
-		\caption{Bonding coordinations of the \suldiox~broken down into specific interactions.
+%\begin{figure}[h!]
+	%\begin{center}
+		%\includegraphics[scale=1.0]{images/coordination/coordination-breakdown.png}
+		%\caption{Bonding coordinations of the \suldiox~broken down into specific interactions.
 The total number of bonds to the molecule (top) are shown regardless of coordination type.
@@
 Between the two \suldiox~oxygen atoms, the molecule mostly does not bond to the water hydrogens, or forms a single hydrogen-bond.}
-		\label{fig:coordination-breakdown}
-	\end{center}
-\end{figure}
+		%\label{fig:coordination-breakdown}
+	%\end{center}
+%\end{figure}
 
@@
 
-These increased
+These increased binding through the \suldiox~oxygens at the hot temperatures should have a visible effect on the radial distribution function (RDF) between \suldiox-oxygen and water-hydrogen.
+Figure \ref{fig:rdf} shows the RDF of the \suldiox~sulfur and oxygen atoms with their water bonding partners, water oxygen and hydrogen, respectively, at both cold and hot temperatures.
+Both the first and second peaks of the O$_{SO_2}$-H$_{H_2O}$ RDF increase at the higher temperature, indicating a large number of water hydrogens near the \suldiox~oxygens in both hydration shells.
+This corroborates the increased coordination through the \suldiox~oxygens to the surface waters at the higher temperature.
+
+The populations of the single-oxygen coordinations are similar at both temperatures, somewhat contradicting the results of the RDF.
+We expect an increase of all \suldiox-oxygen interactions with surface water hydrogens, but the non-sulfur coordinations do not obey this trend.
+This indicates that the non-sulfur coordinations do not behave as most of the other coordinations, and their small contribution to the overall population in the distribution of Figure \ref{fig:bonding-coordinations} indicates that they are anomalous, or only briefly encountered during the simulations.
+
+The same temperature effect of shifting population to higher coordinations also manifests in the \suldiox~sulfur.
+The single-``S'' coordinations have similar populations at both temperatures, but the change from cold to hot temperature shifts the unbound \suldiox~coordinations up towards to the doubly-bound sulfur coordinations (i.e.
+``SS'', ``SSO'', etc.).
+The increased temperature causes more binding to the \suldiox~by allowing for multiple bonds through both the \suldiox~sulfur, and through the oxygens.
+Thus, the hotter temperature allows for increased coordination of the \suldiox~to the water surface.
+
+When the \suldiox~is on the cold water surface, the ``S'' and ``SO'' coordination populations are nearly equally represented.
+As the temperature increases to the hot system, the ``S'' coordination decreases, while the ``SO'' increases.
+We see from the RDFs that colder temperatures favor binding through the \suldiox~sulfur because of the larger first peak in the S$_{SO_2}$-O$_{H_2O}$ RDF.
+From this increase we expect the ``S''-coordinated \suldiox~to be more prevalent at cold temperatures, as is the case for the coordination distribution.
+However, the O$_{SO_2}$-H$_{H_2O}$ RDF indicates increased hydrogen bonding through the \suldiox~oxygens at higher temperatures manifested by a larger first peak in Figure \ref{fig:rdf} for the hot temperature.
+Thus the minimally ``SO''-coordinated distributions (those with at least a single bond to an oxygen and the sulfur, i.e.
+``SO'', ``SSO'', ``SSOO'', etc.) are more prevalent at the high temperature.
+Overall, we find evidence from both the bonding coordination distribution, and from the RDFs to suggest a modified bonding behavior at different temperatures.
+At an increased temperature the surface \suldiox~binds more through its oxygens than the cold counterpart.
+However, like the \suldiox~in the cold temperature, it likely has a persistent interaction through the sulfur atom to a neighboring water oxygen throughout most of the simulated trajectories.

### 2011-07-26 Background.tex
Filling in the images

@@
 
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/double-cycle-type2-small.png}
+		\caption{}
+		\label{fig:cyclic-example}
+	\end{center}
+\end{figure}
+
 The optimized geometry of the \suldiox-hydrates suggest cyclic bonding structures, but it remains a different entirely when a \suldiox~is placed in a dynamics environment such as in the course of molecular dynamics simulations of an aqueous surface.
@@
 Also, further bonding may occur to any of the participating molecules extraneous to the cyclic structure.
-Cycle types \Rmnum{2}, \Rmnum{3}, \Rmnum{4} in Figure \ref{fig:cyclic-structures} are cyclic structures in which the \suldiox~is a member of the cycle.
-Types \Rmnum{1} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle formed near the \suldiox.
+Cycle types \Rmnum{1}, \Rmnum{2}, \Rmnum{3} in Figure \ref{fig:cyclic-structures} are cyclic structures in which the \suldiox~is a member of the cycle.
+Types \Rmnum{4} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle formed near the \suldiox.
 Type \Rmnum{3} is of particular interest because it is one of the most dominant \suldiox~bonding coordinations (``SO'', as shown later) configured as a cyclic structure.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/cycle-types-small.png}
+		\caption{}
+		\label{fig:cyclic-structures}
+	\end{center}
+\end{figure}
 
@@
 We also show the distribution of cyclic structures encountered during MD simulations to further understand the behaviors of \suldiox-hydrates at the water surface.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/triple-cycle-types-small.png}
+		\caption{}
+		\label{fig:type-3-varieties}
+	\end{center}
+\end{figure}

### 2011-07-26 Introduction.tex
Filling in the images

@@
 
-The chemistry of our atmosphere is vastly complex and poorly understood.
+%The chemistry of our atmosphere is vastly complex and poorly understood.
 Reactions taking place on particle surfaces and liquid boundaries dominate many atmospheric chemical pathways, and guide terrestrial and industrial byproducts to their ultimate forms and destinations.

### 2011-07-26 cycle-background.tex
Filling in the images

\subsection{Cyclic Hydrate Complexes}

During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule. Computationally, there are various methods to define characteristics of hydrate complex formation.\cite{} A simple metric of molecular complexation is a count of the number of bonds formed with neighboring molecules.

A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox. However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form. In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries. Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works. As shown in Figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.

%We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems. The first step in the determination of cycles is translating the atoms (vertices) and bonds (edges) to a graph. Figure \ref{fig:so2-complex-graph} shows the procedure visually, where here the root vertex of the graph is the \suldiox-sulfur, and each atom is numbered by the distance, in atoms, to the root vertex. We use only the closest neighboring water molecules that may take part in a cyclic complex in building the graph. After forming the graph, a breadth-first search (BFS) is performed starting at the root vertex.\cite{Knuth1997} The BFS algorithm uses a coloring technique where all unexamined vertices are colored white, queued vertices are gray, and examined vertices are black. During the BFS, encountering a gray vertex as a target of an adjacent edge during examination indicates that a cycle has been found in the graph. At each vertex dequeue for examination, the parent of every vertex is recorded such that the lineage from gray target to root vertex can be reconstructed to determine the participating atoms in the cycle.

%\begin{figure}[h!]
%\begin{center}
%\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
%\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges. The bonds are all given equal weights and the graph is undirected. The numbering shows the relative distance from any atom to the root node (\suldiox-sulfur 1), and also enumerates the iteration through the BFS process at which the atom is discovered. A BFS on this graph would result in discovering the doubly-bound hydrogen (5) as a gray target, and the gray source would be either of the connected oxygens (4). At each iteration the predecessor in the BFS is recorded in order to reconstruct the connectivity of the cycle. This cycle is composed of 8 atoms, and 3 molecules}
%\label{fig:so2-complex-graph}
%\end{center}
%\end{figure}

### 2011-07-26 Background.tex
Finished adding images for rough draft round 1

@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/double-cycle-type2-small.png}
+		\includegraphics[scale=1.0]{images/cycles/double-cycle-type2-small.png}
 		\caption{}
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/cycle-types-small.png}
+		\includegraphics[scale=1.0]{images/cycles/cycle-types-small.png}
 		\caption{}
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/triple-cycle-types-small.png}
+		\includegraphics[scale=1.0]{images/cycles/triple-cycle-types-small.png}
 		\caption{}

### 2011-07-26 Coordination-Results.tex
Filling in the images
Finished adding images for rough draft round 1

@@
 Binding to a second oxygen to the ``SOO'' coordination (7\% of the cold, 12\% of the hot systems) is less likely than losing an interaction through an oxygen back to the ``S'' coordination.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/coordinations/so2-coordinations-percents-small.png}
+		\caption{}
+		\label{fig:bonding-coordinations}
+	\end{center}
+\end{figure}
 
@@
 Most of the unbound population in the hot simulations appears to have shifted to the states with double-oxygen and double-sulfur coordinations.
-Figure \ref{fig:coordination-breakdown} shows the distribution of the number of bonds through the \suldiox~oxygens (left) and through the \suldiox~sulfur (right).
-When comparing from cold to hot distributions, the greatest shift occurs in the number of \suldiox~oxygen bonds as the hot \suldiox~has more multiply-bound oxygens (double and triple interactions).
+Figure \ref{fig:coordination-breakdown} shows the distribution of the number of bonds through the \suldiox~sulfur (top) and through the \suldiox~oxygens (bottom).
+When comparing from cold to hot distributions, the greatest shift occurs in the number of \suldiox~oxygen bonds as the hot \suldiox~has more bonding to the oxygens.
 The increased temperature allows the \suldiox~oxygens to bind to more water hydrogens than in the cold systems.
 
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/coordinations/coordination-breakdown-small.png}
+		\caption{}
+		\label{fig:coordination-breakdown}
+	\end{center}
+\end{figure}
+
 These increased binding through the \suldiox~oxygens at the hot temperatures should have a visible effect on the radial distribution function (RDF) between \suldiox-oxygen and water-hydrogen.
-Figure \ref{fig:rdf} shows the RDF of the \suldiox~sulfur and oxygen atoms with their water bonding partners, water oxygen and hydrogen, respectively, at both cold and hot temperatures.
+Figure \ref{fig:rdf} shows the RDF of the \suldiox~sulfur and oxygen atoms with their water bonding partners, water oxygen and hydrogen, respectively, at both cold (blue) and hot (green) temperatures.
 Both the first and second peaks of the O$_{SO_2}$-H$_{H_2O}$ RDF increase at the higher temperature, indicating a large number of water hydrogens near the \suldiox~oxygens in both hydration shells.
 This corroborates the increased coordination through the \suldiox~oxygens to the surface waters at the higher temperature.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/rdf/rdf-small.png}
+		\caption{}
+		\label{fig:rdf}
+	\end{center}
+\end{figure}
 

### 2011-07-26 Background.tex
Minor edits before sending to Baer

@@
 
-The optimized geometry of the \suldiox-hydrates suggest cyclic bonding structures, but it remains a different entirely when a \suldiox~is placed in a dynamics environment such as in the course of molecular dynamics simulations of an aqueous surface.
+The optimized geometry of the \suldiox-hydrates suggest cyclic bonding structures, but it remains a different story entirely when a \suldiox~is placed in a dynamic environment such as in the course of molecular dynamics simulations of an aqueous surface.
 Although geometry optimization shows the formation of these cyclic complexes, do they form during a dynamic bonding process on a simulated water surface?
@@
 Baer et al.
-presented a detailed geometric and spectroscopic breakdown of types \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the arrangement of the waters involved.
-Type \Rmnum{3}A is arranged with each water contributing an OH bond to the structure of the cycle.
+presented a detailed geometric and spectroscopic breakdown of type \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the arrangement of the waters involved.
+Type \Rmnum{3}-A (shown in Figure \ref{fig:type-3-varieties}A) is arranged with each water contributing an OH bond to the structure of the cycle.
 Type \Rmnum{3}B involves a single water contributing an OH, whereas the other two waters contribute only the oxygen atom or the entire water molecule to the structure, respectively.

### 2011-07-26 Coordination-Results.tex
Minor edits before sending to Baer

@@
 The bonding coordination of the surface \suldiox~was determined at each timestep of the simulations.
-Figure \ref{fig:bonding-coordinations} shows the distribution of coordinations of the surface \suldiox, by percent of simulation time, across all the simulations for both the cold (blue) and hot (red) trajectories.
+Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, by percentage of all bonding coordinations encountered, across all the simulations for both the cold (blue) and hot (red) trajectories.
 A first visual inspection clearly indicates the two most common coordinations: ``S'' and ``SO'', followed by the unbound coordination, and ``SOO''.
 This trend holds at both temperatures.
-Also, a \suldiox~configuration with bonds forming solely through the oxygens without any sulfur interactions, or any configuration with more than a single interaction the the sulfur atom are rare in comparison to the single sulfur coordinations.
+Also, a \suldiox~configuration with bonds forming solely through the oxygens without any sulfur interactions, or any configuration with more than a single interaction through the sulfur atom are rare in comparison to the single sulfur coordinations.
 These less common coordinations likely form as transitory configurations when the waters briefly move closer or further away from the surface \suldiox.
@@
 The distribution paints a likely picture of the surface adsorption process of a \suldiox~first coming in contact with surface waters.
-Starting in an unbound state, two methods of initial contact are possible: through the \suldiox-sulfur of through one of the oxygens.
-Because the population of non-sulfur interactions (i.e.
+Starting in an unbound state, two methods of initial contact are possible: through the \suldiox-sulfur or through one of the oxygens.
+Because the population of non-sulfur bonding interactions (not including the ``unbound'' state, i.e.
 ``O'', ``OO'', ``OOO'', etc.) only account for 3\% of all the configurations sampled, either there is a very brief transition state where the \suldiox~first binds through an oxygen and then quickly adds, or switches to a sulfur interaction with the surface waters, or the first contact is made through the sulfur atom.
-Because the ``S'' coordination has such a large population, accounting for 35\% of the trajectories in the cold systems, and 32\% in the hot systems, it is most likely that the first binding contact to the waters is through the \suldiox~sulfur to a water oxygen.
+The ``S'' coordination has a large population, accounting for 35\% of the trajectories in the cold systems, and 32\% in the hot systems.
+Thus, it is most likely that the first binding contact to the waters is through the \suldiox~sulfur to a water oxygen.
 Furthermore, once in the ``S'' coordination, the \suldiox~can either unbind, add a second sulfur interaction, or bind through an oxygen.
@@
 
-Baer et al performed this coordination analysis for their single-temperature study, but discriminating between \suldiox~binding through the two different oxygens.
+Baer et al performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens.
 They concluded that there is asymmetric hydrogen bonding through the \suldiox~oxygens, with one binding more often than the other.
@@
 
-The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot.
+The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in bonding coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot.
 In the cold temperature, the unbound, single oxygen, and ``S'' coordinations are more populated than in the hot systems.
-The increased temperature causes all the coordinations with a single sulfur interaction or more (with the exception of ``S'') to increase over the equivalent cold temperature coordination populations.
-We see that the cold \suldiox~spends nearly twice as much time unbound, and only slightly more time in the ``S'' coordination.
-Most of the unbound population in the hot simulations appears to have shifted to the states with double-oxygen and double-sulfur coordinations.
+We see that the cold \suldiox~spends nearly twice as much time unbound, and only slightly more time in the ``S'' coordination than in the hot system.
+The increased temperature of the hot systems causes all the coordinations with more than a single sulfur interaction (``SS'', ``SSO'', etc.) to increase over the equivalent cold temperature coordination populations.
+Most of the unbound population in the hot simulations appears to have shifted to the states with double-oxygen or double-sulfur coordinations.
 Figure \ref{fig:coordination-breakdown} shows the distribution of the number of bonds through the \suldiox~sulfur (top) and through the \suldiox~oxygens (bottom).
-When comparing from cold to hot distributions, the greatest shift occurs in the number of \suldiox~oxygen bonds as the hot \suldiox~has more bonding to the oxygens.
-The increased temperature allows the \suldiox~oxygens to bind to more water hydrogens than in the cold systems.
+When comparing from cold to hot distributions, the change in temperature increases the number of bonds to the \suldiox~sulfur and oxygen atoms.
+Whereas in the cold system the \suldiox~very rarely forms two interactions through the sulfur, in the hot system we see that the two-sulfur interaction bonding coordination account for nearly 5\% of all coordinations.
+There is a similar increase in bonding to the \suldiox~oxygens in the hot system.
+The increased temperature of the hot systems allows the \suldiox to bind to more surface waters than in the cold systems.
 
@@
 
-These increased binding through the \suldiox~oxygens at the hot temperatures should have a visible effect on the radial distribution function (RDF) between \suldiox-oxygen and water-hydrogen.
+The increased binding through the \suldiox~oxygens at the hot temperatures should have a visible effect on the radial distribution function (RDF) between \suldiox-oxygen and water-hydrogen.
 Figure \ref{fig:rdf} shows the RDF of the \suldiox~sulfur and oxygen atoms with their water bonding partners, water oxygen and hydrogen, respectively, at both cold (blue) and hot (green) temperatures.
-Both the first and second peaks of the O$_{SO_2}$-H$_{H_2O}$ RDF increase at the higher temperature, indicating a large number of water hydrogens near the \suldiox~oxygens in both hydration shells.
+Both the first and second peaks of the O$_{SO_2}$-H$_{H_2O}$ RDF increase at the higher temperature, indicating a higher number of water hydrogens near the \suldiox~oxygens in both hydration shells.
 This corroborates the increased coordination through the \suldiox~oxygens to the surface waters at the higher temperature.
@@
 The populations of the single-oxygen coordinations are similar at both temperatures, somewhat contradicting the results of the RDF.
-We expect an increase of all \suldiox-oxygen interactions with surface water hydrogens, but the non-sulfur coordinations do not obey this trend.
+We expect an increase of all \suldiox-oxygen interactions with surface water hydrogens at the higher temperature, but the non-sulfur coordinations do not obey this trend.
 This indicates that the non-sulfur coordinations do not behave as most of the other coordinations, and their small contribution to the overall population in the distribution of Figure \ref{fig:bonding-coordinations} indicates that they are anomalous, or only briefly encountered during the simulations.
 
-The same temperature effect of shifting population to higher coordinations also manifests in the \suldiox~sulfur.
-The single-``S'' coordinations have similar populations at both temperatures, but the change from cold to hot temperature shifts the unbound \suldiox~coordinations up towards to the doubly-bound sulfur coordinations (i.e.
+The same temperature effect of shifting population to higher bonding coordinations also manifests in the \suldiox~sulfur.
+The single-``S'' coordinations (``S'', ``SO'', ``SOO'', etc.) have similar populations at both temperatures, but the change from cold to hot temperature shifts the unbound \suldiox~coordinations up towards to the doubly-bound sulfur coordinations (i.e.
 ``SS'', ``SSO'', etc.).
@@
 However, the O$_{SO_2}$-H$_{H_2O}$ RDF indicates increased hydrogen bonding through the \suldiox~oxygens at higher temperatures manifested by a larger first peak in Figure \ref{fig:rdf} for the hot temperature.
-Thus the minimally ``SO''-coordinated distributions (those with at least a single bond to an oxygen and the sulfur, i.e.
+Thus, those \suldiox~bonding coordinations with at least a single bond to an oxygen and to the sulfur, (i.e.
 ``SO'', ``SSO'', ``SSOO'', etc.) are more prevalent at the high temperature.

### 2011-08-08 Coordination-Results.tex
New plots of the transition graphs, shrank the lifespans, added graphviz citations to the bibtex, and TONS OF TEXT about the coordination results (distro, transitions, lifespans).

@@
 The bonding coordination of the surface \suldiox~was determined at each timestep of the simulations.
-Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, by percentage of all bonding coordinations encountered, across all the simulations for both the cold (blue) and hot (red) trajectories.
-A first visual inspection clearly indicates the two most common coordinations: ``S'' and ``SO'', followed by the unbound coordination, and ``SOO''.
-This trend holds at both temperatures.
-Also, a \suldiox~configuration with bonds forming solely through the oxygens without any sulfur interactions, or any configuration with more than a single interaction through the sulfur atom are rare in comparison to the single sulfur coordinations.
-These less common coordinations likely form as transitory configurations when the waters briefly move closer or further away from the surface \suldiox.
-
-The distribution paints a likely picture of the surface adsorption process of a \suldiox~first coming in contact with surface waters.
-Starting in an unbound state, two methods of initial contact are possible: through the \suldiox-sulfur or through one of the oxygens.
-Because the population of non-sulfur bonding interactions (not including the ``unbound'' state, i.e.
-``O'', ``OO'', ``OOO'', etc.) only account for 3\% of all the configurations sampled, either there is a very brief transition state where the \suldiox~first binds through an oxygen and then quickly adds, or switches to a sulfur interaction with the surface waters, or the first contact is made through the sulfur atom.
-The ``S'' coordination has a large population, accounting for 35\% of the trajectories in the cold systems, and 32\% in the hot systems.
-Thus, it is most likely that the first binding contact to the waters is through the \suldiox~sulfur to a water oxygen.
-Furthermore, once in the ``S'' coordination, the \suldiox~can either unbind, add a second sulfur interaction, or bind through an oxygen.
-The histogram of Figure \ref{fig:bonding-coordinations} suggests that the preferred route is to further coordinate through an oxygen to achieve the ``SO'' coordination.
-``SO'' accounts for 35\% of the cold coordinations, and 36\% of the hot trajectories.
-Binding to a second oxygen to the ``SOO'' coordination (7\% of the cold, 12\% of the hot systems) is less likely than losing an interaction through an oxygen back to the ``S'' coordination.
+Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, by percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories.
+A first visual inspection reveals several trends in the bonding of the \suldiox.
+Clearly the ``SO'' coordination is the most populous at both temperatures.
+The second and third most populated coordinations are the ``S'' and ``SOO'', however the distributions differ between temperatures.
+In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally.
+The hotter temperature simulation shifts the distribution such that the ``S'' occurs 5\% less frequently than in the cold, and the ``SOO'' occurs nearly 10\% more times.
+The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``SOO'', respectively.
+These results for the hot (room temperature) system coincide with those of the previous single-temperature simulation study by Baer et al.
+at a similar temperature.\cite{Baer2010}
 
@@
 
-We thus have a likely route for adsorption and binding of an unbound \suldiox~to a water surface.
-The unbound \suldiox~initially binds mostly to a water through the sulfur atom, and then will either unbind, or add an oxygen interaction to oscillate between unbound, ``S'', and ``SO'' coordinations.
-It may then undergo brief sojourns to other coordinations because of the motions and rearrangements of the surface water molecules.
-These brief transitory coordinations are overwhelmed by the dominant ``S'' and ``SO'' states.
-The most probable deviation from those two coordinations is to the ``SOO'', where a second oxygen interaction is formed.
-Otherwise the \suldiox~unbinds from the surface to repeat the adsorption process.
+	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from the distribution of coordinations.
+Clearly the cold \suldiox~spends more time than the hot \suldiox~(10\%) completely unbound from the surface waters.
+This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria set for the analysis.
+Furthermore, the most frequently occuring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
+Baer et al.
+also concluded that these three coordinations werethe most frequent for a room temperature simulation, and specifically isolated the ``SO'' and ``SOO'' as most common in their study.\cite{Baer2010}
 
+	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (i.e.
+``O'', ``OO'', etc.) represent the least formed bonding coordinations.
+A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~oxygen-only bonding to waters.
+In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox-sulfur points further into the \wat~bulk than the oxygens.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, and to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
+
+% Baer's assymetric oxygen binding
 Baer et al performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens.
@@
 ``O'', ``SO'', etc.).
+Furthermore, a triple-oxygen coordination (i.e.
+``OOO'', ``SOOO'', etc.) is very rarely encountered.
+Three \suldiox~oxygen bonds only form if both \suldiox~oxygens are interacting with water hydrogens.
+Our finding that triple-oxygen coordinations rarely form complements Baer et al's conclusion about the assymetry in the oxygen interactions.
 
-%To further assess the effect of temperature on \suldiox~behavior we turn to analysis of the bonds formed between \suldiox~and the water surface.
-The bonding coordinations were determined for the \suldiox~based on the bondlength criteria noted in the introduction.
-Bonds formed between \suldiox-sulfur and \wat-oxygen are labeled 'S' coordinated, and bonds formed between \suldiox-oxygen and \wat-hydrogen are labeled 'O' coordinated.
-For each bond an additional letter label is appended to the coordination.
-For example, an 'SOO' coordinated \suldiox~forms three bonds to waters: two bonds are formed through the \suldiox-oxygens, and one through the \suldiox~sulfur.
-This naming scheme is taken from a similar convention used for describing water bonding interactions, but applied now to the \suldiox~molecule.\cite{Buch2005}
+Having established the preference for an interaction through the \suldiox~sulfur atom, we look to the right side of Figure \ref{fig:bonding-coordinations} at the double-sulfur coordinations (i.e.
+``SS'', ``SSO'', ``SSOO'', etc.).
+We see from the datat that single-sulfur coordination are overwhelmingly preferred over double-sulfur ones.
+Also, adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' coordinations are almost completely unrepresented (less than 1\%), and the change from ``SOO'' to ``SOOO'' produces a very large drop in the coordination distribution of Figure \ref{fig:bonding-coordinations}.
 
-%Figure \ref{fig:so2-bonding-coordinations} shows the distribution of bonding coordinations throughout the set of simulated trajectories for both the cold and hot temperatures.
-The two most common coordination types are the S and SO, followed by a completely unbound \suldiox~molecule.
-Bonding is primarily found through the \suldiox-sulfur atom, with higher coordinationsThe distribution shows that the \suldiox~mostly forms one or two bonding interactions with the waters.
-The bonding of the \suldiox~at both low and high temperatures follows similar trends, however the higher temperature shifts the distribution of coordinations slightly to forming more bonds than the cold temperature \suldiox.
-Figure \ref{fig:coordination-breakdown} breaks down the coordination distribution to show more details of the \suldiox~interactions.
-
-%The total number of bonds formed to the \suldiox~are shown in the top plot of figure \ref{fig:coordination-breakdown}.
-The \suldiox~spends more than 70\% of the trajectories forming single or double interactions with surface waters.
-The center and bottom plots of figure \ref{fig:coordination-breakdown} show the distribution of bonds formed through the \suldiox~sulfur and oxygen atoms, respectively.
-The sulfur interaction distribution (center) shows the \suldiox~clearly spending most of the simulations forming a single bond through the sulfur atom.
-This supports the overall distributions heavily weighted by the S and SO coordination types.
-Looking at the oxygen interaction distribution plot (bottom) of figure \ref{fig:coordination-breakdown}, the oxygen remains mostly unbound or singly bound.
-By increasing the temperature the amount of unbound oxygen coordinations decreases 10\%, but the increase in the singly bound state is only by 4\%, while the double oxygen interactions increase by 6\% over the colder temperature distribution.
-Higher temperatures shift the distribution of interactions with \suldiox~to higher coordination types.
-
-%The sfg experiment performed by our group on \suldiox~surface adsorption suggested an SO coordination type based on...\cite{Tarbuck2005,Tarkbuch2006} This was further supported by DFT simulations of single surface \suldiox~molecules bound to water at room temperature.\cite{Baer2010} Our simulations show that the SO interaction type is nearly as prevalent as the single S-type coordination at both temperatures.
-The colder temperature favors the single bond to the \suldiox~through the sulfur atom more than the high temperature, but the increase in coordination number of the \suldiox~is subtle and does not drastically change with temperature.
-
-% Figure showing all the different types of so2 coordinations
-%\begin{figure}[h!]
-	%\begin{center}
-		%\includegraphics[scale=1.0]{images/coordination/so2-bonding-coordinations.png}
-		%\caption{Distribution of bonding coordinations of the surface \suldiox~molecule.
-Each set of bars represents the amount of time spent in the particular coordination type, including the unbound coordination.
-The results for the cold (blue) and hot (red) temperatures are shown together.
-At the high temperature the \suldiox~bonding becomes more complex with more bonds to the sulfur and oxygen atoms.
-The lower coordination states are more dominated in the cold temperature.
-The naming scheme is similar to that of a previous scheme for water coordinations,\cite{Buch2005} applied to the \suldiox~molecule.}
-		%\label{fig:so2-bonding-coordinations}
-	%\end{center}
-%\end{figure}
-
-%\begin{figure}[h!]
-	%\begin{center}
-		%\includegraphics[scale=1.0]{images/coordination/coordination-breakdown.png}
-		%\caption{Bonding coordinations of the \suldiox~broken down into specific interactions.
-The total number of bonds to the molecule (top) are shown regardless of coordination type.
-The total number of bonds through either the sulfur atom (center) or through both of the oxygens (bottom).
-The \suldiox~primarily forms one or two bonds to the surface waters.
-The sulfur atom mostly bonds to a single water oxygen.
-Between the two \suldiox~oxygen atoms, the molecule mostly does not bond to the water hydrogens, or forms a single hydrogen-bond.}
-		%\label{fig:coordination-breakdown}
-	%\end{center}
-%\end{figure}
+We can now form a picture of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study.
+The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond assymetrically through one of the oxygens either once or twice to water hydrogens.
+The \suldiox~oxygen bonds will form and break throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.
 
@@
 The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in bonding coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot.
-In the cold temperature, the unbound, single oxygen, and ``S'' coordinations are more populated than in the hot systems.
-We see that the cold \suldiox~spends nearly twice as much time unbound, and only slightly more time in the ``S'' coordination than in the hot system.
-The increased temperature of the hot systems causes all the coordinations with more than a single sulfur interaction (``SS'', ``SSO'', etc.) to increase over the equivalent cold temperature coordination populations.
-Most of the unbound population in the hot simulations appears to have shifted to the states with double-oxygen or double-sulfur coordinations.
-Figure \ref{fig:coordination-breakdown} shows the distribution of the number of bonds through the \suldiox~sulfur (top) and through the \suldiox~oxygens (bottom).
-When comparing from cold to hot distributions, the change in temperature increases the number of bonds to the \suldiox~sulfur and oxygen atoms.
-Whereas in the cold system the \suldiox~very rarely forms two interactions through the sulfur, in the hot system we see that the two-sulfur interaction bonding coordination account for nearly 5\% of all coordinations.
-There is a similar increase in bonding to the \suldiox~oxygens in the hot system.
-The increased temperature of the hot systems allows the \suldiox to bind to more surface waters than in the cold systems.
+In the cold temperature, the unbound, ``S'', and ``SO'' coordinations are more populated than in the hot systems.
+The increased temperature decreases the time spent unbound to the surface waters, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature coordination populations.
+We see that the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to the stateswith double-oxygen and double-sulfur coordinations.
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{images/coordinations/coordination-breakdown-small.png}
-		\caption{}
-		\label{fig:coordination-breakdown}
-	\end{center}
-\end{figure}
-
-The increased binding through the \suldiox~oxygens at the hot temperatures should have a visible effect on the radial distribution function (RDF) between \suldiox-oxygen and water-hydrogen.
-Figure \ref{fig:rdf} shows the RDF of the \suldiox~sulfur and oxygen atoms with their water bonding partners, water oxygen and hydrogen, respectively, at both cold (blue) and hot (green) temperatures.
-Both the first and second peaks of the O$_{SO_2}$-H$_{H_2O}$ RDF increase at the higher temperature, indicating a higher number of water hydrogens near the \suldiox~oxygens in both hydration shells.
-This corroborates the increased coordination through the \suldiox~oxygens to the surface waters at the higher temperature.
+On the cold water surface, the ``S'' and ``SO'' are more populated than for the hot system.
+The relative decrease of these coordinations are matched in the hot surface by an increase of the ``SOO'' configuration.
+This speaks to a dramatic difference in the surface behavior of \suldiox~at the two temperatures.
+The cold \suldiox~spends nearly equal time between the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''.
+Thus the addition, or removal of a bond through the \suldiox-oxygen is equally probable, as long as the pivotal sulfur interaction remains intact.
+Figure \ref{fig:rdf} shows the radial distribution functions (RDF) of \suldiox~to water atoms for both cold (blue) and hot (red) temperatures.
+The \suldiox-S to \wat-O RDFs are nearly equal except for a slightly taller first peak of the cold system.
+Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF would indicate that with bonding mostly through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface waters.
+As the temperature increases in the hot system, the added bonding through a second oxygen interaction may be pulling the bonding waters further from the \suldiox-sulfur.
+This suggests that the cold \suldiox~bonds to water's surface more through its sulfur atom (as shown in the \suldiox-S RDF) and favors the more sulfur-centric bonding coordinations (i.e.
+``S'', ``SO'').
+The increased temperature allows the \suldiox~to bond more extensively through its oxygens to the ``SOO'' coordination.
+In turn, the increase binding through the oxygens may pull the \suldiox~further into the water interface and then allow for increased brief bonding through the sulfur, up to double-sulfur coordinations (i.e.
+``SS'', ``SSO'', etc.).
 
@@
 
-The populations of the single-oxygen coordinations are similar at both temperatures, somewhat contradicting the results of the RDF.
-We expect an increase of all \suldiox-oxygen interactions with surface water hydrogens at the higher temperature, but the non-sulfur coordinations do not obey this trend.
-This indicates that the non-sulfur coordinations do not behave as most of the other coordinations, and their small contribution to the overall population in the distribution of Figure \ref{fig:bonding-coordinations} indicates that they are anomalous, or only briefly encountered during the simulations.
+\subsection {Bonding Transitions}
 
-The same temperature effect of shifting population to higher bonding coordinations also manifests in the \suldiox~sulfur.
-The single-``S'' coordinations (``S'', ``SO'', ``SOO'', etc.) have similar populations at both temperatures, but the change from cold to hot temperature shifts the unbound \suldiox~coordinations up towards to the doubly-bound sulfur coordinations (i.e.
+During the course of each simulation, the \suldiox~bonding coordination was determined and recorded for each timestep.
+From the coordination data we extract not only the populations of the various bonding coordinations encountered, but also the frequencies of transitions between the different coordinations (i.e.
+the number of times an \suldiox~switched from one coordination type to another).
+With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} A and B, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one coordination to another.
+The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently).
+Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequently occuring transitions.
+Additionally, the transition lines are numbered to the right of each line with the number of times each transition occured.
+Only the most populated nodes and transitions between them are shown.
+This accounts for the difference of inbound and outbound nodes to certain nodes not equalling zero, as would happen for a complete graph representation of all coordinations and transitions.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/coordinations/coordination-transitions.png}
+		\caption{}
+		\label{fig:coordination-transitions}
+	\end{center}
+\end{figure}
+
+As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} than transitions to less populated coordinations.
+However, some insights to the bonding process are made clearer from these graphs.
+In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes.
+The number of transitions within this ``S-SO-SOO'' cluster of coordinations is an order of magnitude larger than any other transition.
+This indicates that while the \suldiox~is bound in any of the three most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other two coordinations in this group.
+The \suldiox~is rather active and constantly forming and breaking bonds to with its oxygens.
+
+In the graph of the hot system in Figure \ref{fig:coordination-transitions}B, the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size.
+One very surprising result is in the transition from ``SOO-SOOO''.
+The transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations.
+This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox~oxygen.
+As noted earlier, it is likely that the ``SOO'' \suldiox, assymetrically binding twice through a single oxygen, is pulled further into the water interface.
+It is likely more surrounded by waters, and in the hot system it can often form a brief third hydrogen bond to a water through the second \suldiox~oxygen.
+Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox returns back to the ``SOO'' coordination.
+
+Given the information we have about the frequencies of bonding coordination transitions, it is possible to draw a likely route of adsorption beginning with an unbound \suldiox.
+From the unbound state, the \suldiox~can bind to waters either through the sulfur or an oxygen to enter the ``S'' or ``O'' coordinations, respectively.
+At both temperatures we find that the coordination transition in Figure \ref{fig:coordination-transitions} from unbound to ``O'' occurs almost three times more than the transition from unbound to ``S''.
+Either the ``O'' state is more easily formed, but also quickly broken, accounting for the higher transition frequency, or the ``O'' coordination is in fact the first step in adsorption of the unbound \suldiox, leading to a transition to the most frequent coordination, ``SO'', through addition of an \suldiox~sulfur interaction to a water oxygen.
+
+To verify the claim that the ``O'' coordination forms from, and breaks quickly to the unbound state as is suggested by the transition frequency plots in Figure \ref{fig:coordination-transitions}, rather than enduring longer, we plot the lifespans of the various coordinations in Figure \ref{fig:coordination-lifespans}.
+Each point in the plot represent a time during the simulated trajectories in which the \suldiox~formed the respective coordination.
+The vertical ``lifespan'' position is calculated directly from the amount of time spent in the given coordination before changing to another.
+Both cold (blue) and hot (red) simulation data are plotted.
+The data of Figure \ref{fig:coordination-lifespans} show that most coordination configurations last a very brief time, with the majority forming for under 0.5 ps.
+The three most populous coordinations, ``S'', ``SO'', ``SOO'' (as determined from Figure \ref{fig:bonding-coordinations} by percentage) in both temperatures have coordination lifetimes of up to 1 ps, in some instances lasting up to 1.5 ps.
+The brevity of the lifespans overall speaks to the dynamic nature of the interface.
+The length of time in each coordination parallels the populations of the coordinations, and suggests an ordering of stabilities of the various bonding coordinations.
+The ``unbound'' configuration stands out as an anomaly amonst the lifespans of the other bonding coordinations.
+The few data points above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but also detaches far enough to avoid the quick rebinding and coordination change to ``S'' or ``O''.
+Those few long lived unbound data points indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement  has drawn it back to the surface to rebind.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/coordinations/coordination-lifespans.png}
+		\caption{}
+		\label{fig:coordination-lifespans}
+	\end{center}
+\end{figure}
+
+Returning now to the transition plots of Figure \ref{fig:coordination-transitions}, the behavior of the unbound transition to both ``S'' and ``O'' states can now be better characterized.
+Figure \ref{fig:coordination-lifespans} shows that the ``O'' state lifetimes are shorter than the ``S'' coordinated \suldiox.
+Figure \ref{fig:coordination-transitions} shows that the unbound-``O'' transition occurs almost three times more than the transition to the ``S''.
+Thus the unbound \suldiox~forms a bond to a neighboring water through its oxygen, but that bond is short-lived and quickly breaks sending the \suldiox~back to being unbound, or it transitions to the ``SO'' state by forming another bond through the sulfur.
+The unbound-``S'' transition does not occur as often because the ``S'' state is likely more stable than the ``O''.
+Once in the ``S'' coordination the \suldiox~does not quickly break the interaction from its sulfur to water, but rather remains for up to 1 ps in the ``S'' coordination before most likely forming an oxygen bond to make the ``SO'' coordination.
+We have outlined the likely behavior of \suldiox~as it transitions from the unbound state to binding with surface waters.
+
+Once the \suldiox~is bound to the water surface, the pathway leading back to the unbound state is not often traversed.
+As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the cluster of ``S-SO-SOO'' coordination nodes.
+This suggests that the \suldiox~sulfur interaction to a water oxygen has a much longer lifespan than the oxygen bonding to water hydrogens.
+The difference between the ``S-SO-SOO'' coordinations is an addition of oxygen bonds, and the frequent transitions show that the bonds to \suldiox~oxygens are quickly forming and breaking.
+For the \suldiox~sulfur interaction to break, the \suldiox~must enter a non-sulfur coordination (i.e.
+``O'', ``OO'', unbound, etc.) or a coordination with more than a single sulfur interaction (i.e.
 ``SS'', ``SSO'', etc.).
-The increased temperature causes more binding to the \suldiox~by allowing for multiple bonds through both the \suldiox~sulfur, and through the oxygens.
-Thus, the hotter temperature allows for increased coordination of the \suldiox~to the water surface.
-
-When the \suldiox~is on the cold water surface, the ``S'' and ``SO'' coordination populations are nearly equally represented.
-As the temperature increases to the hot system, the ``S'' coordination decreases, while the ``SO'' increases.
-We see from the RDFs that colder temperatures favor binding through the \suldiox~sulfur because of the larger first peak in the S$_{SO_2}$-O$_{H_2O}$ RDF.
-From this increase we expect the ``S''-coordinated \suldiox~to be more prevalent at cold temperatures, as is the case for the coordination distribution.
-However, the O$_{SO_2}$-H$_{H_2O}$ RDF indicates increased hydrogen bonding through the \suldiox~oxygens at higher temperatures manifested by a larger first peak in Figure \ref{fig:rdf} for the hot temperature.
-Thus, those \suldiox~bonding coordinations with at least a single bond to an oxygen and to the sulfur, (i.e.
-``SO'', ``SSO'', ``SSOO'', etc.) are more prevalent at the high temperature.
-Overall, we find evidence from both the bonding coordination distribution, and from the RDFs to suggest a modified bonding behavior at different temperatures.
-At an increased temperature the surface \suldiox~binds more through its oxygens than the cold counterpart.
-However, like the \suldiox~in the cold temperature, it likely has a persistent interaction through the sulfur atom to a neighboring water oxygen throughout most of the simulated trajectories.
+The transitions to coordinations that allow for breaking of the \suldiox~sulfur interactions, or switching the interaction to another water, are infrequent compared to those leading to an oxygen bond transition.
+Thus, the \suldiox~spends most of its time while bound to the surface waters breaking and forming hydrogen-bonds through its oxygens, and interacting with neighboring waters through a more persisten interaction via the \suldiox~sulfur atom.

### 2011-08-22 Background.tex
Finished with Fred's comments on the first half of the manuscript (bonding coordination and background stuff)

@@
 
-A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water hydrogens, or interacts via the sulfur atoms with water oxygens.
-To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters, we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters.
+A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water hydrogens, or interacts via the sulfur atom with water oxygens.
+To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters, those that lie in the topmost region of a gas/water interface, we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters.
 This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in other computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters.
-Thus a bonding coordination of ``OOH'' designates two proton-accepting bonding interactions through the water oxygen, and a single proton-donor bonding interaction through a hydrogen.
+Thus a bonding coordination of ``OOH'' designates two proton-acceptor bonding interactions through the water oxygen, and a single proton-donor bonding interaction through a hydrogen.
 More recently, Baer et al devised a nomenclature that explicitly enumerates the bonding to \suldiox~via the sulfur or oxygen atoms.\cite{Baer2010}
 
-In this work we port Buch et al's nomenclature to \suldiox~in order to show the amount of hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions form from the \suldiox-sulfur to water oxygens.
+In this work we adopt Buch et al's nomenclature to \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens.
 Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox~oxygens.
-Analysis of the distribution of \suldiox~coordinations will give insight to how it binds to the water surface.
+Analysis of the distribution of \suldiox~coordinations will give insight to how \suldiox~binds to the water surface.
 
-During the course of the analysis we use the definition of bonds of Baer et al.
-in determining the \suldiox~bonding coordinations.\cite{Baer2010} The definition is formed from a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox~oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.
+In our analysis we defined bonds using the distance criteria of Baer et al.
+in determining the \suldiox~bonding coordinations.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox~oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.
 
@@
 
-Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations congealing to a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of the surface hydrated \suldiox~by Baer et al.
-showed that the bonding coordination distributions of \suldiox~at the surface are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''.
+Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of surface hydrated \suldiox~by Baer et al.
+showed that the bonding coordination distributions of \suldiox~are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''.
 In the same work they then focused on the most dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.
 
-The cluster geometries found in Baer's study of the surface \suldiox~hydrates imply a cyclic bonding structure through the two or three waters involved.
-This is manifested as a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved.
-Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the \suldiox-sulfur and returning through the \suldiox-oxygen in a two-water cluster.
+The cluster geometries found in this study imply a cyclic bonding structure through the two or three waters involved.
+``Cyclic'' here is used to denote a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved.
+Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the sulfur and returning through the \suldiox-oxygen.
 
@@
 
-The optimized geometry of the \suldiox-hydrates suggest cyclic bonding structures, but it remains a different story entirely when a \suldiox~is placed in a dynamic environment such as in the course of molecular dynamics simulations of an aqueous surface.
-Although geometry optimization shows the formation of these cyclic complexes, do they form during a dynamic bonding process on a simulated water surface?
-Graph theoretical techniques can be employed to determine if cyclic structures form during the simulations.
+\subsection {Graph Theoretical Details}
+The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when a \suldiox~is placed in a dynamic environment such as in the course of molecular dynamics simulations of an aqueous surface.
+Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum.
+Do they also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior?
+To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data.
 Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998}
 
-Here we will only briefly introduce graph theoretical concepts as they have been described well by others, and with varied application to cyclic structures.\cite{Tutte1984, Balakrishnan2000, Harary1973, Huber2007, Garcia2004, Dury2001} A graph consists of a set of vertexes, and edges that connect the vertexes.
-A molecule can be represented with atoms as vertexes, and edges for each intramolecular covalent bond connecting the atoms.
+Here we briefly introduce graph theoretical concepts.
+They have been described well by others with varied application to cyclic structures.\cite{Tutte1984, Balakrishnan2000, Harary1973, Huber2007, Garcia2004, Dury2001} A graph consists of nodes, and edges that connect the nodes.
+A molecule can be represented with atoms as nodes, and edges for each intramolecular covalent bond connecting the atoms.
 The set of edges is then further expanded to include intermolecular interactions such as hydrogen bonds and other bonding bonding interactions.
 Edges may be assigned weights (i.e.
-bond lengths), types, and can be directional pointing towards a specific target from a source.
+bond lengths), types, and can be directional i.e.
+pointing towards a specific target from a source.
 A molecular system including all atoms, bonds, and interactions is thus fully described by a graph.
 
-To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queuing vertexes and all adjacent neighboring vertexes while performing a specified procedure on each visited vertex.
-This is easily performed on adjacency list or connectivity matrix data structures, iterating through atoms of interest in the graph as starting points of the search.
+To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queuing nodes and all neighboring nodes while performing a specified procedure on each visited node.
+This is easily performed on adjacency list or connectivity matrix data structures, iterating through nodes (i.e.
+atoms) of interest in the graph as starting points of the search.
 In BFS terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black).
-Using a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a vertex.
-A benefit of BFS on a graph is the ability to determine the smallest cyclic structure in which a given vertex is a member.
-In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox~sulfur as the root vertex will discover cyclic bonding structures in order of size.
+Using a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node.
+A benefit of BFS on a graph is the ability to determine the smallest cyclic structure in which a given node is a member.
+In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox~sulfur as the root node will discover cyclic bonding structures in order of size.
 Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox.
-Furthermore, it is possible to reconstruct a cycle's structure to find the size of the cycle (number of atoms involved), the number of unique waters in the cycle, and the type of cyclic structure encountered.
+Furthermore, it is possible to reconstruct a cycle's structure by finding its size (number of contributing atoms), and the number of unique waters in the cycle.
+This allows us to distinguish between various types of cyclic structures encountered.
 
-The various types of cyclic bonding arrangements are shown in Figure \ref{fig:cyclic-structures} for a \suldiox~molecule bonding to waters.
-Although three waters are shown in the molecular cartoon, any number of waters are allowed.
-Also, further bonding may occur to any of the participating molecules extraneous to the cyclic structure.
+Several arrangements of cyclic bonding structures are shown in Figure \ref{fig:cyclic-structures} for a \suldiox~molecule with three waters.
+Cycles with fewer or greater numbers of waters are also possible and encountered during MD.
 Cycle types \Rmnum{1}, \Rmnum{2}, \Rmnum{3} in Figure \ref{fig:cyclic-structures} are cyclic structures in which the \suldiox~is a member of the cycle.
 Types \Rmnum{4} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle formed near the \suldiox.
-Type \Rmnum{3} is of particular interest because it is one of the most dominant \suldiox~bonding coordinations (``SO'', as shown later) configured as a cyclic structure.
+Type \Rmnum{3} is of particular interest because the \suldiox~in the cycle has the most frequently occuring bonding coordination (``SO'', as shown later).
 
@@
 		\includegraphics[scale=1.0]{images/cycles/cycle-types-small.png}
-		\caption{}
+		\caption{\suldiox~in involved in various cyclic structures encountered during MD simulations with water.
+The cartoons show the five types of cyclic structures, numbered for reference.
+The cyclic structures may involve any number of waters, but here each structure is shown with three waters.
+Type \Rmnum{3} is formed by a \suldiox~with the ``SO'' bonding coordination, which is the most dominant bonding coordination encountered.}
 		\label{fig:cyclic-structures}
@@
 Baer et al.
-presented a detailed geometric and spectroscopic breakdown of type \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the arrangement of the waters involved.
+presented a detailed geometric and spectroscopic breakdown of type \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the atomic contributions of the waters to the cyclic structure.
 Type \Rmnum{3}-A (shown in Figure \ref{fig:type-3-varieties}A) is arranged with each water contributing an OH bond to the structure of the cycle.
 Type \Rmnum{3}B involves a single water contributing an OH, whereas the other two waters contribute only the oxygen atom or the entire water molecule to the structure, respectively.
-This nuance of the three-water type \Rmnum{3} structures, and the overall distribution are presented in more detail later.
+This nuance of the type \Rmnum{3} structures involving three waters, and the overall distribution are presented in more detail later.
 We also show the distribution of cyclic structures encountered during MD simulations to further understand the behaviors of \suldiox-hydrates at the water surface.
@@
 		\includegraphics[scale=1.0]{images/cycles/triple-cycle-types-small.png}
-		\caption{}
+		\caption{Type \Rmnum{3} cyclic structure (see Figure \ref{fig:cyclic-structures}) are found to occur in two varieties that are distinguished by the water atoms contributing to the cyclic structure.
+In ``type A'' the cycle is formed by an OH bond contributed by each water, and the SO bond of the \suldiox.
+``type B'' involves an oxygen atom from one water, an OH bond from a second water, and all atoms of the third water.}
 		\label{fig:type-3-varieties}

### 2011-08-22 ComputationalMethods.tex
Finished with Fred's comments on the first half of the manuscript (bonding coordination and background stuff)

@@
 Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1.0 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the cold and hot simulations, respectively.
-It was verified that these computational parameters yield a reasonable description of bulk room temperature water when simulating a neat-water system.
+These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
 
@@
 The initial 1 ns equilibration trajectory was not included in the final analysis.
-This resulted in 20,000 time steps of system trajectory for analysis in each replica of the system.
+This resulted in 20,000 time steps of system trajectory for analysis in the sets of hot and cold replicas of the system.

### 2011-08-22 Coordination-Results.tex
Finished with Fred's comments on the first half of the manuscript (bonding coordination and background stuff)

@@
 Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, by percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories.
-A first visual inspection reveals several trends in the bonding of the \suldiox.
+A first visual inspection reveals several trends.
 Clearly the ``SO'' coordination is the most populous at both temperatures.
@@
 In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally.
-The hotter temperature simulation shifts the distribution such that the ``S'' occurs 5\% less frequently than in the cold, and the ``SOO'' occurs nearly 10\% more times.
+The hotter temperature simulation shifts the distribution such that the ``S'' occurs 5\% less frequently than in the cold, and the ``SOO'' occurs nearly 10\% more often.
 The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``SOO'', respectively.
@@
 		\includegraphics[scale=1.0]{images/coordinations/so2-coordinations-percents-small.png}
-		\caption{}
+		\caption{The \suldiox~bonding coordinations occur with different frequencies at the surface under hot and cold temperature conditions.
+Shown above is the distribution of bonding coordinations of the cold (blue) and hot (red) \suldiox~over the entire set of trajectories.
+The values are the percentage of MD timesteps spent in the given bonding coordintion.}
 		\label{fig:bonding-coordinations}
@@
 
-	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from the distribution of coordinations.
-Clearly the cold \suldiox~spends more time than the hot \suldiox~(10\%) completely unbound from the surface waters.
-This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria set for the analysis.
+	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from this distribution of coordinations.
+Clearly the cold \suldiox~spends more time than the hot \suldiox, 13\% and 3\%, respectively, completely unbound from the surface waters.
+This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis.
 Furthermore, the most frequently occuring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
 Baer et al.
-also concluded that these three coordinations werethe most frequent for a room temperature simulation, and specifically isolated the ``SO'' and ``SOO'' as most common in their study.\cite{Baer2010}
+also concluded that these three coordinations were the most frequent for a room temperature simulation, and specifically identified the ``SO'' and ``SOO'' as most common in their study.\cite{Baer2010}
 
 	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (i.e.
-``O'', ``OO'', etc.) represent the least formed bonding coordinations.
-A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~oxygen-only bonding to waters.
-In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox-sulfur points further into the \wat~bulk than the oxygens.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, and to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
+``O'', ``OO'', etc.) represent the least frequently formed bonding coordinations.
+A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters.
+In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, and to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
 
@@
 ``SS'', ``SSO'', ``SSOO'', etc.).
-We see from the datat that single-sulfur coordination are overwhelmingly preferred over double-sulfur ones.
-Also, adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' coordinations are almost completely unrepresented (less than 1\%), and the change from ``SOO'' to ``SOOO'' produces a very large drop in the coordination distribution of Figure \ref{fig:bonding-coordinations}.
+We see from the data that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones.
+Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' are represented less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurences.
 
@@
 The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond assymetrically through one of the oxygens either once or twice to water hydrogens.
-The \suldiox~oxygen bonds will form and break throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.
+The \suldiox~oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.
 
@@
 In the cold temperature, the unbound, ``S'', and ``SO'' coordinations are more populated than in the hot systems.
-The increased temperature decreases the time spent unbound to the surface waters, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature coordination populations.
-We see that the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to the stateswith double-oxygen and double-sulfur coordinations.
+The increased temperature decreases the time spent in the unbound coordination, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature coordination populations.
+We see that the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to coordinations with double-oxygen and double-sulfur bonds.
 
@@
 This speaks to a dramatic difference in the surface behavior of \suldiox~at the two temperatures.
-The cold \suldiox~spends nearly equal time between the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''.
-Thus the addition, or removal of a bond through the \suldiox-oxygen is equally probable, as long as the pivotal sulfur interaction remains intact.
+The cold \suldiox~spends nearly equal time in the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''.
+Thus, the addition or removal of a bond through the \suldiox-oxygen to a neighboring coordination (i.e.
+addition of an oxygen bond from ``SO'' to ``SOO'', or removal of the bond from ``SO'' to ``S'') is equally probable, as long as the sulfur interaction with the water oxygen does not break.
+
 Figure \ref{fig:rdf} shows the radial distribution functions (RDF) of \suldiox~to water atoms for both cold (blue) and hot (red) temperatures.
-The \suldiox-S to \wat-O RDFs are nearly equal except for a slightly taller first peak of the cold system.
-Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF would indicate that with bonding mostly through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface waters.
-As the temperature increases in the hot system, the added bonding through a second oxygen interaction may be pulling the bonding waters further from the \suldiox-sulfur.
-This suggests that the cold \suldiox~bonds to water's surface more through its sulfur atom (as shown in the \suldiox-S RDF) and favors the more sulfur-centric bonding coordinations (i.e.
+The S-O$_{H_2O}$ RDFs are nearly equal except for a slightly taller first peak in the cold system.
+Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF would indicate that since bonding occurs frequently through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface waters.
+In the hot systems, the bonding coordinations with two oxygen bonds (i.e.
+``SOO'', etc.) occur more frequently than in the cold system.
+This additional bonding through a second oxygen interaction may slightly shift neighboring hydrating waters away from the \suldiox~sulfur towards the oxygen end of the molecule.
+This suggests that the cold \suldiox~bonds to water's surface closer through its sulfur atom (as shown in the S-O$_{H_2O}$ RDF) and favors the more ``sulfur-centric'' bonding coordinations (i.e.
 ``S'', ``SO'').
-The increased temperature allows the \suldiox~to bond more extensively through its oxygens to the ``SOO'' coordination.
-In turn, the increase binding through the oxygens may pull the \suldiox~further into the water interface and then allow for increased brief bonding through the sulfur, up to double-sulfur coordinations (i.e.
+The increased temperature of the hot system allows the \suldiox~to bond more extensively through its oxygens to the ``SOO'' coordination.
+The greater interactions through the oxygens, and higher bonding coordination, may pull the \suldiox~further into the water interface and then allow for increased bonding through the sulfur, up to double-sulfur coordinations (i.e.
 ``SS'', ``SSO'', etc.).
@@
 		\includegraphics[scale=1.0]{images/rdf/rdf-small.png}
-		\caption{}
+		\caption{Radial distribution functions (RDF) show the subtle change in how the hydrating waters around \suldiox~reposition as the temperature is increased from cold (blue) to hot (green).
+Shown are the two correlations between sulfur and water oxygens (left) and the \suldiox-oxygen and water hydrogens (right).}
 		\label{fig:rdf}
@@
 
-During the course of each simulation, the \suldiox~bonding coordination was determined and recorded for each timestep.
-From the coordination data we extract not only the populations of the various bonding coordinations encountered, but also the frequencies of transitions between the different coordinations (i.e.
-the number of times an \suldiox~switched from one coordination type to another).
+During the course of each cold and hot simulation, the \suldiox~bonding coordination was determined and recorded for each timestep.
+From the coordination data we extract not only the populations of the various bonding coordinations, but also the frequencies of transitions between the different coordinations (i.e.
+the number of times each \suldiox~switched from one coordination type to another).
 With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} A and B, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one coordination to another.
@@
 Additionally, the transition lines are numbered to the right of each line with the number of times each transition occured.
-Only the most populated nodes and transitions between them are shown.
-This accounts for the difference of inbound and outbound nodes to certain nodes not equalling zero, as would happen for a complete graph representation of all coordinations and transitions.
 
@@
 In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes.
-The number of transitions within this ``S-SO-SOO'' cluster of coordinations is an order of magnitude larger than any other transition.
+The number of transitions within this ``S-SO-SOO'' grouping of coordinations is an order of magnitude larger than any other transition.
 This indicates that while the \suldiox~is bound in any of the three most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other two coordinations in this group.
-The \suldiox~is rather active and constantly forming and breaking bonds to with its oxygens.
+Clearly, the \suldiox~is rather active and constantly forming and breaking bonds with its oxygens.
 
@@
 This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox~oxygen.
-As noted earlier, it is likely that the ``SOO'' \suldiox, assymetrically binding twice through a single oxygen, is pulled further into the water interface.
-It is likely more surrounded by waters, and in the hot system it can often form a brief third hydrogen bond to a water through the second \suldiox~oxygen.
-Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox returns back to the ``SOO'' coordination.
+As noted earlier, it is likely that ``SOO'' coordinated \suldiox, assymetrically binding twice through a single oxygen, is being pulled further into the water interface.
+It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox~oxygen.
+Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox~returns back to the ``SOO'' coordination.
 
 Given the information we have about the frequencies of bonding coordination transitions, it is possible to draw a likely route of adsorption beginning with an unbound \suldiox.
-From the unbound state, the \suldiox~can bind to waters either through the sulfur or an oxygen to enter the ``S'' or ``O'' coordinations, respectively.
+From the unbound coordination, the \suldiox~can bind to waters either through the sulfur or an oxygen to enter the ``S'' or ``O'' coordinations, respectively.
 At both temperatures we find that the coordination transition in Figure \ref{fig:coordination-transitions} from unbound to ``O'' occurs almost three times more than the transition from unbound to ``S''.
-Either the ``O'' state is more easily formed, but also quickly broken, accounting for the higher transition frequency, or the ``O'' coordination is in fact the first step in adsorption of the unbound \suldiox, leading to a transition to the most frequent coordination, ``SO'', through addition of an \suldiox~sulfur interaction to a water oxygen.
+We find two possibilities that may explain this difference.
+The ``O'' coordination may form more easily, but also break quickly after formation, accounting for the higher transition frequency.
+Otherwise, the ``O'' coordination may be the first step in adsorption of an unbound \suldiox, where a subsequent addition of an \suldiox~sulfur interaction to a water oxygen would lead to a transition to the most frequent coordination, ``SO''.
+In the latter case, any adsorption of \suldiox~may proceed through an oxygen binding, accounting for the increased unbound-``O'' transition frequency.
 
-To verify the claim that the ``O'' coordination forms from, and breaks quickly to the unbound state as is suggested by the transition frequency plots in Figure \ref{fig:coordination-transitions}, rather than enduring longer, we plot the lifespans of the various coordinations in Figure \ref{fig:coordination-lifespans}.
-Each point in the plot represent a time during the simulated trajectories in which the \suldiox~formed the respective coordination.
+To verify if the ``O'' coordination forms from, and breaks quickly to the unbound coordination as is suggested by the transition frequency plots in Figure \ref{fig:coordination-transitions}, we plot the lifespans of the various coordinations in Figure \ref{fig:coordination-lifespans}.
+Each point in the plot represents a time during the simulation in which the \suldiox~formed the respective coordination.
 The vertical ``lifespan'' position is calculated directly from the amount of time spent in the given coordination before changing to another.
@@
 The three most populous coordinations, ``S'', ``SO'', ``SOO'' (as determined from Figure \ref{fig:bonding-coordinations} by percentage) in both temperatures have coordination lifetimes of up to 1 ps, in some instances lasting up to 1.5 ps.
-The brevity of the lifespans overall speaks to the dynamic nature of the interface.
-The length of time in each coordination parallels the populations of the coordinations, and suggests an ordering of stabilities of the various bonding coordinations.
+The brevity of lifespans overall speaks to the dynamic nature of the interface.
+The length of time in each coordination parallels the populations of the coordinations, and suggests an ordering of steady states among various bonding coordinations.
 The ``unbound'' configuration stands out as an anomaly amonst the lifespans of the other bonding coordinations.
-The few data points above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but also detaches far enough to avoid the quick rebinding and coordination change to ``S'' or ``O''.
-Those few long lived unbound data points indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement  has drawn it back to the surface to rebind.
+The few lifespans above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''.
+Those few long lived unbound data points indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement occurs and it is drawn back to the surface to rebind.
 
@@
 
-Returning now to the transition plots of Figure \ref{fig:coordination-transitions}, the behavior of the unbound transition to both ``S'' and ``O'' states can now be better characterized.
-Figure \ref{fig:coordination-lifespans} shows that the ``O'' state lifetimes are shorter than the ``S'' coordinated \suldiox.
+Returning to the transition plots of Figure \ref{fig:coordination-transitions}, the behavior of the unbound transition to both ``S'' and ``O'' coordination can now be better characterized.
+Figure \ref{fig:coordination-lifespans} shows that the ``O'' coordinated lifetimes are shorter than the ``S'' coordinated ones.
 Figure \ref{fig:coordination-transitions} shows that the unbound-``O'' transition occurs almost three times more than the transition to the ``S''.
-Thus the unbound \suldiox~forms a bond to a neighboring water through its oxygen, but that bond is short-lived and quickly breaks sending the \suldiox~back to being unbound, or it transitions to the ``SO'' state by forming another bond through the sulfur.
-The unbound-``S'' transition does not occur as often because the ``S'' state is likely more stable than the ``O''.
-Once in the ``S'' coordination the \suldiox~does not quickly break the interaction from its sulfur to water, but rather remains for up to 1 ps in the ``S'' coordination before most likely forming an oxygen bond to make the ``SO'' coordination.
-We have outlined the likely behavior of \suldiox~as it transitions from the unbound state to binding with surface waters.
+The unbound \suldiox~forms a bond to a neighboring water through its oxygen, but that bond is short-lived and either quickly breaks (resulting in unbound \suldiox), or it transitions to the ``SO'' coordination by forming another bond through the sulfur.
+The unbound-``S'' transition does not occur as often.
+This may be because the ``S'' coordination is more stable than the ``O''.
+Once in the ``S'' coordination the \suldiox~does not quickly break the interaction from its sulfur to water, but rather remains for up to 1 ps in the ``S'' coordination before (most likely) forming an oxygen bond to make the ``SO'' coordination.
+We have outlined the likely behavior of \suldiox~as it transitions from the gas phase in an unbound coordination to binding to an aqueous interface by forming bonds and interactions with surface waters.
 
-Once the \suldiox~is bound to the water surface, the pathway leading back to the unbound state is not often traversed.
-As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the cluster of ``S-SO-SOO'' coordination nodes.
+Once the \suldiox~begins interacting with the water surface, the pathway leading back to the unbound coordination is not often traversed.
+As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the cluster of ``S-SO-SOO'' coordinations.
 This suggests that the \suldiox~sulfur interaction to a water oxygen has a much longer lifespan than the oxygen bonding to water hydrogens.
-The difference between the ``S-SO-SOO'' coordinations is an addition of oxygen bonds, and the frequent transitions show that the bonds to \suldiox~oxygens are quickly forming and breaking.
+The difference between the ``S-SO-SOO'' coordinations is an addition or removal of oxygen bonds, and the frequent transitions between them show that the bonds to \suldiox~oxygens are quickly forming and breaking.
 For the \suldiox~sulfur interaction to break, the \suldiox~must enter a non-sulfur coordination (i.e.

### 2011-08-23 Cycle-Results.tex
Adding in result write up for the cyclic structures

\subsection{Cyclic \suldiox~Hydrate Structures}

Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule. The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier and depicted in Figure \ref{fig:cyclic-structures}. Only the most commonly occuring subset of the cyclic structures were analyzed.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/cycles/SO-cycle-breakdown-with-cartoons-small.png}
		\caption{}
		\label{fig:cyclic-breakdown}
	\end{center}
\end{figure}
The plot in Figure \ref{fig:cycle-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures. Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4). The two tallest data points, left-most in the plot, show the overall time spent in all types of cyclic structures. Clearly the hot system \suldiox~spends more time in a cyclic structure than at the cold temperature. The plot is based on two criteria for determining if a cyclic structure has formed. (1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations. (2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen. As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on order of discovery. The \suldiox~will be inevitably involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS. The larger and more extended cyclic structures involving more waters affect the behavior of the hydrogen-bonding network of the water surface. However, we focus here only on the smallest cycles involving the \suldiox~as the most affect the \suldiox~bonding and hydrating behaviors.

It is remarkable that the difference of the time spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\%hot). The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such. Thus, in addition to having found the most likely bonding coordination during the simulated life of \suldiox, we have also found that the hydrates of the \suldiox~form a cyclic structure for much of the time it is bound to the water surface.

Now we look at the different types of cyclic hydrates, distinguished by the number of waters involved in the bonding structure. In Figure \ref{fig:cyclic-breakdown} the single and double water cycles are the least frequently encountered structures, accounting for less than 10\% of both temperature simulations. Formation of the single type is likely energetically unfavorable because of the proximity of the single water to the \suldiox. The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In the static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form. However, the MD simulations here have introduced many waters into a dynamic environment allowing for extended bonding networks, and the results show clearly that the double-water cyclic hydrate is formed much less often (less than 5\% at both temperatures) than the triple-water form.

The results for triple and quadruple-water structures show that larger hydrate cycles are favored at higher temperatures. Although the higher number of hydrating waters (>4) are not shown, those contribute minimally to the overall distribution. The majority of the cyclic hydrates are formed with three waters in the triple type. This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules. The two triple cycle structure are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of the contributions to the overall distribution. It is notable that each temperature has a different dominant type of triple-water cyclic structure. The cold system forms more of the type-B, and the hotter system forms primarily triple type-A.

\subsection {Cyclic Structure Energy Calculations}

\subsection {Cyclic Hydrate Structure Lifetimes}

Knowing that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and also often taking part in some type of cyclic structure, how long does the cyclic hydrate form before breaking to an acyclic structure? To answer the questions of cyclic lifespans, a method was devised to define a lifetime of a cycle. For each trajectory, the coordinate data was analyzed to determine if a cyclic hydrate structure was formed based on the bonding distance criteria established earlier in this manuscript. A timeline was then produced where each timestep was given a value of 1 or 0 if a \suldiox~cycle was found or not found, respectively. This resulted in a time-function, $C(t)$, similar in nature to a digital signal, as shown in black in Figure \ref{fig:debouncing}. Because

### 2011-08-24 Cycle-Results.tex
new cycle debouncing image... almost done with new additions for the cycle results stuff, sans energy text.

@@
 A timeline was then produced where each timestep was given a value of 1 or 0 if a \suldiox~cycle was found or not found, respectively.
-This resulted in a time-function, $C(t)$, similar in nature to a digital signal, as shown in black in Figure \ref{fig:debouncing}.
-Because
+This resulted in a time-function, $C(t)$, similar in nature to a digital signal.
+The resulting cycle formation time-function is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
+At the far left of the plot the signal is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed.
+The cyclic structure is very dynamic constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly.
+This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the signal lasting less than 10 fs each.
+Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to ``debounce'' the time-signal to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
+The time-signal, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width.
+The resulting smoothed signal, $C_s(t)$, was then cutoff with the following criteria:
+
+\[
+  f(t) = \left\{ 
+  \begin{array}{l l}
+    0 & \quad C_s(t) < 0.2\\
+    1 & \quad C_s(t) \geq 0.2\\
+  \end{array} \right.
+\]
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/cycles/cycle-debouncing-small.png}
+		\caption{}
+		\label{fig:debouncing}
+	\end{center}
+\end{figure}

### 2011-08-24 Cycle-Results.tex
Almost done adding content to the cycle results. Only waiting to write up the energy results.

@@
 The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier and depicted in Figure \ref{fig:cyclic-structures}.
-Only the most commonly occuring subset of the cyclic structures were analyzed.
+Only the most commonly occurring subset of the cyclic structures were analyzed.
 
@@
 The resulting cycle formation time-function is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
-At the far left of the plot the signal is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed.
+At the far left of the plot the function is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed.
 The cyclic structure is very dynamic constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly.
-This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the signal lasting less than 10 fs each.
-Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to ``debounce'' the time-signal to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
-The time-signal, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width.
-The resulting smoothed signal, $C_s(t)$, was then cutoff with the following criteria:
+This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the function lasting less than 10 fs each.
+Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to ``debounce'' the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
+The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width.
+The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:
 
@@
 
+where $f(t)$ is the debounced time-function that better represents the lifespans of cycles, eliminating the very brief discontinuities and excursions outside the bonding criteria that cause the rapid ``bouncing'' of the cyclic time-function, $C(t)$.
+Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final ``debounced'' function, $f(t)$ (green).
+
 \begin{figure}[h!]
@@
 \end{figure}
+
+The distribution of cycle lifetimes (i.e.
+contiguous spans of time spent with $f(t) = 1$) was determined.
+A plot of the cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations.
+The most frequent lifespan for both temperature data sets lasts between 0-1 ps.
+This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time.
+Even with the ``debouncing'' procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps.
+The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend also extending to the longer lifespans as well.
+The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps.
+All of the distribution shows only a $<1.5$\% difference between cold and hot cyclic lifetimes.
+Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time.
+
+Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles.
+Above 4 ps, most of the cycles that form are found in the hot system.
+The 8 ps cycles are notable in that they form  for just under half of an entire single simulated trajectory.
+
+We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently.
+The \suldiox~is likely forming and breaking bonds with waters extraneous to the cyclic hydrate structures (i.e.
+not directly involved in the bonds of the cycle).
+For the longer-lived cycles, the extraneous bonding to the \suldiox~may have little effect on the cyclic hydrate waters.
+However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e.
+``S'', ``SS'', ``O'', etc), the cyclic structure is necessarily broken.
+Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure.
+Figure \ref{fig:bonding-coordinations} shows the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively.
+Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold).
+The increased temperature appears to cause the \suldiox~to bond in such a way that is more conducive to the formation of cyclic structures.
+Furthermore, the distribution of bonding coordinations in the hot system shifts the transition between bonding coordinations away from cycle-breaking coordinations (i.e.
+those coordinations that necessarily disallow cycle formation).
+This possibly accounts for the slightly increased populations of longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/cycles/cyclic-lifespans-inset-small.png}
+		\caption{}
+		\label{fig:cycle-lifespans}
+	\end{center}
+\end{figure}

### 2011-08-24 Cycle-Results.tex
Finished the cycle results sans energies. Now on to introduction, 				 conclusion, and energies!

@@
 \end{figure}
-The plot in Figure \ref{fig:cycle-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures.
+The plot in Figure \ref{fig:cyclic-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures.
 Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4).

### 2011-08-24 Background.tex
Working out edits. Added all figure captions.

@@
 		\includegraphics[scale=1.0]{images/cycles/double-cycle-type2-small.png}
-		\caption{}
+		\caption{During the course of MD, the covalent and intermolecular bonds between a \suldiox~and the hydrating waters may form into a ring, resulting in a cyclic hydrate structure.
+Depicted here is one example of such a cyclic structure, formed by the covalent SO and OH bonds of the two molecule types, and the intermolecular hydrogen bonds, and S-O$_{H_2O}$ interactions.
+The atoms involved are the S (yellow), O (red), and H (white).}
 		\label{fig:cyclic-example}

### 2011-08-24 Coordination-Results.tex
Working out edits. Added all figure captions.

@@
 		\includegraphics[scale=1.0]{images/coordinations/coordination-transitions.png}
-		\caption{}
+		\caption{The bonding coordinations of \suldiox~on a water surface rapidly change as bonds break and form throughout the MD trajectories.
+Depicted here is a graph of the bonding coordinations, shown as colored nodes of varying size.
+Larger and more darkly colored coordination nodes are those more often encountered during MD (see Figure \ref{fig:bonding-coordinations}).
+The directed edges between the nodes represent the number of times the coordination transition occured.
+Thicker lines correspond to a greater number of transitions between coordinations.
+The number of time a given transition occured is labeled to the right of the corresponding line.}
 		\label{fig:coordination-transitions}
@@
 		\includegraphics[scale=1.0]{images/coordinations/coordination-lifespans.png}
-		\caption{}
+		\caption{Given that certain bonding coordinations of the surface \suldiox~are more often encountered than others (see Figure \ref{fig:bonding-coordinations}), the plot here shows the time spent in each bonding coordination.
+Each point in the plot represents a single span of simulation time that the \suldiox~had the given coordination.
+The amount of time spent corresponds to the vertical position along the lifespan axis, in ps.}
 		\label{fig:coordination-lifespans}

### 2011-08-24 Cycle-Results.tex
Working out edits. Added all figure captions.

--- 2011-08-24 Cycle-Results.tex
@@
 		\includegraphics[scale=1.0]{images/cycles/SO-cycle-breakdown-with-cartoons-small.png}
-		\caption{}
+		\caption{Cyclic hydrate structures form throughout the MD simulations involving the \suldiox~and one or more water molecules.
+Shown here are the different cycles by number of water (up to 4), and their occurence rate as a percentage of the total simulation time.
+The far right part of the plot shows the contributions of the two types of triple-water cycles (see Figure \ref{fig:type-3-varieties}).
+The cycles had to meet the ``minimally-SO'' bonding criteria (at least one bond to the sulfur and one oxygen bond on the \suldiox) to be used in the analysis.}
 		\label{fig:cyclic-breakdown}
@@
 \end{figure}
-The plot in Figure \ref{fig:cycle-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures.
+
+The plot in Figure \ref{fig:cyclic-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures.
 Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4).
@@
 
-It is remarkable that the difference of the time spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\%hot).
+It is remarkable that the difference of the time spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\% hot).
 The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such.
@@
 		\includegraphics[scale=1.0]{images/cycles/cycle-debouncing-small.png}
-		\caption{}
+		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures.
+A ``debouncing'' procedure was developed (described above in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
+A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure.
+A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the ``debounced'' cycle formation time function, $f(t)$ (green).}
 		\label{fig:debouncing}
@@
 		\includegraphics[scale=1.0]{images/cycles/cyclic-lifespans-inset-small.png}
-		\caption{}
+		\caption{The lifetime of a cyclic hydrate structure after it forms, and before it distorts in shape due to the movement of the liquid surface molecules, is most often short ($<$ 1 ps).
+Shown here is a plot of cycle hydrate structure lifetimes after having applied the ``debouncing'' procedure (described in the text).
+Both cold (blue) and hot (red) simulations are graphed.
+The inset expands the view of the region above 1 ps.}
 		\label{fig:cycle-lifespans}

### 2011-08-25 Cycle-Results.tex
Added reference to yet unpublished, but submitted, work of mine - Shamay2011. Also finished all the beta-draft edits and ready to hand over to Fred for some critical analysis.

@@
 Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule.
-The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier and depicted in Figure \ref{fig:cyclic-structures}.
-Only the most commonly occurring subset of the cyclic structures were analyzed.
+The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier in the text and depicted in Figure \ref{fig:cyclic-structures}.
+Only the most commonly occurring subset of the cyclic structures were analyzed based on two selection criteria for determining formation of a cyclic structure: (1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations.
+(2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen.
+As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on order of discovery.
+The \suldiox~will be involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS.
+The larger and more extended cyclic structures involving more waters affect the behavior of the hydrogen-bonding network of the water surface.
+However, we focus here only on the smallest cycles involving the \suldiox~as they most affect the \suldiox~bonding and hydration.
 
@@
 Clearly the hot system \suldiox~spends more time in a cyclic structure than at the cold temperature.
-The plot is based on two criteria for determining if a cyclic structure has formed.
-(1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations.
-(2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen.
-As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on order of discovery.
-The \suldiox~will be inevitably involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS.
-The larger and more extended cyclic structures involving more waters affect the behavior of the hydrogen-bonding network of the water surface.
-However, we focus here only on the smallest cycles involving the \suldiox~as the most affect the \suldiox~bonding and hydrating behaviors.
-
 It is remarkable that the difference of the time spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\% hot).
 The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such.
-Thus, in addition to having found the most likely bonding coordination during the simulated life of \suldiox, we have also found that the hydrates of the \suldiox~form a cyclic structure for much of the time it is bound to the water surface.
+Thus, in addition to having earlier found the most likely bonding coordination during the simulated life of \suldiox, we have also found that the hydrates of the \suldiox~form cyclic structures for much of the time while bound to the water surface.
 
 Now we look at the different types of cyclic hydrates, distinguished by the number of waters involved in the bonding structure.
-In Figure \ref{fig:cyclic-breakdown} the single and double water cycles are the least frequently encountered structures, accounting for less than 10\% of both temperature simulations.
-Formation of the single type is likely energetically unfavorable because of the proximity of the single water to the \suldiox.
+In Figure \ref{fig:cyclic-breakdown} the single and double water cycles are the least frequently encountered structures, accounting for less than 10\% of both cold and hot temperature simulation times.
+Formation of the single-water type is likely energetically unfavorable because of the proximity of the water to the \suldiox~required to form the bonding cycle.
 The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In the static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form.
@@
 The results for triple and quadruple-water structures show that larger hydrate cycles are favored at higher temperatures.
-Although the higher number of hydrating waters (>4) are not shown, those contribute minimally to the overall distribution.
-The majority of the cyclic hydrates are formed with three waters in the triple type.
-This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules.
-The two triple cycle structure are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of the contributions to the overall distribution.
+Although the higher number of hydrating waters ($>$ 4) are not shown, those contribute minimally to the overall distribution.
+The majority of the cyclic hydrates are formed with three waters in the triple-water type.
+This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}).
+The triple-water type cycle results were broken down into contributions from the type-A and type-B triple-water structures.
+The two triple cycle structures are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of their contributions to the overall distribution.
 It is notable that each temperature has a different dominant type of triple-water cyclic structure.
@@
 
-Knowing that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and also often taking part in some type of cyclic structure, how long does the cyclic hydrate form before breaking to an acyclic structure?
+Knowing that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and also often taking part in some type of cyclic structure, how long does the cyclic hydrate form before breaking to an acyclic hydrate structure?
 To answer the questions of cyclic lifespans, a method was devised to define a lifetime of a cycle.
 For each trajectory, the coordinate data was analyzed to determine if a cyclic hydrate structure was formed based on the bonding distance criteria established earlier in this manuscript.
-A timeline was then produced where each timestep was given a value of 1 or 0 if a \suldiox~cycle was found or not found, respectively.
+A timeline was then produced where each timestep was given a value of 1 or 0 if a \suldiox~cycle was formed or not, respectively.
 This resulted in a time-function, $C(t)$, similar in nature to a digital signal.
-The resulting cycle formation time-function is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
-At the far left of the plot the function is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed.
-The cyclic structure is very dynamic constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly.
-This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the function lasting less than 10 fs each.
-Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to ``debounce'' the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
+A representative portion of the resulting cycle formation time-function is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
+At the far left of the plot the function is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed later in time.
+The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly.
+This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the $C(t)$ function lasting less than 10 fs each.
+Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to debounce the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
 The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width.
@@
 where $f(t)$ is the debounced time-function that better represents the lifespans of cycles, eliminating the very brief discontinuities and excursions outside the bonding criteria that cause the rapid ``bouncing'' of the cyclic time-function, $C(t)$.
-Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final ``debounced'' function, $f(t)$ (green).
+Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).
 
@@
 		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures.
-A ``debouncing'' procedure was developed (described above in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
+A debouncing procedure was developed (described above in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
 A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure.
-A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the ``debounced'' cycle formation time function, $f(t)$ (green).}
+A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
 		\label{fig:debouncing}
@@
 contiguous spans of time spent with $f(t) = 1$) was determined.
-A plot of the cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations.
+The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations.
 The most frequent lifespan for both temperature data sets lasts between 0-1 ps.
 This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time.
-Even with the ``debouncing'' procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps.
+Even with the debouncing procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps.
 The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend also extending to the longer lifespans as well.
@@
 Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time.
-
 Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles.
@@
 Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure.
-Figure \ref{fig:bonding-coordinations} shows the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively.
+Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively.
 Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold).
+Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox), compared to the cold system coordination distribution.
 The increased temperature appears to cause the \suldiox~to bond in such a way that is more conducive to the formation of cyclic structures.
-Furthermore, the distribution of bonding coordinations in the hot system shifts the transition between bonding coordinations away from cycle-breaking coordinations (i.e.
-those coordinations that necessarily disallow cycle formation).
-This possibly accounts for the slightly increased populations of longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
+This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
 
@@
 		\caption{The lifetime of a cyclic hydrate structure after it forms, and before it distorts in shape due to the movement of the liquid surface molecules, is most often short ($<$ 1 ps).
-Shown here is a plot of cycle hydrate structure lifetimes after having applied the ``debouncing'' procedure (described in the text).
+Shown here is a plot of cycle hydrate structure lifetimes after having applied the debouncing procedure (described in the text).
 Both cold (blue) and hot (red) simulations are graphed.

### 2011-08-26 Coordination-Results.tex
Wrote an introduction, added paragraph about "why we need to talk about the cycle-signal bouncing and why it needs to be taken care of"

@@
 A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters.
-In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, and to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
+In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, and to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
 

### 2011-08-26 Cycle-Results.tex
Wrote an introduction, added paragraph about "why we need to talk about the cycle-signal bouncing and why it needs to be taken care of"

@@
 		\caption{Cyclic hydrate structures form throughout the MD simulations involving the \suldiox~and one or more water molecules.
-Shown here are the different cycles by number of water (up to 4), and their occurence rate as a percentage of the total simulation time.
+Shown here are the different cycles by number of water (up to 4), and their occurrence rate as a percentage of the total simulation time.
 The far right part of the plot shows the contributions of the two types of triple-water cycles (see Figure \ref{fig:type-3-varieties}).
@@
 For each trajectory, the coordinate data was analyzed to determine if a cyclic hydrate structure was formed based on the bonding distance criteria established earlier in this manuscript.
-A timeline was then produced where each timestep was given a value of 1 or 0 if a \suldiox~cycle was formed or not, respectively.
+A timeline was then produced where each timestep was given a value of 1 or 0 depending on if a \suldiox~hydrate bonding cycle was present or not, respectively.
 This resulted in a time-function, $C(t)$, similar in nature to a digital signal.
-A representative portion of the resulting cycle formation time-function is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
-At the far left of the plot the function is in the ``no cycle'' state, and then switches ``on'' as a cycle is formed later in time.
+
+In an electro-mechanical system, a mechanical switch triggering an electrical signal often outputs a noisy signal that has a period of signal bounce before settling to a final value.
+This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing.
+Many mechanical, electrical, and software solutions have been devised for this problem of signal bounce.
+Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before breaking, where the $C(t)$ function bounces before settling into a final value.
+The bounce in the function, manifested as a series of rapid switches between cycle presence and absence, signals a transition period due to the distortions of the bonds in the cyclic hydrates.
+We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions.
+The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states.
+$f(t)$ may then be used to calculate statistics about the cycle lifespans.
+The statistics thus calculated will no longer be skewed by the multiple short changes of state associated with the function bounce.
+
+A representative portion of the resulting cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
+At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time.
 The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly.
-This is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the $C(t)$ function lasting less than 10 fs each.
-Because of the very rapid breaking and reformation of the cyclic structure, a procedure was used to debounce the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
+This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the $C(t)$ function lasting less than 10 fs each.
+The following procedure was used to debounce the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
 The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width.
@@
 
-where $f(t)$ is the debounced time-function that better represents the lifespans of cycles, eliminating the very brief discontinuities and excursions outside the bonding criteria that cause the rapid ``bouncing'' of the cyclic time-function, $C(t)$.
+where $f(t)$ is the debounced time-function that better represents the lifespans of cycles, eliminating the very brief discontinuities and excursions outside the bonding criteria that cause the rapid bouncing of the cyclic time-function, $C(t)$.
 Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).

### 2011-08-26 Background.tex
starting some more edits

@@
 
-A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water hydrogens, or interacts via the sulfur atom with water oxygens.
+A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water hydrogens, or interacts via the sulfur atom with water-oxygens.
 To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters, those that lie in the topmost region of a gas/water interface, we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters.
 This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in other computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters.
-Thus a bonding coordination of ``OOH'' designates two proton-acceptor bonding interactions through the water oxygen, and a single proton-donor bonding interaction through a hydrogen.
-More recently, Baer et al devised a nomenclature that explicitly enumerates the bonding to \suldiox~via the sulfur or oxygen atoms.\cite{Baer2010}
+Thus a bonding coordination of ``OOH'' designates two proton-acceptor bonding interactions through the water-oxygen, and a single proton-donor bonding interaction through a hydrogen.
+More recently, Baer et al.
+devised a nomenclature that explicitly enumerates the bonding to \suldiox~via the sulfur or oxygen atoms.\cite{Baer2010}
 
 In this work we adopt Buch et al's nomenclature to \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens.
-Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox~oxygens.
+Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water-oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox-oxygens.
 Analysis of the distribution of \suldiox~coordinations will give insight to how \suldiox~binds to the water surface.
@@
 In our analysis we defined bonds using the distance criteria of Baer et al.
-in determining the \suldiox~bonding coordinations.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox~oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.
+in determining the \suldiox~bonding coordinations.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox-oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.
 
@@
 Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum.
-Do they also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior?
+Do the cyclic structures also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior?
 To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data.
@@
 bond lengths), types, and can be directional i.e.
-pointing towards a specific target from a source.
+pointing towards a target node from a source node.
 A molecular system including all atoms, bonds, and interactions is thus fully described by a graph.
@@
 Using a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node.
-A benefit of BFS on a graph is the ability to determine the smallest cyclic structure in which a given node is a member.
-In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox~sulfur as the root node will discover cyclic bonding structures in order of size.
+A benefit of BFS on a graph is the ability to determine the smallest cyclic structure containing a given node.
+In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox~sulfur as the starting, or root node or the search, will discover cyclic bonding structures in order of size.
 Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox.
@@
 Types \Rmnum{4} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle formed near the \suldiox.
-Type \Rmnum{3} is of particular interest because the \suldiox~in the cycle has the most frequently occuring bonding coordination (``SO'', as shown later).
+Type \Rmnum{3} is of particular interest because the \suldiox~in this cycle has the most frequently occuring bonding coordination (``SO'', as shown later).
 
@@
 		\includegraphics[scale=1.0]{images/cycles/cycle-types-small.png}
-		\caption{\suldiox~in involved in various cyclic structures encountered during MD simulations with water.
+		\caption{\suldiox~in various cyclic structures encountered during MD simulations with water.
 The cartoons show the five types of cyclic structures, numbered for reference.
@@
 Baer et al.
-presented a detailed geometric and spectroscopic breakdown of type \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the atomic contributions of the waters to the cyclic structure.
+presented a detailed geometric and spectroscopic breakdown of type \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the set of water atoms involved in the cyclic structure.
 Type \Rmnum{3}-A (shown in Figure \ref{fig:type-3-varieties}A) is arranged with each water contributing an OH bond to the structure of the cycle.
@@
 		\includegraphics[scale=1.0]{images/cycles/triple-cycle-types-small.png}
-		\caption{Type \Rmnum{3} cyclic structure (see Figure \ref{fig:cyclic-structures}) are found to occur in two varieties that are distinguished by the water atoms contributing to the cyclic structure.
+		\caption{Type \Rmnum{3} cyclic structures (see Figure \ref{fig:cyclic-structures}) are found to occur in two varieties that are distinguished by the water atoms contributing to the cyclic structure.
 In ``type A'' the cycle is formed by an OH bond contributed by each water, and the SO bond of the \suldiox.

### 2011-08-26 ComputationalMethods.tex
starting some more edits

@@
 On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).~\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions.
-The electronic charge density was desccribed using an auxilliary basis set of plane waves.
-Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1.0 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the cold and hot simulations, respectively.
+The electronic charge density was described using an auxiliary basis set of plane waves.
+Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the cold and hot simulations, respectively.
 These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
@@
 Initially, 10 equilibrated boxes of side-lengths 10.0\angs~, with 36 randomly packed water molecules were used.
-5 of the boxes were used for each of the cold and hot simulations.
+Five of the boxes were used for each of the cold and hot simulations.
 A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis.
@@
 Periodic boundaries were set on the two short axes to form an infinite slab.
-The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1.0 fs.
+The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1 fs.
 The initial 1 ns equilibration trajectory was not included in the final analysis.

### 2011-08-26 Coordination-Results.tex
starting some more edits

@@
 Clearly the ``SO'' coordination is the most populous at both temperatures.
-The second and third most populated coordinations are the ``S'' and ``SOO'', however the distributions differ between temperatures.
+The second and third most populated coordinations are the ``S'' and ``SOO'', however their distributions differ between temperatures.
 In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally.
@@
 	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from this distribution of coordinations.
-Clearly the cold \suldiox~spends more time than the hot \suldiox, 13\% and 3\%, respectively, completely unbound from the surface waters.
+Clearly the cold \suldiox~spends more time than the hot \suldiox, 13\% versus 3\%, respectively, completely unbound from the surface waters.
 This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis.
@@
 	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (i.e.
-``O'', ``OO'', etc.) represent the least frequently formed bonding coordinations.
+``O'', ``OO'', etc.) represent the least frequently formed.
 A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters.

### 2011-08-26 Introduction.tex

--- 2011-06-15 Introduction.tex
@@
 
-The chemistry of our atmosphere is vastly complex and poorly understood.
-Reactions taking place on particle surfaces and liquid boundaries dominate many atmospheric chemical pathways, and guide terrestrial and industrial byproducts to their ultimate forms and destinations.
-How do atmospheric conditions affect the chemistry of environmental constituents?
+One of chemistry's many remaining ``black box'' events is the adsorption of a gas into a water surface.
+Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions.
+How does gas bind to a water surface, and what steps are involved in adsorption?
+How does a gas molecule near a water surface affect the water to which it will bind?
+What is the structure of hydrating waters in the surface region, and how does that hydrated solute molecule behave differently than as a gas?
+Experiments to address these questions provide valuable information, but can never fully describe these microscopic events and behaviors.
+The information can be determined computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.
+
+As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems.
+The molecular structure, high solubility in water, and its abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
+A complete picture of the \suldiox~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in water interfacial regions.
+
+In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structure that forms around a surface-bound \suldiox~on water.
+We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the fluidity of the \suldiox~hydrate structures, and the behaviors on the water surface.
+Unlike small cluster studies that use DFT calculations on geometry optimized structures to find likely geometries in a static vacuum setting, The quantum MD technique allows us to simulate the process much more realistically, and with greater accuracy than our previous classical MD work could allow.\cite{Shamay2011} This computational study greatly enhances the picture developed in our previous work on \suldiox.
+In our previous computational study we used classical MD to determine net orientational behavior of \suldiox~binding to a water surface, and of the orientation of the waters in response to the presence of an adsorbing gas.
+Understanding the net behavior of the molecules in the aqueous interfacial region throughout adsorption was the first step to understanding the specific details of gas binding and surface behavior.
+
+Quantum MD techniques accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to look at the specific bonding interactions that form within the surface hydrates, and in the extended bonding further into the water.
+The surfaces in this work were simulated at two temperatures: room temperature at 298K, and the more atmospherically relevant cold 270K.
+This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Conclusions from the experiments regarding the specific nature of those complexes could only be inferred.
+
+We believe this to be the first temperature study using quantum MD to look at the binding of a small gas molecule to a water surface.
+We show how temperature affects the bonding behavior of the surface adsorbed \suldiox~to neighboring waters.
+A proposed set of steps in a binding mechanism for \suldiox~adsorbing to a water surface is presented.
+We also look at \suldiox~behavior when already bound by the surface waters.
+Lastly, our analysis shows some interesting extended bonding behavior of \suldiox~hydrates, and that the hydrating waters and the \suldiox~often form cyclic ring structures through intermolecular bonds.

### 2011-08-26 Coordination-Results.tex
New round of edits given out to others

@@
 Shown above is the distribution of bonding coordinations of the cold (blue) and hot (red) \suldiox~over the entire set of trajectories.
-The values are the percentage of MD timesteps spent in the given bonding coordintion.}
+The values are the percentage of MD timesteps spent in the given bonding coordination.}
 		\label{fig:bonding-coordinations}
@@
 This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis.
-Furthermore, the most frequently occuring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
+Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
 Baer et al.
@@
 Three \suldiox~oxygen bonds only form if both \suldiox~oxygens are interacting with water hydrogens.
-Our finding that triple-oxygen coordinations rarely form complements Baer et al's conclusion about the assymetry in the oxygen interactions.
+Our finding that triple-oxygen coordinations rarely form complements Baer et al's conclusion about the asymmetry in the oxygen interactions.
 
@@
 We see from the data that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones.
-Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' are represented less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurences.
+Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' represent less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
 
 We can now form a picture of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study.
-The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond assymetrically through one of the oxygens either once or twice to water hydrogens.
+The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens.
 The \suldiox~oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.
@@
 In the cold temperature, the unbound, ``S'', and ``SO'' coordinations are more populated than in the hot systems.
-The increased temperature decreases the time spent in the unbound coordination, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature coordination populations.
+The increased temperature decreases the time spent in the unbound coordination, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature populations.
 We see that the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to coordinations with double-oxygen and double-sulfur bonds.
@@
 The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently).
-Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequently occuring transitions.
-Additionally, the transition lines are numbered to the right of each line with the number of times each transition occured.
+Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequently occurring transitions.
+Additionally, the transition lines are numbered to the right of each line with the number of times each transition occurred.
 
@@
 Larger and more darkly colored coordination nodes are those more often encountered during MD (see Figure \ref{fig:bonding-coordinations}).
-The directed edges between the nodes represent the number of times the coordination transition occured.
+The directed edges between the nodes represent the number of times the coordination transition occurred.
 Thicker lines correspond to a greater number of transitions between coordinations.
-The number of time a given transition occured is labeled to the right of the corresponding line.}
+The number of time a given transition occurred is labeled to the right of the corresponding line.}
 		\label{fig:coordination-transitions}
@@
 In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes.
-The number of transitions within this ``S-SO-SOO'' grouping of coordinations is an order of magnitude larger than any other transition.
+The number of transitions within this ``S-SO-SOO'' group of coordinations is an order of magnitude larger than any other transition.
 This indicates that while the \suldiox~is bound in any of the three most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other two coordinations in this group.
@@
 This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox~oxygen.
-As noted earlier, it is likely that ``SOO'' coordinated \suldiox, assymetrically binding twice through a single oxygen, is being pulled further into the water interface.
+As noted earlier, it is likely that ``SOO'' coordinated \suldiox, asymmetrically binding twice through a single oxygen, is being pulled further into the water interface.
 It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox~oxygen.
@@
 The length of time in each coordination parallels the populations of the coordinations, and suggests an ordering of steady states among various bonding coordinations.
-The ``unbound'' configuration stands out as an anomaly amonst the lifespans of the other bonding coordinations.
+The ``unbound'' configuration stands out as an anomaly amongst the lifespans of the other bonding coordinations.
 The few lifespans above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''.
@@
 Once the \suldiox~begins interacting with the water surface, the pathway leading back to the unbound coordination is not often traversed.
-As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the cluster of ``S-SO-SOO'' coordinations.
+As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the ``S-SO-SOO'' group of coordinations.
 This suggests that the \suldiox~sulfur interaction to a water oxygen has a much longer lifespan than the oxygen bonding to water hydrogens.
@@
 The transitions to coordinations that allow for breaking of the \suldiox~sulfur interactions, or switching the interaction to another water, are infrequent compared to those leading to an oxygen bond transition.
-Thus, the \suldiox~spends most of its time while bound to the surface waters breaking and forming hydrogen-bonds through its oxygens, and interacting with neighboring waters through a more persisten interaction via the \suldiox~sulfur atom.
+Thus, the \suldiox~spends most of its time while bound to the surface waters breaking and forming hydrogen-bonds through its oxygens, and interacting with neighboring waters through a more persistent interaction via the \suldiox~sulfur atom.

### 2011-08-26 Cycle-Results.tex
New round of edits given out to others

@@
 (2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen.
-As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on order of discovery.
+As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on the search's order of cycle discovery.
 The \suldiox~will be involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS.
@@
 Clearly the hot system \suldiox~spends more time in a cyclic structure than at the cold temperature.
-It is remarkable that the difference of the time spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\% hot).
+It is remarkable that the difference of the times spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\% hot).
 The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such.
@@
 It is notable that each temperature has a different dominant type of triple-water cyclic structure.
-The cold system forms more of the type-B, and the hotter system forms primarily triple type-A.
+The cold system forms more of the type-B, and the hotter system forms primarily type-A.
 
@@
 Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold).
-Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox), compared to the cold system coordination distribution.
-The increased temperature appears to cause the \suldiox~to bond in such a way that is more conducive to the formation of cyclic structures.
+Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox).
+This is relative to the cold system coordination distribution that has equal ``S'' and ``SOO'' populations.
+The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures.
 This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.

### 2011-08-31 Introduction.tex
old?

@@
 
-As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems.
-The molecular structure, high solubility in water, and its abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
+As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems.\cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} The molecular structure, high solubility in water, and its abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
 A complete picture of the \suldiox~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in water interfacial regions.

### 2011-08-31 Cycle-Results.tex
About to start final edits to the pre-draft sans energy stuff

@@
 
-Knowing that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and also often taking part in some type of cyclic structure, how long does the cyclic hydrate form before breaking to an acyclic hydrate structure?
-To answer the questions of cyclic lifespans, a method was devised to define a lifetime of a cycle.
-For each trajectory, the coordinate data was analyzed to determine if a cyclic hydrate structure was formed based on the bonding distance criteria established earlier in this manuscript.
-A timeline was then produced where each timestep was given a value of 1 or 0 depending on if a \suldiox~hydrate bonding cycle was present or not, respectively.
-This resulted in a time-function, $C(t)$, similar in nature to a digital signal.
+We know that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and is also often taking part in some type of cyclic structure.
+With this in mind, it becomes interesting to ask: how long does a cyclic hydrate form before breaking to an acyclic hydrate structure?
+To answer the question of cyclic lifespan, a method was devised to define the lifetime of a cycle.
+For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript.
+A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a particular \suldiox~hydrate bonding cycle was present or not, respectively.
+This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.
 
-In an electro-mechanical system, a mechanical switch triggering an electrical signal often outputs a noisy signal that has a period of signal bounce before settling to a final value.
+In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value.
 This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing.
-Many mechanical, electrical, and software solutions have been devised for this problem of signal bounce.
-Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before breaking, where the $C(t)$ function bounces before settling into a final value.
-The bounce in the function, manifested as a series of rapid switches between cycle presence and absence, signals a transition period due to the distortions of the bonds in the cyclic hydrates.
+Many mechanical, electrical, and software solutions have been devised to suppress the transients or ``debounce'' the signal.
+Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before dissolution, where the $C(t)$ function bounces before settling into a final value.
+The bounce in the function, manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence.
+Physically, the bond lengths within the cycle (or the cycle-to-be) are fluctuating back and forth across our bond-length criteria.
+It is thus an artifact of our algorithmic determination of the presence or absence of a bond.
+Removing this artifact will allow us to examine the longer-time bonding behavior and this is accomplished as follows.
+
 We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions.
@@
 $f(t)$ may then be used to calculate statistics about the cycle lifespans.
-The statistics thus calculated will no longer be skewed by the multiple short changes of state associated with the function bounce.
-
-A representative portion of the resulting cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
+A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
 At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time.
 The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly.
-This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of very sharp spikes in the $C(t)$ function lasting less than 10 fs each.
-The following procedure was used to debounce the time-function to disregard very brief moments when the cycle breaks or forms for less than 20 fs.
-The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width.
+This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each.
+The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration.
 The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:
@@
 
-where $f(t)$ is the debounced time-function that better represents the lifespans of cycles, eliminating the very brief discontinuities and excursions outside the bonding criteria that cause the rapid bouncing of the cyclic time-function, $C(t)$.
+where $f(t)$ is the debounced time-function that represents the lifespans of cycles.
 Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).
@@
 We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently.
-The \suldiox~is likely forming and breaking bonds with waters extraneous to the cyclic hydrate structures (i.e.
+The \suldiox~is likely forming and breaking bonds with waters external to the cyclic hydrate structures (i.e.
 not directly involved in the bonds of the cycle).
-For the longer-lived cycles, the extraneous bonding to the \suldiox~may have little effect on the cyclic hydrate waters.
+For the longer-lived cycles, the external bonding to the \suldiox~may have little effect on the cyclic hydrate waters.
 However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e.

### 2011-08-31 Introduction.tex
About to start final edits to the pre-draft sans energy stuff

--- 2011-07-26 Introduction.tex
@@
 
-%The chemistry of our atmosphere is vastly complex and poorly understood.
-Reactions taking place on particle surfaces and liquid boundaries dominate many atmospheric chemical pathways, and guide terrestrial and industrial byproducts to their ultimate forms and destinations.
-How do atmospheric conditions affect the chemistry of environmental constituents?
+One of Chemistry's many remaining ``black box'' events is the adsorption of gas molecules onto a water surface.
+Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions.
+How does gas initially bind to a water surface, and what steps are involved in the subsequent adsorption?
+How does an unbound gas molecule near a water surface affect the water to which it will bind?
+What is the structure of hydrating waters in the surface region, and how does a hydrated solute molecule behave differently than as a gas?
+Experiments to address these questions provide valuable information, but can never fully characterize these microscopic events and behaviors.
+However, these systems can be fully characterized computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.
+
+As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems.
+Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
+A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in interfacial water regions.
+
+In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water.
+We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behaviors of the water surface.
+The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.
+It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated.
+In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas.
+Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas binding and surface behavior.
+
+%Unlike small cluster studies that use DFT calculations on geometry optimized structures to find likely geometries in a static vacuum setting, The quantum MD technique allows us to simulate the process much more realistically, and with greater accuracy than our previous classical MD work could allow.\cite{Shamay2011} This computational study greatly enhances the picture developed in our previous work on \suldiox.
+
+Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.
+Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 270K.
+This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Conclusions regarding the specific nature of those complexes were inferred from the experiments.
+With our current computational studies we now have insights about the specific microscopic geometries and behaviors of the hydrating complexes.
+
+We believe this to be the first temperature study using quantum MD to study the binding of small gas molecules on a water surface.
+We show how temperature affects the bonding behavior of the surface adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface.
+We also examine \suldiox~behavior when bound to the surface waters.
+Lastly, our analysis demonstrates an extended bonding of \suldiox~hydrate structures, as \suldiox~hydrates are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.

### 2011-08-31 Background.tex

@@
 
-A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water hydrogens, or interacts via the sulfur atom with water-oxygens.
-To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters, those that lie in the topmost region of a gas/water interface, we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters.
-This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in other computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters.
+A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water-hydrogens, or interacts via the sulfur atom with water-oxygens.
+To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters (those that lie in the topmost region of a gas/water interface), we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters.
+This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in more recent computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters.
 Thus a bonding coordination of ``OOH'' designates two proton-acceptor bonding interactions through the water-oxygen, and a single proton-donor bonding interaction through a hydrogen.
@@
 \subsection {Graph Theoretical Details}
-The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when a \suldiox~is placed in a dynamic environment such as in the course of molecular dynamics simulations of an aqueous surface.
+The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of molecular dynamics simulations of an aqueous surface.
 Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum.
@@
 A molecule can be represented with atoms as nodes, and edges for each intramolecular covalent bond connecting the atoms.
-The set of edges is then further expanded to include intermolecular interactions such as hydrogen bonds and other bonding bonding interactions.
+The set of edges is then further expanded to include intermolecular interactions such as hydrogen bonds and other bonding interactions.
 Edges may be assigned weights (i.e.
-bond lengths), types, and can be directional i.e.
+bond lengths), types, and can be directional, i.e.
 pointing towards a target node from a source node.
@@
 A benefit of BFS on a graph is the ability to determine the smallest cyclic structure containing a given node.
-In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox~sulfur as the starting, or root node or the search, will discover cyclic bonding structures in order of size.
+In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox-sulfur as the starting, or root node or the search, will discover cyclic bonding structures in order of size.
 Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox.
@@
 Cycle types \Rmnum{1}, \Rmnum{2}, \Rmnum{3} in Figure \ref{fig:cyclic-structures} are cyclic structures in which the \suldiox~is a member of the cycle.
-Types \Rmnum{4} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle formed near the \suldiox.
+Types \Rmnum{4} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle types formed near the \suldiox.
 Type \Rmnum{3} is of particular interest because the \suldiox~in this cycle has the most frequently occuring bonding coordination (``SO'', as shown later).
@@
 Type \Rmnum{3}-A (shown in Figure \ref{fig:type-3-varieties}A) is arranged with each water contributing an OH bond to the structure of the cycle.
-Type \Rmnum{3}B involves a single water contributing an OH, whereas the other two waters contribute only the oxygen atom or the entire water molecule to the structure, respectively.
+Type \Rmnum{3}-B involves a single water contributing an OH, whereas the other two waters contribute only the oxygen atom or the entire water molecule to the structure, respectively.
 This nuance of the type \Rmnum{3} structures involving three waters, and the overall distribution are presented in more detail later.

### 2011-08-31 ComputationalMethods.tex

@@
 The electronic charge density was described using an auxiliary basis set of plane waves.
-Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the cold and hot simulations, respectively.
+Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively.
 These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
 
-Initially, 10 equilibrated boxes of side-lengths 10.0\angs~, with 36 randomly packed water molecules were used.
+Initially, 10 equilibrated boxes of side-lengths 10.0\angs, with 36 randomly packed water molecules were used.
 Five of the boxes were used for each of the cold and hot simulations.
@@
 A copy of the initial system cubes were then expanded along one axis (z-axis) to 25\angs.
-The system energy was minimized.
+The system energy was minimized through a geometry optimization.
 Subsequently, the system was equilibrated for 1 ns in canonical ensemble (NVT) conditions.
@@
 The initial 1 ns equilibration trajectory was not included in the final analysis.
-This resulted in 20,000 time steps of system trajectory for analysis in the sets of hot and cold replicas of the system.
+This resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system.

### 2011-08-31 Coordination-Results.tex

@@
 The bonding coordination of the surface \suldiox~was determined at each timestep of the simulations.
-Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, by percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories.
+Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, as a percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories.
 A first visual inspection reveals several trends.

### 2011-08-31 Introduction.tex

@@
 As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems.
-Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
+\cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
 A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in interfacial water regions.
@@
 We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behaviors of the water surface.
-The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.
-It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated.
+The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated.
 In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas.
@@
 
-Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.
-Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 270K.
-This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Conclusions regarding the specific nature of those complexes were inferred from the experiments.
-With our current computational studies we now have insights about the specific microscopic geometries and behaviors of the hydrating complexes.
+Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.\cite{Baer2010} Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 273K.
+This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Although conclusions regarding the specific nature of those complexes could only be inferred from the experiments, our current computational studies now provide us with insights about the specific microscopic geometries and behaviors of the hydrating complexes.
 
@@
 We show how temperature affects the bonding behavior of the surface adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface.
-We also examine \suldiox~behavior when bound to the surface waters.
-Lastly, our analysis demonstrates an extended bonding of \suldiox~hydrate structures, as \suldiox~hydrates are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.
+We also examine \suldiox~binding behavior when bound to the surface waters.
+Lastly, our analysis demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.

### 2011-09-08 Background.tex
More of Fred's final edits

@@
 
-In this work we adopt Buch et al's nomenclature to \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens.
+In this work we adopt a nomenclature scheme for \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens.
 Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water-oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox-oxygens.
@@
 
-In our analysis we defined bonds using the distance criteria of Baer et al.
-in determining the \suldiox~bonding coordinations.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox-oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.
+To determine \suldiox~bonding coordinations, we defined intermolecular bonds using the distance criteria of Baer et al.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox-oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.
 
@@
 showed that the bonding coordination distributions of \suldiox~are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''.
-In the same work they then focused on the most dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.
+In the same work they then focused on the dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.
+However, none of the studies have yet focused directly on the presence of an extended hydrating structure involving closed rings of molecular interactions involving the solvated \suldiox~molecules.
 
@@
 ``Cyclic'' here is used to denote a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved.
-Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the sulfur and returning through the \suldiox-oxygen.
+Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the sulfur and returning through a \suldiox-oxygen.
 
@@
 \subsection {Graph Theoretical Details}
-The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of molecular dynamics simulations of an aqueous surface.
+The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface.
 Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum.
@@
 A benefit of BFS on a graph is the ability to determine the smallest cyclic structure containing a given node.
-In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox-sulfur as the starting, or root node or the search, will discover cyclic bonding structures in order of size.
+In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox-sulfur as the starting, or root node for the search, will discover cyclic bonding structures in order of size.
 Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox.

### 2011-09-08 ComputationalMethods.tex
More of Fred's final edits

@@
 
-On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).~\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions.
+On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions.
 The electronic charge density was described using an auxiliary basis set of plane waves.
-Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP)~\cite{LEE1988}, and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.~\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively.
+Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively.
 These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
@@
 The initial 1 ns equilibration trajectory was not included in the final analysis.
-This resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system.
+This simulation process resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system, for a total of 100,000 timesteps at each temperature.

### 2011-09-08 Coordination-Results.tex
More of Fred's final edits

@@
 
-The bonding coordination of the surface \suldiox~was determined at each timestep of the simulations.
+The bonding coordination of each \suldiox~was determined at each timestep of the simulations.
 Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, as a percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories.
@@
 
-	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (i.e.
+	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (e.g.
 ``O'', ``OO'', etc.) represent the least frequently formed.
 A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters.
-In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, and to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
+In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
 
@@
 Baer et al performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens.
-They concluded that there is asymmetric hydrogen bonding through the \suldiox~oxygens, with one binding more often than the other.
-This is supported by the findings here where all the double oxygen coordinations (i.e.
-``OO'', ``SOO'', etc.) represent a much lower percentage of the coordinations than the single oxygen counterparts (i.e.
+They concluded that there is asymmetric hydrogen bonding through the \suldiox-oxygens, with one oxygen binding more often than the other.
+This is supported by the findings here where all the double oxygen coordinations (e.g.
+``OO'', ``SOO'', etc.) represent a much lower percentage of the coordinations than the single oxygen counterparts (e.g.
 ``O'', ``SO'', etc.).
-Furthermore, a triple-oxygen coordination (i.e.
+Furthermore, a triple-oxygen coordination (e.g.
 ``OOO'', ``SOOO'', etc.) is very rarely encountered.
-Three \suldiox~oxygen bonds only form if both \suldiox~oxygens are interacting with water hydrogens.
+Three \suldiox-oxygen bonds only form if both \suldiox-oxygens are interacting with water hydrogens.
 Our finding that triple-oxygen coordinations rarely form complements Baer et al's conclusion about the asymmetry in the oxygen interactions.
 
-Having established the preference for an interaction through the \suldiox~sulfur atom, we look to the right side of Figure \ref{fig:bonding-coordinations} at the double-sulfur coordinations (i.e.
+Having established the preference for an interaction through the \suldiox-sulfur atom, we look to the right side of Figure \ref{fig:bonding-coordinations} at the double-sulfur coordinations (e.g.
 ``SS'', ``SSO'', ``SSOO'', etc.).
 We see from the data that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones.
-Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' represent less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
+Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' together represent less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
 
@@
 The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens.
-The \suldiox~oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.
+The \suldiox-oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.
 
@@
 The cold \suldiox~spends nearly equal time in the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''.
-Thus, the addition or removal of a bond through the \suldiox-oxygen to a neighboring coordination (i.e.
+Thus, the addition or removal of a bond through the \suldiox-oxygen to a neighboring coordination (e.g.
 addition of an oxygen bond from ``SO'' to ``SOO'', or removal of the bond from ``SO'' to ``S'') is equally probable, as long as the sulfur interaction with the water oxygen does not break.
@@
 The S-O$_{H_2O}$ RDFs are nearly equal except for a slightly taller first peak in the cold system.
-Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF would indicate that since bonding occurs frequently through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface waters.
-In the hot systems, the bonding coordinations with two oxygen bonds (i.e.
+Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF indicates that since bonding occurs frequently through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface waters.
+In the hot systems, the bonding coordinations with two oxygen bonds (e.g.
 ``SOO'', etc.) occur more frequently than in the cold system.
-This additional bonding through a second oxygen interaction may slightly shift neighboring hydrating waters away from the \suldiox~sulfur towards the oxygen end of the molecule.
-This suggests that the cold \suldiox~bonds to water's surface closer through its sulfur atom (as shown in the S-O$_{H_2O}$ RDF) and favors the more ``sulfur-centric'' bonding coordinations (i.e.
+This additional bonding through a second oxygen interaction may slightly shift neighboring hydrating waters away from the \suldiox-sulfur towards the oxygen end of the molecule.
+This suggests that the cold \suldiox~bonds closer to the surface waters through its sulfur atom (as shown in the S-O$_{H_2O}$ RDF) and favors the more ``sulfur-centric'' bonding coordinations (e.g.
 ``S'', ``SO'').
 The increased temperature of the hot system allows the \suldiox~to bond more extensively through its oxygens to the ``SOO'' coordination.
-The greater interactions through the oxygens, and higher bonding coordination, may pull the \suldiox~further into the water interface and then allow for increased bonding through the sulfur, up to double-sulfur coordinations (i.e.
+The greater interactions through the oxygens, and higher bonding coordination, may pull the \suldiox~further into the water interface and then allow for increased bonding through the sulfur, up to double-sulfur coordinations (e.g.
 ``SS'', ``SSO'', etc.).
@@
 
-During the course of each cold and hot simulation, the \suldiox~bonding coordination was determined and recorded for each timestep.
+During the course of each simulation, the \suldiox~bonding coordination was determined and recorded for each timestep.
 From the coordination data we extract not only the populations of the various bonding coordinations, but also the frequencies of transitions between the different coordinations (i.e.
@@
 The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently).
-Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequently occurring transitions.
+Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequent transitions.
 Additionally, the transition lines are numbered to the right of each line with the number of times each transition occurred.
@@
 Thicker lines correspond to a greater number of transitions between coordinations.
-The number of time a given transition occurred is labeled to the right of the corresponding line.}
+The number of times a given transition occurred is labeled to the right of the corresponding line.}
 		\label{fig:coordination-transitions}
@@
 As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} than transitions to less populated coordinations.
-However, some insights to the bonding process are made clearer from these graphs.
+Insights to the bonding process are made clearer from these graphs.
 In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes.
@@
 One very surprising result is in the transition from ``SOO-SOOO''.
-The transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations.
-This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox~oxygen.
+This transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations.
+This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox-oxygen.
 As noted earlier, it is likely that ``SOO'' coordinated \suldiox, asymmetrically binding twice through a single oxygen, is being pulled further into the water interface.
-It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox~oxygen.
-Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox~returns back to the ``SOO'' coordination.
+It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox-oxygen.
+Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox~returns to the ``SOO'' coordination.
 
@@
 The ``O'' coordination may form more easily, but also break quickly after formation, accounting for the higher transition frequency.
-Otherwise, the ``O'' coordination may be the first step in adsorption of an unbound \suldiox, where a subsequent addition of an \suldiox~sulfur interaction to a water oxygen would lead to a transition to the most frequent coordination, ``SO''.
+Otherwise, the ``O'' coordination may be the first step in adsorption of an unbound \suldiox, where a subsequent addition of an \suldiox-sulfur interaction to a water oxygen would lead to a transition to the most frequent coordination, ``SO''.
 In the latter case, any adsorption of \suldiox~may proceed through an oxygen binding, accounting for the increased unbound-``O'' transition frequency.
@@
 The vertical ``lifespan'' position is calculated directly from the amount of time spent in the given coordination before changing to another.
-Both cold (blue) and hot (red) simulation data are plotted.
+Both cold (blue) and hot (red) data are plotted.
 The data of Figure \ref{fig:coordination-lifespans} show that most coordination configurations last a very brief time, with the majority forming for under 0.5 ps.
@@
 The ``unbound'' configuration stands out as an anomaly amongst the lifespans of the other bonding coordinations.
-The few lifespans above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''.
+The few lifespans above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that it also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''.
 Those few long lived unbound data points indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement occurs and it is drawn back to the surface to rebind.
@@
 Each point in the plot represents a single span of simulation time that the \suldiox~had the given coordination.
-The amount of time spent corresponds to the vertical position along the lifespan axis, in ps.}
+The total amount of consecutive timesteps in a coordination corresponds to the vertical position along the lifespan axis, in ps.}
 		\label{fig:coordination-lifespans}
@@
 As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the ``S-SO-SOO'' group of coordinations.
-This suggests that the \suldiox~sulfur interaction to a water oxygen has a much longer lifespan than the oxygen bonding to water hydrogens.
-The difference between the ``S-SO-SOO'' coordinations is an addition or removal of oxygen bonds, and the frequent transitions between them show that the bonds to \suldiox~oxygens are quickly forming and breaking.
-For the \suldiox~sulfur interaction to break, the \suldiox~must enter a non-sulfur coordination (i.e.
-``O'', ``OO'', unbound, etc.) or a coordination with more than a single sulfur interaction (i.e.
+This suggests that the \suldiox-sulfur interaction to a water oxygen has a much longer lifespan than the oxygen bonding to water hydrogens.
+The difference between the ``S-SO-SOO'' coordinations is an addition or removal of oxygen bonds, and the frequent transitions between them show that the bonds to \suldiox-oxygens are quickly forming and breaking.
+For the \suldiox-sulfur interaction to break, the \suldiox~must enter a non-sulfur coordination (e.g.
+``O'', ``OO'', unbound, etc.) or a coordination with more than a single sulfur interaction (e.g.
 ``SS'', ``SSO'', etc.).
-The transitions to coordinations that allow for breaking of the \suldiox~sulfur interactions, or switching the interaction to another water, are infrequent compared to those leading to an oxygen bond transition.
-Thus, the \suldiox~spends most of its time while bound to the surface waters breaking and forming hydrogen-bonds through its oxygens, and interacting with neighboring waters through a more persistent interaction via the \suldiox~sulfur atom.
+The transitions to coordinations that allow for breaking of the \suldiox-sulfur interactions, or switching the interaction to another water, are infrequent compared to those leading to an oxygen bond transition.
+Thus, the \suldiox~spends most of its time while bound to the surface waters breaking and forming hydrogen-bonds through its oxygens, and interacting with neighboring waters through a more persistent interaction via the \suldiox-sulfur atom.

### 2011-09-08 Cycle-Results.tex
More of Fred's final edits

@@
 Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule.
-The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form during MD, as posited earlier in the text and depicted in Figure \ref{fig:cyclic-structures}.
-Only the most commonly occurring subset of the cyclic structures were analyzed based on two selection criteria for determining formation of a cyclic structure: (1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations.
+The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form, as posited earlier in the text and depicted in Figure \ref{fig:cyclic-structures}.
+Only the most commonly occurring subset of the cyclic structures were analyzed based on two selection criteria: (1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations.
 (2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen.
@@
 The larger and more extended cyclic structures involving more waters affect the behavior of the hydrogen-bonding network of the water surface.
-However, we focus here only on the smallest cycles involving the \suldiox~as they most affect the \suldiox~bonding and hydration.
+We focus here only on the smallest cycles involving the \suldiox~as they most affect the \suldiox~bonding and hydration.
 
@@
 The two tallest data points, left-most in the plot, show the overall time spent in all types of cyclic structures.
-Clearly the hot system \suldiox~spends more time in a cyclic structure than at the cold temperature.
-It is remarkable that the difference of the times spent in a cyclic structure between the two temperatures is 15\% (42\% cold, 57\% hot).
+Clearly the hot \suldiox~spends more time in a cyclic structure than at the cold temperature.
+It is remarkable that the time spent in a cyclic structure shows a 15\% difference (42\% cold, 57\% hot).
 The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such.
@@
 Formation of the single-water type is likely energetically unfavorable because of the proximity of the water to the \suldiox~required to form the bonding cycle.
-The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In the static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form.
+The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In those static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form.
 However, the MD simulations here have introduced many waters into a dynamic environment allowing for extended bonding networks, and the results show clearly that the double-water cyclic hydrate is formed much less often (less than 5\% at both temperatures) than the triple-water form.
@@
 The majority of the cyclic hydrates are formed with three waters in the triple-water type.
-This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}).
-The triple-water type cycle results were broken down into contributions from the type-A and type-B triple-water structures.
+This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that, of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}).
+The triple-water type cycle results were broken-down into contributions from the type-A and type-B triple-water structures.
 The two triple cycle structures are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of their contributions to the overall distribution.
@@
 This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each.
-The time-function, $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration.
+$C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration.
 The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:
@@
 		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures.
-A debouncing procedure was developed (described above in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
+A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
 A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure.
@@
 Even with the debouncing procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps.
-The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend also extending to the longer lifespans as well.
+The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend extending to the longer lifespans as well.
 The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps.
@@
 Above 4 ps, most of the cycles that form are found in the hot system.
-The 8 ps cycles are notable in that they form  for just under half of an entire single simulated trajectory.
+The 8 ps cycles are notable in that they form for just under half the length of one of the simulated trajectories.
 

### 2011-09-08 Introduction.tex
More of Fred's final edits

@@
 What is the structure of hydrating waters in the surface region, and how does a hydrated solute molecule behave differently than as a gas?
-Experiments to address these questions provide valuable information, but can never fully characterize these microscopic events and behaviors.
+Experiments to address these questions provide valuable information, but can never fully characterize microscopic events and behaviors.
 However, these systems can be fully characterized computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.
@@
 \cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
-A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in interfacial water regions.
+A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in water's surface region.
 
 In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water.
-We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behaviors of the water surface.
+We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behavior of the water surface.
 The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated.
 In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas.
-Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas binding and surface behavior.
+Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas-binding and surface behavior.
 
@@
 We believe this to be the first temperature study using quantum MD to study the binding of small gas molecules on a water surface.
-We show how temperature affects the bonding behavior of the surface adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface.
+We show how temperature affects the bonding behavior of the surface-adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface.
 We also examine \suldiox~binding behavior when bound to the surface waters.

### 2011-11-30 Background.tex
Conclusions! Finishing up the final edits before handing over to Geri.

@@
 
-Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of surface hydrated \suldiox~by Baer et al.
+Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008, Jayne1990, Jayne1990a, Donaldson1995} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of surface hydrated \suldiox~by Baer et al.
 showed that the bonding coordination distributions of \suldiox~are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''.
 In the same work they then focused on the dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.
-However, none of the studies have yet focused directly on the presence of an extended hydrating structure involving closed rings of molecular interactions involving the solvated \suldiox~molecules.
+However, none of the studies have yet focused directly on the presence of an extended hydrating structure forming closed rings of molecular interactions involving the solvated \suldiox~molecules.
 
@@
 		\caption{During the course of MD, the covalent and intermolecular bonds between a \suldiox~and the hydrating waters may form into a ring, resulting in a cyclic hydrate structure.
-Depicted here is one example of such a cyclic structure, formed by the covalent SO and OH bonds of the two molecule types, and the intermolecular hydrogen bonds, and S-O$_{H_2O}$ interactions.
+Depicted here is one example of such a cyclic structure, formed by the covalent SO and OH bonds of the two molecule types, the intermolecular hydrogen bonds, and S-O$_{H_2O}$ interactions.
 The atoms involved are the S (yellow), O (red), and H (white).}
@@
 \subsection {Graph Theoretical Details}
-The optimized geometry of the \suldiox-hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface.
+The optimized geometry of the \suldiox~hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface.
 Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum.
@@
 bond lengths), types, and can be directional, i.e.
-pointing towards a target node from a source node.
+pointing from a source node towards a target node.
 A molecular system including all atoms, bonds, and interactions is thus fully described by a graph.
 
-To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A BFS is a recursive algorithm of queuing nodes and all neighboring nodes while performing a specified procedure on each visited node.
+To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A graph search is a recursive algorithm of queuing nodes and all neighboring nodes while performing a specified procedure on each visited node.
 This is easily performed on adjacency list or connectivity matrix data structures, iterating through nodes (i.e.
 atoms) of interest in the graph as starting points of the search.
-In BFS terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black).
-Using a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node.
-A benefit of BFS on a graph is the ability to determine the smallest cyclic structure containing a given node.
+In graph search terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black).
+Performing a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node.
+A benefit of BFS is the ability to determine the smallest cyclic structure containing a given node.
 In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox-sulfur as the starting, or root node for the search, will discover cyclic bonding structures in order of size.
@@
 Type \Rmnum{3}-B involves a single water contributing an OH, whereas the other two waters contribute only the oxygen atom or the entire water molecule to the structure, respectively.
-This nuance of the type \Rmnum{3} structures involving three waters, and the overall distribution are presented in more detail later.
+This nuance of the type \Rmnum{3} structures involving three waters, and the overall distribution of structures are presented in more detail later.
 We also show the distribution of cyclic structures encountered during MD simulations to further understand the behaviors of \suldiox-hydrates at the water surface.
@@
 In ``type A'' the cycle is formed by an OH bond contributed by each water, and the SO bond of the \suldiox.
-``type B'' involves an oxygen atom from one water, an OH bond from a second water, and all atoms of the third water.}
+``Type B'' involves an oxygen atom from one water, an OH bond from a second water, and all atoms of the third water.}
 		\label{fig:type-3-varieties}

### 2011-11-30 Conclusions.tex
Conclusions! Finishing up the final edits before handing over to Geri.

\section {Conclusions}

The adsorption of small gas molecules to water surfaces has been extensively studied over the past few decades. Much has been learned about the energies of hydrate configurations, and the kinetics of gaseous uptake into aqueous systems. Yet, the specific nature of the adsorption process, including the various geometries, hydrate species, and bonding pathways remains largely unknown. As a gas transitions into the liquid water phase, it passes through a fluid interfacial region that remains poorly understood. Our understanding of the processes and chemistry of the interface is still in its infancy, but we gain unique new insights that are key to understanding many environmentally important processes at aqueous surfaces.

Presented herein are the results of ab initio molecular dynamics simulations that focus on how a wandering gaseous \suldiox~molecule first makes contact with a water surface, and subsequently forms extended hydrate structures with interfacial water molecules. The computational studies complement and expand on experimental studies from this laboratory that found surface complexation of \suldiox~at a water surface.\cite{Tarbuck2005,Tarbuck2006,Ota2011} Furthermore, these computations build upon and enrich our understanding of adsorbing \suldiox~behavior from our recently published computational study on interfacial geometries of aqueous surface \suldiox~molecules.\cite{Shamay2011}

Our simulations show that \suldiox~has a preferred means of bonding and interacting with surface water molecules by taking on various bonding coordinations. In this work it was shown that the ``SO'' bonding configuration is the most preferred, with ``S'' and ``SOO'' also contributing greatly to the coordination distribution. Once a \suldiox~has bound to form a surface hydrate it rarely forms multiple bonding interactions through the sulfur atom, and even less frequently takes on a configuration with no sulfur interactions to nearby waters.

This study is one of very few temperature studies looking at the microscopic nature of interfacial gas molecules on water. By changing the temperature, it was found that a hotter water system leads to longer \suldiox~binding to the water surface. The distribution of bonding coordinations was greatly affected by a temperature change, shifting populations of bonding configurations because of the altered \suldiox~and \wat~behavior. At the higher temperature, \suldiox~forms more frequent bonds to interfacial waters through the sulfur and oxygen atoms. Overall, we have determined that the \suldiox~hydrate interactions are transient, binding and unbinding to water molecules rapidly in very dynamic bonding coordinations.

In this work we introduce the use of a graph structure to represent atoms and interconnectedness between molecules, and also to represent transitions between the various bonding coordinations of an adsorbed \suldiox. It was shown that the intermolecular bonds formed through the \suldiox-oxygens are quickly broken and formed, lasting briefly compared to sulfur interactions. From the graph of bonding coordination transitions, we found a likely pathway for \suldiox adsorption starting with an unbound gas-phase \suldiox, ending with a hydrated \suldiox~species bound to surface waters.

The formation of cyclic hydrate structures was probed and it was found that these hydrated bonding ring species form during much of a simulated trajectory. Temperature increases the occurrence of cyclic structures, and also shifts the distribution of the specific types of cycles being formed. Two types of cyclic tri-hydrates were discovered during the course of simulations. The cycle lifetimes were found to be mostly short-lived, with a majority lasting less than 1 ps before breaking and reforming due to the dynamic bonding and motion of the surface waters and \suldiox~molecules. Temperature did not have a very dramatic effect on the cyclic lifespans, but higher temperatures did lead to \suldiox~bonding coordinations that are more likely to form into cyclic hydrates.

These studies build upon our computational and experimental research in this area, seeking to understand how gases adsorb and transit across an aqueous/air interface. Such knowledge is invaluable for understanding land water and environmental aerosol systems where gaseous uptake behavior at a water surface surprises us and often defies our physical intuitions.\cite{Jayne1990,Yang2002,Worsnop1989,Boniface2000}

### 2011-11-30 Coordination-Results.tex
Conclusions! Finishing up the final edits before handing over to Geri.

@@
 the number of times each \suldiox~switched from one coordination type to another).
-With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} A and B, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one coordination to another.
+With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} A and B, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one bonding coordination to another.
 The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently).
@@
 In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes.
-The number of transitions within this ``S-SO-SOO'' group of coordinations is an order of magnitude larger than any other transition.
+The number of transitions within this ``S-SO-SOO'' group of bonding coordinations is an order of magnitude larger than any other transition.
 This indicates that while the \suldiox~is bound in any of the three most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other two coordinations in this group.
-Clearly, the \suldiox~is rather active and constantly forming and breaking bonds with its oxygens.
+Clearly, the \suldiox~is rather active and constantly forming and breaking bonds through its oxygens.
 
@@
 The three most populous coordinations, ``S'', ``SO'', ``SOO'' (as determined from Figure \ref{fig:bonding-coordinations} by percentage) in both temperatures have coordination lifetimes of up to 1 ps, in some instances lasting up to 1.5 ps.
-The brevity of lifespans overall speaks to the dynamic nature of the interface.
+The brevity of lifespans overall speaks to the dynamic nature of the \suldiox~surface binding.
 The length of time in each coordination parallels the populations of the coordinations, and suggests an ordering of steady states among various bonding coordinations.
@@
 Figure \ref{fig:coordination-lifespans} shows that the ``O'' coordinated lifetimes are shorter than the ``S'' coordinated ones.
-Figure \ref{fig:coordination-transitions} shows that the unbound-``O'' transition occurs almost three times more than the transition to the ``S''.
+Figure \ref{fig:coordination-transitions} shows that the unbound-``O'' transition occurs almost three times more than the transition to the ``S'' bonding coordination.
 The unbound \suldiox~forms a bond to a neighboring water through its oxygen, but that bond is short-lived and either quickly breaks (resulting in unbound \suldiox), or it transitions to the ``SO'' coordination by forming another bond through the sulfur.
@@
 As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the ``S-SO-SOO'' group of coordinations.
-This suggests that the \suldiox-sulfur interaction to a water oxygen has a much longer lifespan than the oxygen bonding to water hydrogens.
+This suggests that the \suldiox-sulfur interaction to a water oxygen has a much longer lifespan than the \suldiox-oxygen bonding to water hydrogens.
 The difference between the ``S-SO-SOO'' coordinations is an addition or removal of oxygen bonds, and the frequent transitions between them show that the bonds to \suldiox-oxygens are quickly forming and breaking.

### 2011-11-30 Cycle-Results.tex
Conclusions! Finishing up the final edits before handing over to Geri.

@@
-\subsection{Cyclic \suldiox~Hydrate Structures}
+\section {Cyclic \suldiox~Hydrate Structures}
 
@@
 (2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen.
-As noted earlier in the discussion of the graph-finding routine and the BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on the search's order of cycle discovery.
+As noted earlier in the discussion of the graph BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on the search's order of cycle discovery.
 The \suldiox~will be involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS.
@@
 
-\subsection {Cyclic Structure Energy Calculations}
-
 \subsection {Cyclic Hydrate Structure Lifetimes}
@@
 For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript.
-A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a particular \suldiox~hydrate bonding cycle was present or not, respectively.
+A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively.
 This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.
@@
 This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing.
-Many mechanical, electrical, and software solutions have been devised to suppress the transients or ``debounce'' the signal.
+Many mechanical, electrical, and software solutions have been devised to suppress the transients, or ``debounce'' the signal.
 Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before dissolution, where the $C(t)$ function bounces before settling into a final value.
-The bounce in the function, manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence.
+The bounce in the function manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence.
 Physically, the bond lengths within the cycle (or the cycle-to-be) are fluctuating back and forth across our bond-length criteria.

### 2011-12-01 Background.tex
Finishing up with edits returned from Geri. Getting ready to send out Draft 2.

@@
 
-A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water-hydrogens, or interacts via the sulfur atom with water-oxygens.
-To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters (those that lie in the topmost region of a gas/water interface), we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters.
+A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water-hydrogens, or interacts via the sulfur atom with water-oxygens.\cite{Baer2010,Bishenden1998,Steudel2009} To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters (those that lie in the topmost region of a gas/water interface), we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters.
 This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in more recent computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters.
@@
 
-In this work we adopt a nomenclature scheme for \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens.
+In this work we adopt a similar nomenclature scheme for \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens.
 Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water-oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox-oxygens.
-Analysis of the distribution of \suldiox~coordinations will give insight to how \suldiox~binds to the water surface.
+Analysis of the distribution of these various \suldiox~coordinations will give insight to how \suldiox~binds to the water surface.
 
@@
 
-Hydrated \suldiox~clusters have been studied extensively with many recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008, Jayne1990, Jayne1990a, Donaldson1995} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The first attempt to elucidate the structure of surface hydrated \suldiox~by Baer et al.
-showed that the bonding coordination distributions of \suldiox~are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''.
-In the same work they then focused on the dominant coordination to determine the most likely cluster geometry using two and three-water clusters bound to the \suldiox.
-However, none of the studies have yet focused directly on the presence of an extended hydrating structure forming closed rings of molecular interactions involving the solvated \suldiox~molecules.
+Hydrated \suldiox~clusters have been studied extensively with several recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008, Jayne1990, Jayne1990a, Donaldson1995} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The computational study to elucidate the structure of surface hydrated \suldiox~by Baer et al.
+showed that the bonding coordination distributions of \suldiox~at the surface are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''.
+In the same work they then focused on the dominant coordination to determine the most likely cluster geometry of di- and tri-hydrate species of \suldiox.
+However, that study, and others probing specific hydrate structures, were performed in gas phase, under idealized conditions following geometry optimizations.
+None of the studies have yet focused directly on the presence of an extended hydrating structure involving \suldiox~molecules at a water surface, forming closed rings of molecular interactions.
+Here we apply the ab initio molecular dynamics to recreate a microscopic water surface environment.
+By focusing on the occurence of a specific subset of \suldiox~bonding coordination types, we probe directly a certain hydrated \suldiox~in an aqueous interfacial environment.
 
-The cluster geometries found in this study imply a cyclic bonding structure through the two or three waters involved.
+The cluster geometries predicted in the study by Baer et al.
+imply a cyclic bonding structure through the two or three hydrating waters.
 ``Cyclic'' here is used to denote a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved.
@@
 \subsection {Graph Theoretical Details}
-The optimized geometry of the \suldiox~hydrates suggests cyclic bonding structures, but it remains a different story entirely when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface.
-Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in a static vacuum.
+
+As noted above, the optimized geometry of the \suldiox~hydrates suggests cyclic bonding structures.
+Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in the gas phase.
+A different story entirely has the potential to emerge when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface.
 Do the cyclic structures also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior?
@@
 The cyclic structures may involve any number of waters, but here each structure is shown with three waters.
-Type \Rmnum{3} is formed by a \suldiox~with the ``SO'' bonding coordination, which is the most dominant bonding coordination encountered.}
+Type \Rmnum{3} is formed by a \suldiox~with the ``SO'' bonding coordination, which is the most dominant bonding coordination encountered.
+Each hydrate structure is labeled with its corresponding bonding coordination type using the nomenclature described in the text.}
 		\label{fig:cyclic-structures}

### 2011-12-01 Conclusions.tex
Finishing up with edits returned from Geri. Getting ready to send out Draft 2.

@@
 Much has been learned about the energies of hydrate configurations, and the kinetics of gaseous uptake into aqueous systems.
-Yet, the specific nature of the adsorption process, including the various geometries, hydrate species, and bonding pathways remains largely unknown.
+Yet, the specific molecular nature of the adsorption process, including the various geometries, hydrate species, and bonding pathways remains largely unknown.
 As a gas transitions into the liquid water phase, it passes through a fluid interfacial region that remains poorly understood.
-Our understanding of the processes and chemistry of the interface is still in its infancy, but we gain unique new insights that are key to understanding many environmentally important processes at aqueous surfaces.
+Our understanding of the processes and chemistry of the interface is still in its infancy, but we are beginning to gain unique new insights that are key to understanding many environmentally important processes at aqueous surfaces.
 

### 2011-12-01 Coordination-Results.tex
Finishing up with edits returned from Geri. Getting ready to send out Draft 2.

@@
 We find two possibilities that may explain this difference.
-The ``O'' coordination may form more easily, but also break quickly after formation, accounting for the higher transition frequency.
+The single H-bonding of the ``O'' coordination may form more easily, but also break quickly after formation, accounting for the higher transition frequency.
 Otherwise, the ``O'' coordination may be the first step in adsorption of an unbound \suldiox, where a subsequent addition of an \suldiox-sulfur interaction to a water oxygen would lead to a transition to the most frequent coordination, ``SO''.
-In the latter case, any adsorption of \suldiox~may proceed through an oxygen binding, accounting for the increased unbound-``O'' transition frequency.
+In the latter case, any adsorption of \suldiox~proceeds through an oxygen binding, accounting for the increased unbound-``O'' transition frequency.
 

### 2011-12-01 Introduction.tex
Finishing up with edits returned from Geri. Getting ready to send out Draft 2.

@@
 
-One of Chemistry's many remaining ``black box'' events is the adsorption of gas molecules onto a water surface.
-Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions.
+The molecular nature of adsorption of gas molecules onto a water surface is one of the remaining largely uncharted territories of surface chemistry.
+Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions, and certainly less than what we know about gaseous adsorption on a solid surface.
 How does gas initially bind to a water surface, and what steps are involved in the subsequent adsorption?
@@
 What is the structure of hydrating waters in the surface region, and how does a hydrated solute molecule behave differently than as a gas?
-Experiments to address these questions provide valuable information, but can never fully characterize microscopic events and behaviors.
+Experiments to address these questions provide valuable information, but have to date never fully characterize microscopic events and behaviors.
 However, these systems can be fully characterized computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.
 
-As a model gas, \suldiox~has been used extensively in research because of its importance in commercial and environmental systems.
-\cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} Its molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
+\suldiox~is a particularly important gas to use as a starting point and model system because of its importance in commercial and environmental systems.
+\cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} Its simple molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions.
 A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in water's surface region.
@@
 In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water.
-We simulate a dynamic water surface with an extended hydrogen-bonding network that captures the variability of the \suldiox~hydrate structures, and the behavior of the water surface.
+We simulate a dynamic water surface, complete with all the extended hydrogen-bonding interactions that capture the variability of the \suldiox~hydrate structures, and the behavior of the water surface molecules.
 The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated.
@@
 We also examine \suldiox~binding behavior when bound to the surface waters.
-Lastly, our analysis demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.
+Lastly, our analysis of a specific bonding arrangement demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.

### 2011-12-05 Article.tex
New computational methods section.
Adding new sections
coordination breakdown image and more text about the coordinations.
partly finished the new coordination and temp change writeup.
Finished writing out the hand-notes for coordinations + rdfs, and the background for coordinations and cycles
Filling in the images
old?
removed line numbers
Conclusions! Finishing up the final edits before handing over to Geri.
Finishing up with edits returned from Geri. Getting ready to send out Draft 2.
Draft 3 sent to Geri, along with new abstract and title.

@@
 \documentclass{article}
+%\documentclass{scrartcl}
 
 \usepackage {amsmath}
+\usepackage {textcomp}
 \usepackage {setspace}
@@
 \usepackage[sort&compress]{achemso}
-\usepackage{lineno}
+%\usepackage{lineno}
+
+% For typesetting roman numerals
+\makeatletter
+\newcommand{\rmnum}[1]{\romannumeral #1}
+\newcommand{\Rmnum}[1]{\expandafter\@slowromancap\romannumeral #1@}
+\makeatother
 
 % some formatting tags
-\oddsidemargin  0.0in
-\evensidemargin 0.0in
-\textwidth      6.5in
+\usepackage[
+top    = 0.5in,
+bottom = 1.5in,
+left   = 2.50cm,
+right  = 2.50cm]{geometry}
+%\oddsidemargin  0.0in
+%\evensidemargin 0.0in
+%\topmargin -0.5in
+%\bottommargin -0.5in
+%\textwidth      6.5in
 
-\title{Temperature Effects on Surface Sulfur Dioxide}
+\title{Bonding, Intimacy, and Circles of Friends: Further Computational Studies of Sulfur Dioxide Adsorption to Aqueous Surfaces}
 \author{Eric S. Shamay \and Geraldine L. Richmond}
@@
 
-%\maketitle
+\maketitle
 
 %\onehalfspacing
-\linenumbers 
+%\linenumbers 
 \doublespacing
 
-%\begin{abstract}
-	%...Abstract goes here...
-%\end{abstract}
+\begin{abstract}
+A water surface is a dynamic and constantly evolving terrain producing a vast array of unique molecular properties and interactions with environmental chemical species.
+The dynamics of water surfaces permit life on earth to continue, but also keeps us from developing a complete microscopic picture of the specific behaviors that take place within interfacial aqueous environments.
+This computational study examines a piece of the water puzzle by elucidating the bonding, dynamic interactions, and hydrate structures of sulfur dioxide gas adsorbing to a water surface.
+Results described herein address the specific ways in which sulfur dioxide gas molecules bind to a water surface, and paint a more complete picture of the adsorption pathway than was previously developed from experimental and computational studies.
+Ab initio molecular dynamics have been employed to study sulfur dioxide and water interactions at two environmentally relevant temperatures on a water surface.
+The results of this study on a common environmental and industrially important gas provide molecular insight to aid our understanding of interactions on aqueous surfaces, and gaseous adsorption processes.
+\end{abstract}
 
-%\include{Introduction}
-%\include{Methods}
-%\include{Results}
-%\include{Conclusions}
+\include{Introduction}
+\include{Background}
+\include{ComputationalMethods}
+\include{Coordination-Results}
+\include{Cycle-Results}
+\include{Conclusions}
 

### 2011-12-05 Background.tex
Draft 3 sent to Geri, along with new abstract and title.

@@
 Here we apply the ab initio molecular dynamics to recreate a microscopic water surface environment.
-By focusing on the occurence of a specific subset of \suldiox~bonding coordination types, we probe directly a certain hydrated \suldiox~in an aqueous interfacial environment.
+By focusing on the occurrence of a specific subset of \suldiox~bonding coordination types, we probe directly a certain hydrated \suldiox~in an aqueous interfacial environment.
 
-The cluster geometries predicted in the study by Baer et al.
+The gas phase cluster geometries predicted in the study by Baer et al.
 imply a cyclic bonding structure through the two or three hydrating waters.
@@
 
-As noted above, the optimized geometry of the \suldiox~hydrates suggests cyclic bonding structures.
-Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters in the gas phase.
+As noted above, the optimized geometry of the gas phase \suldiox~hydrates suggests cyclic bonding structures.
+Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters.
 A different story entirely has the potential to emerge when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface.
@@
 Types \Rmnum{4} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle types formed near the \suldiox.
-Type \Rmnum{3} is of particular interest because the \suldiox~in this cycle has the most frequently occuring bonding coordination (``SO'', as shown later).
+Type \Rmnum{3} is of particular interest because the \suldiox~in this cycle has the most frequently occurring bonding coordination (``SO'', as shown later).
 

### 2011-12-05 Coordination-Results.tex
Draft 3 sent to Geri, along with new abstract and title.

@@
 ``SS'', ``SSO'', ``SSOO'', etc.).
-We see from the data that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones.
+From the data it is clear that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones.
 Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' together represent less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
 
-We can now form a picture of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study.
+A picture can now be formed of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study.
 The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens.
@@
 The increased temperature decreases the time spent in the unbound coordination, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature populations.
-We see that the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to coordinations with double-oxygen and double-sulfur bonds.
+As shown, the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to coordinations with double-oxygen and double-sulfur bonds.
 
@@
 
-To verify if the ``O'' coordination forms from, and breaks quickly to the unbound coordination as is suggested by the transition frequency plots in Figure \ref{fig:coordination-transitions}, we plot the lifespans of the various coordinations in Figure \ref{fig:coordination-lifespans}.
+To verify if the ``O'' coordination forms from, and breaks quickly to the unbound coordination as is suggested by the transition frequency plots in Figure \ref{fig:coordination-transitions}, the lifespans of the various coordinations are plotted in Figure \ref{fig:coordination-lifespans}.
 Each point in the plot represents a time during the simulation in which the \suldiox~formed the respective coordination.
@@
 The few lifespans above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that it also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''.
-Those few long lived unbound data points indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement occurs and it is drawn back to the surface to rebind.
+Those data points of a few long lived unbound species indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement occurs and it is drawn back to the surface to rebind.
 
@@
 This suggests that the \suldiox-sulfur interaction to a water oxygen has a much longer lifespan than the \suldiox-oxygen bonding to water hydrogens.
-The difference between the ``S-SO-SOO'' coordinations is an addition or removal of oxygen bonds, and the frequent transitions between them show that the bonds to \suldiox-oxygens are quickly forming and breaking.
+The difference between the ``S-SO-SOO'' coordinations is an addition or removal of oxygen bonds.
+The frequent transitions between them show that the bonds to \suldiox-oxygens are quickly forming and breaking.
 For the \suldiox-sulfur interaction to break, the \suldiox~must enter a non-sulfur coordination (e.g.

### 2011-12-05 Introduction.tex
Draft 3 sent to Geri, along with new abstract and title.

@@
 
-The molecular nature of adsorption of gas molecules onto a water surface is one of the remaining largely uncharted territories of surface chemistry.
+The molecular nature of the adsorption of gas molecules onto a water surface is one of the remaining largely uncharted territories of surface chemistry.
 Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions, and certainly less than what we know about gaseous adsorption on a solid surface.
-How does gas initially bind to a water surface, and what steps are involved in the subsequent adsorption?
+How does a gas initially bind to a water surface, and what steps are involved in the subsequent adsorption?
 How does an unbound gas molecule near a water surface affect the water to which it will bind?

### 2011-12-07 submission/Article.tex
Submitted the paper for review. Added the submission directory

\documentclass{article}

\usepackage {amsmath}
\usepackage {textcomp}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}

% For typesetting roman numerals
\makeatletter
\newcommand{\rmnum}[1]{\romannumeral #1}
\newcommand{\Rmnum}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother

% some formatting tags
\usepackage[
top    = 0.5in,
bottom = 1.5in,
left   = 2.50cm,
right  = 2.50cm]{geometry}

\title{Staying Hydrated: The Molecular Journey of Gaseous Sulfur Dioxide to a Water Surface}
\author{Eric S. Shamay \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

\maketitle

\doublespacing

\begin{abstract}
A water surface is a dynamic and constantly evolving terrain producing a vast array of unique molecular properties and interactions with chemical species in the environment. The complex dynamics of water surfaces permit life on earth to continue, but also complicate the development of a complete microscopic picture of the specific behaviors that take place within interfacial aqueous environments. This computational study examines a piece of the water puzzle by elucidating the bonding, dynamic interactions, and hydrate structures of sulfur dioxide gas adsorbing to a water surface. Results described herein address the specific ways in which sulfur dioxide gas molecules bind to a water surface, and paint a more complete picture of the adsorption pathway than was previously developed from experimental and computational studies. Ab initio molecular dynamics have been employed to study sulfur dioxide and water interactions at two environmentally relevant temperatures on a water surface. The results of this study on a common environmental and industrially important gas provide molecular insight to aid our understanding of interactions on aqueous surfaces, and gaseous adsorption processes.
\end{abstract}

\section {Introduction}

The molecular nature of the adsorption of gas molecules onto a water surface is one of the remaining largely uncharted territories of surface chemistry. Although gas uptake into aqueous systems occurs often environmentally and industrially, we still know very little about the process and the details of the adsorption reactions, and certainly less than what we know about gaseous adsorption on a solid surface. How does a gas initially bind to a water surface, and what steps are involved in the subsequent adsorption? How does an unbound gas molecule near a water surface affect the water to which it will bind? What is the structure of hydrating waters in the surface region, and how does a hydrated solute molecule behave differently than as a gas? Experiments to address these questions provide valuable information, but have to date never fully characterize microscopic events and behaviors. However, these systems can be fully characterized computationally, and when coupled to the previous experimental work can provide a much more complete picture of gaseous adsorption to aqueous surfaces.

\suldiox~is a particularly important gas to use as a starting point and model system because of its importance in commercial and environmental systems. \cite{Boniface2000,Jayne1990,Johns2011,Heber1997,Faloona2009,Clegg2001} Its simple molecular structure, high solubility in water, and relative abundance make it a pivotal compound in numerous aqueous atmospheric reactions. A complete picture of the \suldiox/\wat~adsorption process will aide in understanding gaseous adsorption on the many aqueous surfaces in the environment, as well as in understanding the fundamental nature of gases in water's surface region.

In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water. We simulate a dynamic water surface, complete with all the extended hydrogen-bonding interactions that capture the variability of the \suldiox~hydrate structures, and the behavior of the water surface molecules. The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated. In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas. Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas-binding and surface behavior.

%Unlike small cluster studies that use DFT calculations on geometry optimized structures to find likely geometries in a static vacuum setting, The quantum MD technique allows us to simulate the process much more realistically, and with greater accuracy than our previous classical MD work could allow.\cite{Shamay2011} This computational study greatly enhances the picture developed in our previous work on \suldiox.

Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.\cite{Baer2010} Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 273K. This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Although conclusions regarding the specific nature of those complexes could only be inferred from the experiments, our current computational studies now provide us with insights about the specific microscopic geometries and behaviors of the hydrating complexes.

We believe this to be the first temperature study using quantum MD to study the binding of small gas molecules on a water surface. We show how temperature affects the bonding behavior of the surface-adsorbed \suldiox~to neighboring waters, and propose a sequential binding mechanism for \suldiox~adsorbing to a water surface. We also examine \suldiox~binding behavior when bound to the surface waters. Lastly, our analysis of a specific bonding arrangement demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.

%background
\subsection {Bonding Coordination}

A hydrated \suldiox~in an aqueous environment forms hydrogen bonds through the oxygens to nearby water-hydrogens, or interacts via the sulfur atom with water-oxygens.\cite{Baer2010,Bishenden1998,Steudel2009} To further our analysis of the way in which \suldiox~coordinates its bonding to surface waters (those that lie in the topmost region of a gas/water interface), we adopt a naming scheme to denote the way in which the \suldiox~is hydrated by the surrounding waters. This naming scheme mimics a notational system developed during a previous study on water coordination by Buch et al,\cite{Buch 2005} and was subsequently used in more recent computational work.\cite{Walker2006b} In this naming system, a letter is used to designate the atom on a water molecule through which a hydrogen bond is formed to neighboring waters. Thus a bonding coordination of ``OOH'' designates two proton-acceptor bonding interactions through the water-oxygen, and a single proton-donor bonding interaction through a hydrogen. More recently, Baer et al. devised a nomenclature that explicitly enumerates the bonding to \suldiox~via the sulfur or oxygen atoms.\cite{Baer2010}

In this work we adopt a similar nomenclature scheme for \suldiox~in order to quantify hydrogen bonding through the acceptor \suldiox-oxygens, and the weaker bonding interactions from the \suldiox-sulfur to water-oxygens. Thus, an ``SOO'' coordinated \suldiox~molecule forms a single interaction through the sulfur atom to a neighboring water-oxygen, and two hydrogen bonds through either a single \suldiox-oxygen, or distributed with one hydrogen bond on each of the \suldiox-oxygens. Analysis of the distribution of these various \suldiox~coordinations will give insight to how \suldiox~binds to the water surface.

To determine \suldiox~bonding coordinations, we defined intermolecular bonds using the distance criteria of Baer et al.\cite{Baer2010} The bond-length definition is based on a set of distance criteria where a bonding interaction between a \wat-oxygen and \suldiox-sulfur is formed at a distance less than 3.5 \angs, and an \suldiox-oxygen hydrogen bond to a \wat-hydrogen is formed at a distance less than 2.2 \angs.

\subsection {Cyclic Bonding Structures}

Hydrated \suldiox~clusters have been studied extensively with several recent experiments and computations forming a clearer picture of \suldiox~bulk and surface behaviors.\cite{Baer2010, Tarbuck2005, Tarbuck2006, Ota2011, Bishenden1998, Hirabayashi2006, Steudel2009, Yang2002, Hayashi1985, Moin2011, Eckl2008, Jayne1990, Jayne1990a, Donaldson1995} At a water surface, it is now known that \suldiox~forms a complex with water during adsorption, and then subsequently absorbs into the interfacial region by reaction to form ionic sulfur species.\cite{Tarbuck2005, Tarbuck2006, Ota2011} The computational study to elucidate the structure of surface hydrated \suldiox~by Baer et al. showed that the bonding coordination distributions of \suldiox~at the surface are altered relative to the bulk region, and that two coordinations dominate the distribution of bonding types: the ``SO'' and the ``SOO''. In the same work they then focused on the dominant coordination to determine the most likely cluster geometry of di- and tri-hydrate species of \suldiox. However, that study, and others probing specific hydrate structures, were performed in gas phase, under idealized conditions following geometry optimizations. None of the studies have yet focused directly on the presence of an extended hydrating structure involving \suldiox~molecules at a water surface, forming closed rings of molecular interactions. Here we apply the ab initio molecular dynamics to recreate a microscopic water surface environment. By focusing on the occurrence of a specific subset of \suldiox~bonding coordination types, we probe directly a certain hydrated \suldiox~in an aqueous interfacial environment.

The gas phase cluster geometries predicted in the study by Baer et al. imply a cyclic bonding structure through the two or three hydrating waters. ``Cyclic'' here is used to denote a closed loop formed by the intermolecular hydrogen bonds, S-O interactions, and covalent bonds of the molecules involved. Figure \ref{fig:cyclic-example} depicts one such cyclic structure showing the bonds beginning on the sulfur and returning through a \suldiox-oxygen.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{double-cycle-type2-small.png}
		\caption{During the course of MD, the covalent and intermolecular bonds between a \suldiox~and the hydrating waters may form into a ring, resulting in a cyclic hydrate structure. Depicted here is one example of such a cyclic structure, formed by the covalent SO and OH bonds of the two molecule types, the intermolecular hydrogen bonds, and S-O$_{H_2O}$ interactions. The atoms involved are the S (yellow), O (red), and H (white).}
		\label{fig:cyclic-example}
	\end{center}
\end{figure}

\subsection {Graph Theoretical Details}

As noted above, the optimized geometry of the gas phase \suldiox~hydrates suggests cyclic bonding structures. Geometry optimization shows the formation of these cyclic hydrate structures with two or three waters. A different story entirely has the potential to emerge when \suldiox~is placed in a dynamic environment such as in the course of MD simulations of an aqueous surface. Do the cyclic structures also form in the course of a dynamic bonding process on a simulated water surface, where extended hydrating structures influence \suldiox~and water behavior? To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data. Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998}

Here we briefly introduce graph theoretical concepts. They have been described well by others with varied application to cyclic structures.\cite{Tutte1984, Balakrishnan2000, Harary1973, Huber2007, Garcia2004, Dury2001} A graph consists of nodes, and edges that connect the nodes. A molecule can be represented with atoms as nodes, and edges for each intramolecular covalent bond connecting the atoms. The set of edges is then further expanded to include intermolecular interactions such as hydrogen bonds and other bonding interactions. Edges may be assigned weights (i.e. bond lengths), types, and can be directional, i.e. pointing from a source node towards a target node. A molecular system including all atoms, bonds, and interactions is thus fully described by a graph.

To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A graph search is a recursive algorithm of queuing nodes and all neighboring nodes while performing a specified procedure on each visited node. This is easily performed on adjacency list or connectivity matrix data structures, iterating through nodes (i.e. atoms) of interest in the graph as starting points of the search. In graph search terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black). Performing a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node. A benefit of BFS is the ability to determine the smallest cyclic structure containing a given node. In the case of \suldiox~hydrate structures, beginning the BFS with the \suldiox-sulfur as the starting, or root node for the search, will discover cyclic bonding structures in order of size. Here we are only concerned with the smallest cyclic structure involving those waters in the first and second hydration shells around the \suldiox. Furthermore, it is possible to reconstruct a cycle's structure by finding its size (number of contributing atoms), and the number of unique waters in the cycle. This allows us to distinguish between various types of cyclic structures encountered.

Several arrangements of cyclic bonding structures are shown in Figure \ref{fig:cyclic-structures} for a \suldiox~molecule with three waters. Cycles with fewer or greater numbers of waters are also possible and encountered during MD. Cycle types \Rmnum{1}, \Rmnum{2}, \Rmnum{3} in Figure \ref{fig:cyclic-structures} are cyclic structures in which the \suldiox~is a member of the cycle. Types \Rmnum{4} and \Rmnum{5} do not involve the \suldiox~in the bonding cycle, but are commonly encountered as the smallest cycle types formed near the \suldiox. Type \Rmnum{3} is of particular interest because the \suldiox~in this cycle has the most frequently occurring bonding coordination (``SO'', as shown later).

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{cycle-types-small.png}
		\caption{\suldiox~in various cyclic structures encountered during MD simulations with water. The cartoons show the five types of cyclic structures, numbered for reference. The cyclic structures may involve any number of waters, but here each structure is shown with three waters. Type \Rmnum{3} is formed by a \suldiox~with the ``SO'' bonding coordination, which is the most dominant bonding coordination encountered. Each hydrate structure is labeled with its corresponding bonding coordination type using the nomenclature described in the text.}
		\label{fig:cyclic-structures}
	\end{center}
\end{figure}

Baer et al. presented a detailed geometric and spectroscopic breakdown of type \Rmnum{3} cycles with two and three waters from their DFT calculations.\cite{Baer2010} Given the information of the number of waters, atoms, and bonds involved in the bonding cycles, we find that of the three-water type \Rmnum{3} cycles, there exist two structural varieties, shown in Figure \ref{fig:type-3-varieties}, that differ in the set of water atoms involved in the cyclic structure. Type \Rmnum{3}-A (shown in Figure \ref{fig:type-3-varieties}A) is arranged with each water contributing an OH bond to the structure of the cycle. Type \Rmnum{3}-B involves a single water contributing an OH, whereas the other two waters contribute only the oxygen atom or the entire water molecule to the structure, respectively. This nuance of the type \Rmnum{3} structures involving three waters, and the overall distribution of structures are presented in more detail later. We also show the distribution of cyclic structures encountered during MD simulations to further understand the behaviors of \suldiox-hydrates at the water surface.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{triple-cycle-types-small.png}
		\caption{Type \Rmnum{3} cyclic structures (see Figure \ref{fig:cyclic-structures}) are found to occur in two varieties that are distinguished by the water atoms contributing to the cyclic structure. In ``type A'' the cycle is formed by an OH bond contributed by each water, and the SO bond of the \suldiox. ``Type B'' involves an oxygen atom from one water, an OH bond from a second water, and all atoms of the third water.}
		\label{fig:type-3-varieties}
	\end{center}
\end{figure}

%comp methods
\section{Computational Methods}

On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions. The electronic charge density was described using an auxiliary basis set of plane waves. Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively. These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.

Initially, 10 equilibrated boxes of side-lengths 10.0\angs, with 36 randomly packed water molecules were used. Five of the boxes were used for each of the cold and hot simulations. A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis. A copy of the initial system cubes were then expanded along one axis (z-axis) to 25\angs. The system energy was minimized through a geometry optimization. Subsequently, the system was equilibrated for 1 ns in canonical ensemble (NVT) conditions. Periodic boundaries were set on the two short axes to form an infinite slab. The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1 fs. The initial 1 ns equilibration trajectory was not included in the final analysis. This simulation process resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system, for a total of 100,000 timesteps at each temperature.

% coordination results
\section {Sulfur Dioxide Bonding Coordinations}

The bonding coordination of each \suldiox~was determined at each timestep of the simulations. Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, as a percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories. A first visual inspection reveals several trends. Clearly the ``SO'' coordination is the most populous at both temperatures. The second and third most populated coordinations are the ``S'' and ``SOO'', however their distributions differ between temperatures. In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally. The hotter temperature simulation shifts the distribution such that the ``S'' occurs 5\% less frequently than in the cold, and the ``SOO'' occurs nearly 10\% more often. The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``SOO'', respectively. These results for the hot (room temperature) system coincide with those of the previous single-temperature simulation study by Baer et al. at a similar temperature.\cite{Baer2010}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{so2-coordinations-percents-small.png}
		\caption{The \suldiox~bonding coordinations occur with different frequencies at the surface under hot and cold temperature conditions. Shown above is the distribution of bonding coordinations of the cold (blue) and hot (red) \suldiox~over the entire set of trajectories. The values are the percentage of MD timesteps spent in the given bonding coordination.}
		\label{fig:bonding-coordinations}
	\end{center}
\end{figure}

	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from this distribution of coordinations. Clearly the cold \suldiox~spends more time than the hot \suldiox, 13\% versus 3\%, respectively, completely unbound from the surface waters. This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis. Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures. Baer et al. also concluded that these three coordinations were the most frequent for a room temperature simulation, and specifically identified the ``SO'' and ``SOO'' as most common in their study.\cite{Baer2010}

	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (e.g. ``O'', ``OO'', etc.) represent the least frequently formed. A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters. In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.

% Baer's assymetric oxygen binding
Baer et al performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens. They concluded that there is asymmetric hydrogen bonding through the \suldiox-oxygens, with one oxygen binding more often than the other. This is supported by the findings here where all the double oxygen coordinations (e.g. ``OO'', ``SOO'', etc.) represent a much lower percentage of the coordinations than the single oxygen counterparts (e.g. ``O'', ``SO'', etc.). Furthermore, a triple-oxygen coordination (e.g. ``OOO'', ``SOOO'', etc.) is very rarely encountered. Three \suldiox-oxygen bonds only form if both \suldiox-oxygens are interacting with water hydrogens. Our finding that triple-oxygen coordinations rarely form complements Baer et al's conclusion about the asymmetry in the oxygen interactions.

Having established the preference for an interaction through the \suldiox-sulfur atom, we look to the right side of Figure \ref{fig:bonding-coordinations} at the double-sulfur coordinations (e.g. ``SS'', ``SSO'', ``SSOO'', etc.). From the data it is clear that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones. Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' together represent less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.

A picture can now be formed of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study. The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens. The \suldiox-oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.

\subsection {Temperature Effects on Bonding Coordinations}

The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in bonding coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot. In the cold temperature, the unbound, ``S'', and ``SO'' coordinations are more populated than in the hot systems. The increased temperature decreases the time spent in the unbound coordination, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature populations. As shown, the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to coordinations with double-oxygen and double-sulfur bonds.

On the cold water surface, the ``S'' and ``SO'' are more populated than for the hot system. The relative decrease of these coordinations are matched in the hot surface by an increase of the ``SOO'' configuration. This speaks to a dramatic difference in the surface behavior of \suldiox~at the two temperatures. The cold \suldiox~spends nearly equal time in the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''. Thus, the addition or removal of a bond through the \suldiox-oxygen to a neighboring coordination (e.g. addition of an oxygen bond from ``SO'' to ``SOO'', or removal of the bond from ``SO'' to ``S'') is equally probable, as long as the sulfur interaction with the water oxygen does not break.

Figure \ref{fig:rdf} shows the radial distribution functions (RDF) of \suldiox~to water atoms for both cold (blue) and hot (red) temperatures. The S-O$_{H_2O}$ RDFs are nearly equal except for a slightly taller first peak in the cold system. Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF indicates that since bonding occurs frequently through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface waters. In the hot systems, the bonding coordinations with two oxygen bonds (e.g. ``SOO'', etc.) occur more frequently than in the cold system. This additional bonding through a second oxygen interaction may slightly shift neighboring hydrating waters away from the \suldiox-sulfur towards the oxygen end of the molecule. This suggests that the cold \suldiox~bonds closer to the surface waters through its sulfur atom (as shown in the S-O$_{H_2O}$ RDF) and favors the more ``sulfur-centric'' bonding coordinations (e.g. ``S'', ``SO''). The increased temperature of the hot system allows the \suldiox~to bond more extensively through its oxygens to the ``SOO'' coordination. The greater interactions through the oxygens, and higher bonding coordination, may pull the \suldiox~further into the water interface and then allow for increased bonding through the sulfur, up to double-sulfur coordinations (e.g. ``SS'', ``SSO'', etc.).

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{rdf-small.png}
		\caption{Radial distribution functions (RDF) show the subtle change in how the hydrating waters around \suldiox~reposition as the temperature is increased from cold (blue) to hot (green). Shown are the two correlations between sulfur and water oxygens (left) and the \suldiox-oxygen and water hydrogens (right).}
		\label{fig:rdf}
	\end{center}
\end{figure}

\subsection {Bonding Transitions}

During the course of each simulation, the \suldiox~bonding coordination was determined and recorded for each timestep. From the coordination data we extract not only the populations of the various bonding coordinations, but also the frequencies of transitions between the different coordinations (i.e. the number of times each \suldiox~switched from one coordination type to another). With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} A and B, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one bonding coordination to another. The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently). Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequent transitions. Additionally, the transition lines are numbered to the right of each line with the number of times each transition occurred.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{coordination-transitions.png}
		\caption{The bonding coordinations of \suldiox~on a water surface rapidly change as bonds break and form throughout the MD trajectories. Depicted here is a graph of the bonding coordinations, shown as colored nodes of varying size. Larger and more darkly colored coordination nodes are those more often encountered during MD (see Figure \ref{fig:bonding-coordinations}). The directed edges between the nodes represent the number of times the coordination transition occurred. Thicker lines correspond to a greater number of transitions between coordinations. The number of times a given transition occurred is labeled to the right of the corresponding line.}
		\label{fig:coordination-transitions}
	\end{center}
\end{figure}

As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} than transitions to less populated coordinations. Insights to the bonding process are made clearer from these graphs. In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes. The number of transitions within this ``S-SO-SOO'' group of bonding coordinations is an order of magnitude larger than any other transition. This indicates that while the \suldiox~is bound in any of the three most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other two coordinations in this group. Clearly, the \suldiox~is rather active and constantly forming and breaking bonds through its oxygens.

In the graph of the hot system in Figure \ref{fig:coordination-transitions}B, the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size. One very surprising result is in the transition from ``SOO-SOOO''. This transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations. This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox-oxygen. As noted earlier, it is likely that ``SOO'' coordinated \suldiox, asymmetrically binding twice through a single oxygen, is being pulled further into the water interface. It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox-oxygen. Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox~returns to the ``SOO'' coordination.

Given the information we have about the frequencies of bonding coordination transitions, it is possible to draw a likely route of adsorption beginning with an unbound \suldiox. From the unbound coordination, the \suldiox~can bind to waters either through the sulfur or an oxygen to enter the ``S'' or ``O'' coordinations, respectively. At both temperatures we find that the coordination transition in Figure \ref{fig:coordination-transitions} from unbound to ``O'' occurs almost three times more than the transition from unbound to ``S''. We find two possibilities that may explain this difference. The single H-bonding of the ``O'' coordination may form more easily, but also break quickly after formation, accounting for the higher transition frequency. Otherwise, the ``O'' coordination may be the first step in adsorption of an unbound \suldiox, where a subsequent addition of an \suldiox-sulfur interaction to a water oxygen would lead to a transition to the most frequent coordination, ``SO''. In the latter case, any adsorption of \suldiox~proceeds through an oxygen binding, accounting for the increased unbound-``O'' transition frequency.

To verify if the ``O'' coordination forms from, and breaks quickly to the unbound coordination as is suggested by the transition frequency plots in Figure \ref{fig:coordination-transitions}, the lifespans of the various coordinations are plotted in Figure \ref{fig:coordination-lifespans}. Each point in the plot represents a time during the simulation in which the \suldiox~formed the respective coordination. The vertical ``lifespan'' position is calculated directly from the amount of time spent in the given coordination before changing to another. Both cold (blue) and hot (red) data are plotted. The data of Figure \ref{fig:coordination-lifespans} show that most coordination configurations last a very brief time, with the majority forming for under 0.5 ps. The three most populous coordinations, ``S'', ``SO'', ``SOO'' (as determined from Figure \ref{fig:bonding-coordinations} by percentage) in both temperatures have coordination lifetimes of up to 1 ps, in some instances lasting up to 1.5 ps. The brevity of lifespans overall speaks to the dynamic nature of the \suldiox~surface binding. The length of time in each coordination parallels the populations of the coordinations, and suggests an ordering of steady states among various bonding coordinations. The ``unbound'' configuration stands out as an anomaly amongst the lifespans of the other bonding coordinations. The few lifespans above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that it also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''. Those data points of a few long lived unbound species indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement occurs and it is drawn back to the surface to rebind.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{coordination-lifespans.png}
		\caption{Given that certain bonding coordinations of the surface \suldiox~are more often encountered than others (see Figure \ref{fig:bonding-coordinations}), the plot here shows the time spent in each bonding coordination. Each point in the plot represents a single span of simulation time that the \suldiox~had the given coordination. The total amount of consecutive timesteps in a coordination corresponds to the vertical position along the lifespan axis, in ps.}
		\label{fig:coordination-lifespans}
	\end{center}
\end{figure}

Returning to the transition plots of Figure \ref{fig:coordination-transitions}, the behavior of the unbound transition to both ``S'' and ``O'' coordination can now be better characterized. Figure \ref{fig:coordination-lifespans} shows that the ``O'' coordinated lifetimes are shorter than the ``S'' coordinated ones. Figure \ref{fig:coordination-transitions} shows that the unbound-``O'' transition occurs almost three times more than the transition to the ``S'' bonding coordination. The unbound \suldiox~forms a bond to a neighboring water through its oxygen, but that bond is short-lived and either quickly breaks (resulting in unbound \suldiox), or it transitions to the ``SO'' coordination by forming another bond through the sulfur. The unbound-``S'' transition does not occur as often. This may be because the ``S'' coordination is more stable than the ``O''. Once in the ``S'' coordination the \suldiox~does not quickly break the interaction from its sulfur to water, but rather remains for up to 1 ps in the ``S'' coordination before (most likely) forming an oxygen bond to make the ``SO'' coordination. We have outlined the likely behavior of \suldiox~as it transitions from the gas phase in an unbound coordination to binding to an aqueous interface by forming bonds and interactions with surface waters.

Once the \suldiox~begins interacting with the water surface, the pathway leading back to the unbound coordination is not often traversed. As shown in Figure \ref{fig:coordination-transitions} the dominant coordination transitions occur between the ``S-SO-SOO'' group of coordinations. This suggests that the \suldiox-sulfur interaction to a water oxygen has a much longer lifespan than the \suldiox-oxygen bonding to water hydrogens. The difference between the ``S-SO-SOO'' coordinations is an addition or removal of oxygen bonds. The frequent transitions between them show that the bonds to \suldiox-oxygens are quickly forming and breaking. For the \suldiox-sulfur interaction to break, the \suldiox~must enter a non-sulfur coordination (e.g. ``O'', ``OO'', unbound, etc.) or a coordination with more than a single sulfur interaction (e.g. ``SS'', ``SSO'', etc.). The transitions to coordinations that allow for breaking of the \suldiox-sulfur interactions, or switching the interaction to another water, are infrequent compared to those leading to an oxygen bond transition. Thus, the \suldiox~spends most of its time while bound to the surface waters breaking and forming hydrogen-bonds through its oxygens, and interacting with neighboring waters through a more persistent interaction via the \suldiox-sulfur atom.

% cycle results
\section {Cyclic \suldiox~Hydrate Structures}

Having examined the bonding coordinations and bonding behavior of \suldiox~with surface waters, we now turn to a secondary behavior of the hydrate structures that form around the surface-bound \suldiox~molecule. The simulation trajectory data was analyzed to determine the presence and characteristics of \suldiox~cyclic hydrate structures that form, as posited earlier in the text and depicted in Figure \ref{fig:cyclic-structures}. Only the most commonly occurring subset of the cyclic structures were analyzed based on two selection criteria: (1) The distances between atoms must match the same bonding/distance criteria as used for determining bonding coordinations. (2) The \suldiox~must be minimally in a bonding coordination of type ``SO'', meaning that the sulfur has at least one bonding interaction, and at least one hydrogen-bond must have formed with an oxygen to a neighboring water-hydrogen. As noted earlier in the discussion of the graph BFS algorithm, the cyclic structures found represent the smallest cycles in which the \suldiox~is a member, based on the search's order of cycle discovery. The \suldiox~will be involved in other larger and more extended cyclic bonding structures beyond the first one discovered via the BFS. The larger and more extended cyclic structures involving more waters affect the behavior of the hydrogen-bonding network of the water surface. We focus here only on the smallest cycles involving the \suldiox~as they most affect the \suldiox~bonding and hydration.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{SO-cycle-breakdown-with-cartoons-small.png}
		\caption{Cyclic hydrate structures form throughout the MD simulations involving the \suldiox~and one or more water molecules. Shown here are the different cycles by number of water (up to 4), and their occurrence rate as a percentage of the total simulation time. The far right part of the plot shows the contributions of the two types of triple-water cycles (see Figure \ref{fig:type-3-varieties}). The cycles had to meet the ``minimally-SO'' bonding criteria (at least one bond to the sulfur and one oxygen bond on the \suldiox) to be used in the analysis.}
		\label{fig:cyclic-breakdown}
	\end{center}
\end{figure}

The plot in Figure \ref{fig:cyclic-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures. Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4). The two tallest data points, left-most in the plot, show the overall time spent in all types of cyclic structures. Clearly the hot \suldiox~spends more time in a cyclic structure than at the cold temperature. It is remarkable that the time spent in a cyclic structure shows a 15\% difference (42\% cold, 57\% hot). The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such. Thus, in addition to having earlier found the most likely bonding coordination during the simulated life of \suldiox, we have also found that the hydrates of the \suldiox~form cyclic structures for much of the time while bound to the water surface.

Now we look at the different types of cyclic hydrates, distinguished by the number of waters involved in the bonding structure. In Figure \ref{fig:cyclic-breakdown} the single and double water cycles are the least frequently encountered structures, accounting for less than 10\% of both cold and hot temperature simulation times. Formation of the single-water type is likely energetically unfavorable because of the proximity of the water to the \suldiox~required to form the bonding cycle. The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In those static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form. However, the MD simulations here have introduced many waters into a dynamic environment allowing for extended bonding networks, and the results show clearly that the double-water cyclic hydrate is formed much less often (less than 5\% at both temperatures) than the triple-water form.

The results for triple and quadruple-water structures show that larger hydrate cycles are favored at higher temperatures. Although the higher number of hydrating waters ($>$ 4) are not shown, those contribute minimally to the overall distribution. The majority of the cyclic hydrates are formed with three waters in the triple-water type. This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that, of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}). The triple-water type cycle results were broken-down into contributions from the type-A and type-B triple-water structures. The two triple cycle structures are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of their contributions to the overall distribution. It is notable that each temperature has a different dominant type of triple-water cyclic structure. The cold system forms more of the type-B, and the hotter system forms primarily type-A.

\subsection {Cyclic Hydrate Structure Lifetimes}

We know that the \suldiox~bound to a water surface is most likely in the ``SO'' bonding coordination, and is also often taking part in some type of cyclic structure. With this in mind, it becomes interesting to ask: how long does a cyclic hydrate form before breaking to an acyclic hydrate structure? To answer the question of cyclic lifespan, a method was devised to define the lifetime of a cycle. For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript. A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively. This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.

In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value. This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing. Many mechanical, electrical, and software solutions have been devised to suppress the transients, or ``debounce'' the signal. Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before dissolution, where the $C(t)$ function bounces before settling into a final value. The bounce in the function manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence. Physically, the bond lengths within the cycle (or the cycle-to-be) are fluctuating back and forth across our bond-length criteria. It is thus an artifact of our algorithmic determination of the presence or absence of a bond. Removing this artifact will allow us to examine the longer-time bonding behavior and this is accomplished as follows.

We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions. The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states. $f(t)$ may then be used to calculate statistics about the cycle lifespans. A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}. At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time. The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly. This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each. $C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration. The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:

\[
  f(t) = \left\{ 
  \begin{array}{l l}
    0 & \quad C_s(t) < 0.2\\
    1 & \quad C_s(t) \geq 0.2\\
  \end{array} \right.
\]

where $f(t)$ is the debounced time-function that represents the lifespans of cycles. Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{cycle-debouncing-small.png}
		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures. A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure. A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure. A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
		\label{fig:debouncing}
	\end{center}
\end{figure}

The distribution of cycle lifetimes (i.e. contiguous spans of time spent with $f(t) = 1$) was determined. The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations. The most frequent lifespan for both temperature data sets lasts between 0-1 ps. This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time. Even with the debouncing procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps. The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend extending to the longer lifespans as well. The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps. All of the distribution shows only a $<1.5$\% difference between cold and hot cyclic lifetimes. Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time. Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles. Above 4 ps, most of the cycles that form are found in the hot system. The 8 ps cycles are notable in that they form for just under half the length of one of the simulated trajectories.

We know from the transition frequency plots of Figure \ref{fig:coordination-transitions} that the bonding coordinations are switching frequently. The \suldiox~is likely forming and breaking bonds with waters external to the cyclic hydrate structures (i.e. not directly involved in the bonds of the cycle). For the longer-lived cycles, the external bonding to the \suldiox~may have little effect on the cyclic hydrate waters. However, any time the \suldiox~switches into an unbound, sulfur-only, or oxygen-only coordination (i.e. ``S'', ``SS'', ``O'', etc), the cyclic structure is necessarily broken. Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure. Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively. Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold). Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox). This is relative to the cold system coordination distribution that has equal ``S'' and ``SOO'' populations. The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures. This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{cyclic-lifespans-inset-small.png}
		\caption{The lifetime of a cyclic hydrate structure after it forms, and before it distorts in shape due to the movement of the liquid surface molecules, is most often short ($<$ 1 ps). Shown here is a plot of cycle hydrate structure lifetimes after having applied the debouncing procedure (described in the text). Both cold (blue) and hot (red) simulations are graphed. The inset expands the view of the region above 1 ps.}
		\label{fig:cycle-lifespans}
	\end{center}
\end{figure}

%conclusions
\section {Conclusions}

The adsorption of small gas molecules to water surfaces has been extensively studied over the past few decades. Much has been learned about the energies of hydrate configurations, and the kinetics of gaseous uptake into aqueous systems. Yet, the specific molecular nature of the adsorption process, including the various geometries, hydrate species, and bonding pathways remains largely unknown. As a gas transitions into the liquid water phase, it passes through a fluid interfacial region that remains poorly understood. Our understanding of the processes and chemistry of the interface is still in its infancy, but we are beginning to gain unique new insights that are key to understanding many environmentally important processes at aqueous surfaces.

Presented herein are the results of ab initio molecular dynamics simulations that focus on how a wandering gaseous \suldiox~molecule first makes contact with a water surface, and subsequently forms extended hydrate structures with interfacial water molecules. The computational studies complement and expand on experimental studies from this laboratory that found surface complexation of \suldiox~at a water surface.\cite{Tarbuck2005,Tarbuck2006,Ota2011} Furthermore, these computations build upon and enrich our understanding of adsorbing \suldiox~behavior from our recently published computational study on interfacial geometries of aqueous surface \suldiox~molecules.\cite{Shamay2011}

Our simulations show that \suldiox~has a preferred means of bonding and interacting with surface water molecules by taking on various bonding coordinations. In this work it was shown that the ``SO'' bonding configuration is the most preferred, with ``S'' and ``SOO'' also contributing greatly to the coordination distribution. Once a \suldiox~has bound to form a surface hydrate it rarely forms multiple bonding interactions through the sulfur atom, and even less frequently takes on a configuration with no sulfur interactions to nearby waters.

This study is one of very few temperature studies looking at the microscopic nature of interfacial gas molecules on water. By changing the temperature, it was found that a hotter water system leads to longer \suldiox~binding to the water surface. The distribution of bonding coordinations was greatly affected by a temperature change, shifting populations of bonding configurations because of the altered \suldiox~and \wat~behavior. At the higher temperature, \suldiox~forms more frequent bonds to interfacial waters through the sulfur and oxygen atoms. Overall, we have determined that the \suldiox~hydrate interactions are transient, binding and unbinding to water molecules rapidly in very dynamic bonding coordinations.

In this work we introduce the use of a graph structure to represent atoms and interconnectedness between molecules, and also to represent transitions between the various bonding coordinations of an adsorbed \suldiox. It was shown that the intermolecular bonds formed through the \suldiox-oxygens are quickly broken and formed, lasting briefly compared to sulfur interactions. From the graph of bonding coordination transitions, we found a likely pathway for \suldiox adsorption starting with an unbound gas-phase \suldiox, ending with a hydrated \suldiox~species bound to surface waters.

The formation of cyclic hydrate structures was probed and it was found that these hydrated bonding ring species form during much of a simulated trajectory. Temperature increases the occurrence of cyclic structures, and also shifts the distribution of the specific types of cycles being formed. Two types of cyclic tri-hydrates were discovered during the course of simulations. The cycle lifetimes were found to be mostly short-lived, with a majority lasting less than 1 ps before breaking and reforming due to the dynamic bonding and motion of the surface waters and \suldiox~molecules. Temperature did not have a very dramatic effect on the cyclic lifespans, but higher temperatures did lead to \suldiox~bonding coordinations that are more likely to form into cyclic hydrates.

These studies build upon our computational and experimental research in this area, seeking to understand how gases adsorb and transit across an aqueous/air interface. Such knowledge is invaluable for understanding land water and environmental aerosol systems where gaseous uptake behavior at a water surface surprises us and often defies our physical intuitions.\cite{Jayne1990,Yang2002,Worsnop1989,Boniface2000}

\section {Acknowledgements}

The authors would like to thank Dr. Daniel Steck of the University of Oregon Department of Physics for the generous use of computing resources, Dr. Kevin E. Johnson (Pacific University) and Dr. Fred Moore (Whitman College) for many informative discussions and insights, the National Science Foundation (Grant CHE-1051215), and the US Department of Education Graduate Assistance in Areas of National Need (GAANN) program (Grant P200A070436) for support of this research.

\bibliography{so2temperature}

\end{document}

### 2012-03-27 resubmission/ResponseToReviews.tex
Preparing the resubmission files. Working on responses to the reviewers.

Journal: The Journal of Physical Chemistry 
Manuscript ID: jp-2011-11804m 
Title: "Staying Hydrated: The Molecular Journey of Gaseous Sulfur Dioxide to a 
Water Surface" 
Author(s): Shamay, Eric; Richmond, Geraldine

Dear Prof. Richmond:

Thank you for submitting your manuscript for publication in The Journal of Physical Chemistry. It has been examined by expert reviewers who have concluded that the work is of potential interest to the readership of The Journal of Physical Chemistry; however, it appears that a major revision, possibly followed by further reviewer evaluation, will be needed prior to its further consideration for publication. Please see the enclosed reviewers' reports for details regarding the requested changes and/or additions.

Additionally, we have noticed that your manuscript referenced above is missing the Table of Contents graphic. Please see the submission requirements for the Journal of Physical Chemistry TOC Graphic at http://pubs.acs.org/paragonplus/submission/jpchax/jpchax_authguide.pdf. 
Note that the graphic for the TOC entry should be in the form of a structure, graphical image, photograph, reaction scheme, or equation. The Author must submit the graphic in the actual size to be used for the TOC, fitting in an area 1.375 in. high and 3.5 in. wide (3.6 cm Ã— 8.9 cm). The allowed file formats for TOC graphics are TIFF and PDF. Please submit a revision of this manuscript including the TOC graphic as the final page of your Article manuscript revision submission. Separate TOC graphics are not allowed.

We ask that you submit your revision no later than 29-Apr-2012.

When submitting your revised manuscript through ACS Paragon Plus, you will be able to respond to the reviewers' comments in the text box provided or by attaching a file containing your detailed responses outlining the changes made and explaining your reasons for disagreeing with any suggestions you choose not to follow.

To Revise Your Manuscript on the Web:

To revise your manuscript, log into ACS Paragon Plus at http://paragonplus.acs.org/login and select "My Authoring Activity". There you will find your manuscript title listed under "Revisions Requested by Editorial Office." Your original files are available to you when you upload your revised manuscript. If you are replacing files, please remove the old version of the file from the manuscript before uploading the new file.

Thank you for considering The Journal of Physical Chemistry as a forum for the publication of your work.

With sincere regards,

Prof. Pavel Jungwirth 
Senior Editor 
The Journal of Physical Chemistry 
Journal Fax: 202-354-5425 
Journal Email: jungwirth-office@jpc.acs.org

Reviewer: 1 
-----------

1,2,3. The authors have taken into consideration the entirety of all the reviewer suggestions regarding functional and basis sets, cutoffs, timesteps, and temperature that affect the simulations. The changes are reflected in the text, and meet the standards suggested in the reviews. All the simulation data has been re-analyzed, and the results are all presented from the new simulations.
% 1. The computational parameters chosen as well as the authorsʼ claim that “these computational parameters were verified to yield a reasonable description of bulk room temperature water …” runs counter to published results.

%2. Given that the system contains water and thus the highest vibrational modes are ~3700 cm-1, it is hard to fathom a timestep of “1 fs” can be used to accurately integrate the equation of motions without energy drift. If this is truly the case, it might behoove the authors to provide some supplemental materials showing the conserved quantity as a function of time for both the NVT and NVE simulations provided the system is not glassy and can sample all the possible conformations.

%3. The author stated that DZVP basis set was chosen, it was never explained why the authors believe that such a basis set is adequate. In fact, work by VandeVondele et al shows that at least a TZV2P basis set with a cutoff of 280 Ry was necessary in order to obtain forces comparable to other more established codes such as CPMD. In addition, works by Siepmann et al on vapor liquid coexistence curves hints that the standard TZV2P basis set with a cutoff of 280 Ry might still be inadequate. Last, what was the convergence criteria used for the SCF?

%4a. The cold system temperature was set to 210K in the resimulated systems. This much larger separation of temperatures is more in line with the reviewer's concerns regarding thermostatting. 
4. Following recent literatures on simulations of liquid water using potentials derived from density functional theory (albeit some with different codes as well as different methods), one firm conclusions can be made is that water near ambient conditions are more glassy than liquid. One way to overcome this is through massive thermostating which in addition also ensures equipartition of the kinetic energy in all degrees of freedom. This raises two issues.  a. First, as the production run was obtained in the microcanonical (NVE) ensemble and the two target temperature is only separated by 27 degrees, can the authors show that temperature for the subsystem of interest is really the targeted temperature thus fulfilling the equipartition theorem.  b. And can the authors show that indeed, the water part of the simulated system is behaving as a liquid and not conformationally trap in a glassy state?

5. “Nose” should be “Nosé”. In addition, the coupling frequency, when not chosen appropriately can cause energy drifts, was not specified.

6. The text has been updated to reflect the equilibration period of 1 ps.
%6. The authors indicate the system was equilibrated for “1 ns”. This just does not seem correct unless the equilibration was done using a cheaper potential other than DFT.

7. From the RDF, it can be seen that both hot and cold simulations converges to 1.0 as distance increases. But if the sulfur dioxide gas molecule is at the liquid-vapor interface, shouldnʼt this value approach something more like 0.5 instead of 1.0?

8. The inclusion of dispersion corrections were not explored for this simulation study. The goal of the project was to monitor specific bonding behaviors of SO2 using previously established methods, not as a study of the results from various constraints or changes to simulation parameters.
%8. Has the authors explore the ramifications of including dispersion correction in their calculation which might play a large in such a system?

Reviewer: 2 
-----------

Recommendation: This paper is probably publishable, but should be reviewed again in revised form before it is accepted.

Reviewer: 3
-----------

1. Because the re-simulated systems were done using B3LYP functional a benchmark comparison is no longer necessary. Small cluster studies of so2 with several waters have been performed using high-accuracy basis and functionals previously, and we do not set out to reproduce that work here, but rather to build on it. The authors believe that the current work with the new simulation parameters achieves a level of accuracy suggested by the reviewers, but that lengthy discussion of this fact is not requisite in this particular simulation study.
%1.	The predominant stabilization interaction between and SO2 molecule and a water molecule is due to dipoles. This essentially places the sulpher close to the water oxygen. This simple dipole interaction reasoning is consistent with everything the authors find. However, there is a component of this interaction that arises due to van der Waal’s, due to the sulpher lone pairs. In this regards BLYP has been used here, which is known not to capture such subtle interactions. Considering this, I would have liked to see at least a set of benchmarks for the functional and basis used in this work, with post Hartree-Fock MP2 and B3LYP used for comparison. The benchmark system could have involved a simple cluster of SO2 with a few water molecules. The absence of this comparison makes the results weak.

2.	The larger concentration of unbound SO2 at a lower temperature is counter-intuitive. (Fig. 4) One would have expected the opposite trend and no physical reasoning is provided to explain this result.

3. Several previous simulation studies have discussed the structure of water under similar simulated conditions. The focus of the work is on the structure of SO2 and its behavior near a water cluster surface. Thus, the authors have made the conscious decision to avoid lengthy discussion of the surrounding water's structure.
%3.	The structure of the surrounding water molecules is never discussed.

4.	It is stated that Quickstep is used. Assume this implies that the Car-Parrinello method is used, but it is not stated in the paper.

5.	The system includes 36 water molecules in a box. The effect of this restriction is important to discuss.

6. K-point integration was not used in the integration, as this was not addressed in similar previous studies. Also, the CP2K simulation package does not allow for this type of calculation. The authors do not believe this is detrmiental to the accuracy of the calculations.
%6.	Was k-point integration used in the calculation? If so how many. If only gamma-point was used, again there needs to be a discussion on the effectiveness of the calculation.

7. In our new simulations, the timestep has been reduced to 0.5 fs. The text has been updated to reflect this.
%7.	The time-step of 1fs appears to be too large for a Car-Parrinello simulation. Since NVE simulations are used, I would be interested in knowing how well the total system energy is conserved (which determines the extent to which the dynamics simulations are accurate).

% 8.	With respect to previous work on molecular computations using graphs, I would like to bring to the author’s attention the following work by O. Sinanoglu, which to my knowledge is one of the first such attempts: Oktay Sinanoglu J. Am. Chem. Soc., 1975, 97 (9), pp 2309–2320. 
8. The authors thank the reviewer for pointing out this previous study using graph theoretical techniques (Oktay Sinanoglu J. Am. Chem. Soc., 1975, 97 (9), pp 2309–2320). That study pertains to synthetic or catalytic pathways of organic systems, but can be generalized to very convoluted and complicated reaction pathways leading to very intricate graphical networks. The systems investigated in that work are much more complicated, and the graphs are designed to convey a great deal more information than in the present text. In our work we wish to investigate only the transitions between different coordination or cyclic-hydrate states, and are not interested in multi-reactant processes. A citation to the work has been added to the text.

### 2012-03-27 resubmission/Article.tex
Preparing the resubmission files. Working on responses to the reviewers.

--- 2011-12-07 submission/Article.tex
@@
 
-Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.\cite{Baer2010} Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 273K.
+Quantum MD techniques are the logical follow-up as they accurately reproduce the hydration geometry around the bound \suldiox~molecules, and allow us to examine in detail the specific bonding interactions that occur within the surface hydrates, and in the extended bonding further into the water.\cite{Baer2010} Two parallel studies are performed in this work; one at 300K, and one at the more atmospherically relevant cold 210K.
 This set of temperatures complements our most recent experimental studies that showed the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures.\cite{Ota2011} Other experiments by our group developed the picture of \suldiox~adsorption, and showed that \suldiox~surface hydrate complexes form when a water surface is exposed to \suldiox~gas.\cite{Tarbuck2005,Tarbuck2006} Although conclusions regarding the specific nature of those complexes could only be inferred from the experiments, our current computational studies now provide us with insights about the specific microscopic geometries and behaviors of the hydrating complexes.
@@
 To study the formation and behavior of cyclic hydrate structures we employ graph theoretical techniques on MD trajectory data.
-Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998}
+Previous use of graphs in molecular computations were applied to finding stable arrangements of water clusters, ice, hydrogen bonding, extracting topological molecular properties, enumerating complex reaction pathways, and cyclic structure studies.\cite{Anick2002, Huber2007, Radhakrishnan1991, Shi2005, Garcia2004, McDonald1998, Sinanogly1975}
 
@@
 The electronic charge density was described using an auxiliary basis set of plane waves.
-Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian DZVP basis set, the exchange-correlation functional of Becke, Lee, Yang, and Parr (BLYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 1 fs was used, with a Nose-Hoover thermostat set at 273K and 300K for the ``cold'' and ``hot'' simulations, respectively.
-These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
+Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian TZV2P basis set with a cutoff of 280 Ry, the exchange-correlation functional of Becke, Lee, Yang, and Parr (B3LYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 0.5 fs was used, with a Nos\'{e}-Hoover thermostat set at 210K and 300K for the ``cold'' and ``hot'' simulations, respectively.
+%These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
 
-Initially, 10 equilibrated boxes of side-lengths 10.0\angs, with 36 randomly packed water molecules were used.
-Five of the boxes were used for each of the cold and hot simulations.
+Initially, 20 equilibrated boxes of side-lengths 8.545\angs, with 24 randomly packed water molecules were used.
+Ten of the boxes were used for each of the cold and hot simulations.
 A sulfur dioxide molecule was randomly placed onto the surface within 2.5\angs~of a water molecule centrally located above the waters in the z-axis.
-A copy of the initial system cubes were then expanded along one axis (z-axis) to 25\angs.
+A copy of the initial system cubes were then expanded along one axis (z-axis) to 30\angs.
 The system energy was minimized through a geometry optimization.
-Subsequently, the system was equilibrated for 1 ns in canonical ensemble (NVT) conditions.
+Subsequently, the system was equilibrated for 1 ps in canonical ensemble (NVT) conditions.
 Periodic boundaries were set on the two short axes to form an infinite slab.
-The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 1 fs.
-The initial 1 ns equilibration trajectory was not included in the final analysis.
-This simulation process resulted in 20,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system, for a total of 100,000 timesteps at each temperature.
+The equilibrated systems were then simulated for a further 20 ps in the microcanonical ensemble (NVE), with trajectory snapshots recorded every 0.5 fs.
+The initial 1 ps equilibration trajectory was not included in the final analysis.
+This simulation process resulted in 40,000 time steps of system trajectory for analysis in each of the hot and cold replicas of the system, for a total of 400,000 timesteps at each temperature.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{so2-coordinations-percents-small.png}
+		\includegraphics[scale=1.0]{coordination-distributions.png}
 		\caption{The \suldiox~bonding coordinations occur with different frequencies at the surface under hot and cold temperature conditions.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{SO-cycle-breakdown-with-cartoons-small.png}
+		\includegraphics[scale=1.0]{cycle-distributions.png}
 		\caption{Cyclic hydrate structures form throughout the MD simulations involving the \suldiox~and one or more water molecules.
@@
 For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript.
-A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively.
+%A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively.
 This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.
 
-In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value.
+%In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value.
 This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing.
@@
 
-We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions.
+%We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions.
 The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states.
@@
 
-\[
-  f(t) = \left\{ 
-  \begin{array}{l l}
-    0 & \quad C_s(t) < 0.2\\
-    1 & \quad C_s(t) \geq 0.2\\
-  \end{array} \right.
-\]
+%\[
+%  f(t) = \left\{ 
+%  \begin{array}{l l}
+%    0 & \quad C_s(t) < 0.2\\
+%    1 & \quad C_s(t) \geq 0.2\\
+%  \end{array} \right.
+%\]
 
-where $f(t)$ is the debounced time-function that represents the lifespans of cycles.
+%where $f(t)$ is the debounced time-function that represents the lifespans of cycles.
 Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).
+
+%\begin{figure}[h!]
+%	\begin{center}
+%		\includegraphics[scale=1.0]{cycle-debouncing-small.png}
+%		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures.
+A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
+A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure.
+A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
+%		\label{fig:debouncing}
+%	\end{center}
+%\end{figure}
+
+The distribution of cycle lifetimes (i.e.
+contiguous spans of time spent with a fully formed cyclic hydrate) was determined.
+The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations.
+The most frequent lifespan for both temperature data sets is less than 50 fs.
+This accounts for the vast majority of cyclic hydrates found (approximately 80\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time.
+The nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than a few fs.
+The difference between hot and cold systems in the $<50$ fs population is less than 3\%, with this trend extending to the longer lifespans as well.
+The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 0.2 ps.
+All of the distribution shows only a $<5$\% difference between cold and hot cyclic lifetimes.
+Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time.
+The cold system has more cycles forming than the hot system at the shortest cycle lifetimes ($<10$ fs).
+Up to 100 fs, the hot temperature cycles show a very slight population increase above the cold temperature cycles.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{cycle-debouncing-small.png}
-		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures.
-A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
-A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure.
-A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
-		\label{fig:debouncing}
+		\includegraphics[scale=1.0]{cycle-lifespans.png}
+		\caption{The lifetime of a cyclic hydrate structure after it forms, and before it distorts in shape due to the movement of the liquid surface molecules, is most often short ($<$ 50 fs).
+Shown here is a plot of cycle hydrate structure lifetimes.
+Both cold (blue) and hot (red) simulations are graphed.
+The inset expands the view of the region above 200 fs.}
+		\label{fig:cycle-lifespans}
 	\end{center}
 \end{figure}
-
-The distribution of cycle lifetimes (i.e.
-contiguous spans of time spent with $f(t) = 1$) was determined.
-The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations.
-The most frequent lifespan for both temperature data sets lasts between 0-1 ps.
-This accounts for the vast majority of cyclic hydrates found (approximately 95\%) indicating that these structures are very transient, and are continuously forming and breaking for very short periods of time.
-Even with the debouncing procedure that would artificially increase the timespan spent either formed or broken, the nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than 1 ps.
-The difference between hot and cold systems in the $<1$ ps population is less than 2\%, with this trend extending to the longer lifespans as well.
-The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 1 ps.
-All of the distribution shows only a $<1.5$\% difference between cold and hot cyclic lifetimes.
-Overall, the distribution shows that when cycles form, at both temperatures, they last a similar amount of time.
-Up to 3 ps, the cold temperature cycles show a very slight population increase above the hot temperature cycles.
-Above 4 ps, most of the cycles that form are found in the hot system.
-The 8 ps cycles are notable in that they form for just under half the length of one of the simulated trajectories.
 
@@
 Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure.
-Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 39\% and 24\% of the bonding coordinations in the cold and hot system, respectively.
-Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 16\% above the cold).
-Of the three most encountered bonding coordinations, the hot system shifts population from the ``S'' (a cycle-breaking coordination) to ``SO'' and ``SOO'' (coordinations that allow for bonding cycle formation through the \suldiox).
-This is relative to the cold system coordination distribution that has equal ``S'' and ``SOO'' populations.
+Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 54\% and 70\% of the bonding coordinations in the cold and hot system, respectively.
+Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 2.5\% above the cold).
+Of the most encountered bonding coordinations the hot system has a much higher proportion of unbound coordinations, and a relative decrease of the ``SOO'' coordination.
 The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures.
 This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{cyclic-lifespans-inset-small.png}
-		\caption{The lifetime of a cyclic hydrate structure after it forms, and before it distorts in shape due to the movement of the liquid surface molecules, is most often short ($<$ 1 ps).
-Shown here is a plot of cycle hydrate structure lifetimes after having applied the debouncing procedure (described in the text).
-Both cold (blue) and hot (red) simulations are graphed.
-The inset expands the view of the region above 1 ps.}
-		\label{fig:cycle-lifespans}
-	\end{center}
-\end{figure}
 

### 2012-03-27 resubmission/Article.tex
more revisions. Need to work on rewriting data analysis

@@
 
-In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on water.
-We simulate a dynamic water surface, complete with all the extended hydrogen-bonding interactions that capture the variability of the \suldiox~hydrate structures, and the behavior of the water surface molecules.
-The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a water slab are incorporated.
+In this study we use ab initio quantum molecular dynamics (MD) techniques to model and simulate the hydrating structures that form around a surface-bound \suldiox~on a water cluster surface.
+We simulate a dynamic water surface of a small cluster, complete with all the extended hydrogen-bonding interactions that capture the variability of the \suldiox~hydrate structures, and the behavior of water's surface molecules.
+The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a larger water cluster are incorporated.
 In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas.
@@
 We also examine \suldiox~binding behavior when bound to the surface waters.
-Lastly, our analysis of a specific bonding arrangement demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to preferentially form an extended cyclic ring structure through intermolecular bonds.
+Lastly, our analysis of a specific bonding arrangement demonstrates an extended bonding structure of \suldiox~hydrates, as they are seen to form an extended cyclic ring structure through intermolecular bonds.
 

### 2012-03-28 resubmission/Article.tex
more of same

@@
 A first visual inspection reveals several trends.
-Clearly the ``SO'' coordination is the most populous at both temperatures.
-The second and third most populated coordinations are the ``S'' and ``SOO'', however their distributions differ between temperatures.
+At both temperatures the \suldiox~spends much of the time unbound without any bonding to the surface waters.
+The cluster of the three most populated coordinations is comprised of the ``S'', ``SO'', and ``SOO''.
+This grouping represents the coordinations with a single bond through the sulfur, and up to two interactions through the \suldiox-oxygens.
+In this coordination group the ``SO'' is the most populous, but in no way dominates the bonding types as the difference in populations is less than 5\% within each temperature.
+The second and third most populated bonding coordinations (excluding the unbound state) are the ``S'' and ``SOO'', however their distributions differ between temperatures.
 In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally.
-The hotter temperature simulation shifts the distribution such that the ``S'' occurs 5\% less frequently than in the cold, and the ``SOO'' occurs nearly 10\% more often.
-The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``SOO'', respectively.
-These results for the hot (room temperature) system coincide with those of the previous single-temperature simulation study by Baer et al.
-at a similar temperature.\cite{Baer2010}
+The hotter temperature simulation shifts the distribution such that the ``S'' occurs approximately 4\% less frequently than in the cold, and the ``SOO'' occurs nearly 8\% less often.
+The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``S'', respectively.
 
@@
 	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from this distribution of coordinations.
-Clearly the cold \suldiox~spends more time than the hot \suldiox, 13\% versus 3\%, respectively, completely unbound from the surface waters.
-This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis.
-Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
+Clearly the hot \suldiox~spends more time than the cold \suldiox, 45\% versus 23\%, respectively, completely unbound from the surface waters.
+This does not necessarily imply a complete desorption into the gas phase, but likely only a brief sojourn away from the surface waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis.
+The unbound coordination data is also supported by our recent spectroscopic experiments that resulted in increased \suldiox~adsorption at colder temperatures.\cite{Ota2011} Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
 Baer et al.
@@
 A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters.
-In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.
+In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, to the extent that a non-sulfur coordination is not often formed during the course of all our simulations.
 
@@
 They concluded that there is asymmetric hydrogen bonding through the \suldiox-oxygens, with one oxygen binding more often than the other.
-This is supported by the findings here where all the double oxygen coordinations (e.g.
-``OO'', ``SOO'', etc.) represent a much lower percentage of the coordinations than the single oxygen counterparts (e.g.
+This is supported by the findings here where the double oxygen coordinations (e.g.
+``OO'', ``SOO'', etc.) represent a lower percentage of the coordinations than the single oxygen counterparts (e.g.
 ``O'', ``SO'', etc.).
@@
 Three \suldiox-oxygen bonds only form if both \suldiox-oxygens are interacting with water hydrogens.
-Our finding that triple-oxygen coordinations rarely form complements Baer et al's conclusion about the asymmetry in the oxygen interactions.
+Our finding that triple-oxygen coordinations rarely form, and that double-oxygen coordination are less frequent than the single-oxygen counterparts, complements Baer et al's conclusion about the asymmetry in the oxygen interactions.
 
@@
 From the data it is clear that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones.
-Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' together represent less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
+Adding a third oxygen atom is also unfavorable as the ``OOO'', ``SOOO'', and ``SSOOO'' together represent less than 3\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
 
 A picture can now be formed of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study.
-The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens.
-The \suldiox-oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.
+Often the \suldiox~will remain unbound from the interface, nearly twice as often when the temperature is hot.
+The \suldiox~will mostly have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens.
+The \suldiox-oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant bonding coordination will be the ``SO'' bonding arrangement.
 
@@
 The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in bonding coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot.
-In the cold temperature, the unbound, ``S'', and ``SO'' coordinations are more populated than in the hot systems.
-The increased temperature decreases the time spent in the unbound coordination, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature populations.
-As shown, the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to coordinations with double-oxygen and double-sulfur bonds.
+In the cold temperature, the bound coordinations are overall more populated than when hot.
+Conversely, the hot temperature system is much more often unbound (nearly half of the simulated time) than when the temperature is cold.
+As shown, the cold \suldiox~spends nearly half as much time unbound as the hot \suldiox, with most of the unbound population in the cold simulations shifting to coordinations with single-sulfur bonds.
 
-On the cold water surface, the ``S'' and ``SO'' are more populated than for the hot system.
+%On the cold water surface, the ``S'' and ``SO'' are more populated than for the hot system.
 The relative decrease of these coordinations are matched in the hot surface by an increase of the ``SOO'' configuration.
@@
 the number of times each \suldiox~switched from one coordination type to another).
-With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} A and B, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one bonding coordination to another.
+With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} left and right, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one bonding coordination to another.
 The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently).
@@
 
-As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} than transitions to less populated coordinations.
+As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} (left) than transitions to less populated coordinations.
 Insights to the bonding process are made clearer from these graphs.
-In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes.
-The number of transitions within this ``S-SO-SOO'' group of bonding coordinations is an order of magnitude larger than any other transition.
-This indicates that while the \suldiox~is bound in any of the three most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other two coordinations in this group.
+In the cold system graph of Figure \ref{fig:coordination-transitions}, we find that the majority of transitions are between the unbound and ``O'' nodes, and between the single-sulfur coordinations (e.g.
+``S-SO'', ``SO-SOO'', etc.) The number of transitions within this group of bonding coordinations is an order of magnitude larger than any other transition.
+This indicates that while the \suldiox~is bound in any of the most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other coordinations.
 Clearly, the \suldiox~is rather active and constantly forming and breaking bonds through its oxygens.
 
-In the graph of the hot system in Figure \ref{fig:coordination-transitions}B, the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size.
+In the graph of the hot system in Figure \ref{fig:coordination-transitions} (right), the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size.
 One very surprising result is in the transition from ``SOO-SOOO''.

### 2012-03-28 resubmission/ResponseToReviews.tex
Finishing up with the resubmission draft, and the responses to the reviewers. Added plots of the temp and conserved quantities for two representative plots at cold/hot to submit with the response to reviews.

@@
+\documentclass{article}
+
+\usepackage {textcomp}
+\usepackage {setspace}
+\usepackage [pdftex]{graphicx}
+
+% some formatting tags
+\usepackage[
+top    = 0.5in,
+bottom = 1.5in,
+left   = 2.50cm,
+right  = 2.50cm]{geometry}
+
+\begin{document}
+
+%\doublespacing
 Journal: The Journal of Physical Chemistry
+
 Manuscript ID: jp-2011-11804m
+
 Title: "Staying Hydrated: The Molecular Journey of Gaseous Sulfur Dioxide to a
+
 Water Surface"
+
 Author(s): Shamay, Eric; Richmond, Geraldine 
+\\ \\
+Dear Prof. Jungwirth,
 
-Dear Prof. Richmond:
+Because of the depth and impact of the reviewer comments about our recently submitted manuscript, the authors have decided to completely re-run the simulations of SO2 on water in order to bring the simulation parameters into line with the reviewer suggestions.
+This major reworking of the simulations and data resulted in major changes to the data presented, manuscript text, results and conclusions.
+We have put forth great effort to respond to each of the reviewer comments.
+Please find our enumerated responses below, matching to each numbered point of the original reviewer comments.
+We believe that with the changes to the work and the newly run simulations, the new manuscript is more accurate and in line with current simulation standards for aqueous systems.
+We hope you will further consider our manuscript for publication in the journal.
+\\ \\
 
-Thank you for submitting your manuscript for publication in The Journal of Physical Chemistry.
-It has been examined by expert reviewers who have concluded that the work is of potential interest to the readership of The Journal of Physical Chemistry; however, it appears that a major revision, possibly followed by further reviewer evaluation, will be needed prior to its further consideration for publication.
-Please see the enclosed reviewers' reports for details regarding the requested changes and/or additions.
+--Eric Shamay \\ \\
 
-Additionally, we have noticed that your manuscript referenced above is missing the Table of Contents graphic.
-Please see the submission requirements for the Journal of Physical Chemistry TOC Graphic at http://pubs.acs.org/paragonplus/submission/jpchax/jpchax_authguide.pdf.
-Note that the graphic for the TOC entry should be in the form of a structure, graphical image, photograph, reaction scheme, or equation.
-The Author must submit the graphic in the actual size to be used for the TOC, fitting in an area 1.375 in.
-high and 3.5 in.
-wide (3.6 cm Ã— 8.9 cm).
-The allowed file formats for TOC graphics are TIFF and PDF.
-Please submit a revision of this manuscript including the TOC graphic as the final page of your Article manuscript revision submission.
-Separate TOC graphics are not allowed.
-
-We ask that you submit your revision no later than 29-Apr-2012.
-
-When submitting your revised manuscript through ACS Paragon Plus, you will be able to respond to the reviewers' comments in the text box provided or by attaching a file containing your detailed responses outlining the changes made and explaining your reasons for disagreeing with any suggestions you choose not to follow.
-
-To Revise Your Manuscript on the Web:
-
-To revise your manuscript, log into ACS Paragon Plus at http://paragonplus.acs.org/login and select "My Authoring Activity".
-There you will find your manuscript title listed under "Revisions Requested by Editorial Office." Your original files are available to you when you upload your revised manuscript.
-If you are replacing files, please remove the old version of the file from the manuscript before uploading the new file.
-
-Thank you for considering The Journal of Physical Chemistry as a forum for the publication of your work.
-
-With sincere regards,
-
-Prof. Pavel Jungwirth 
-Senior Editor 
-The Journal of Physical Chemistry 
-Journal Fax: 202-354-5425 
-Journal Email: jungwirth-office@jpc.acs.org
-
-Reviewer: 1 
------------
+\textbf{Reviewer: 1} \\
 
@@
 The authors have taken into consideration the entirety of all the reviewer suggestions regarding functional and basis sets, cutoffs, timesteps, and temperature that affect the simulations.
-The changes are reflected in the text, and meet the standards suggested in the reviews.
-All the simulation data has been re-analyzed, and the results are all presented from the new simulations.
+The changes to the simulation parameters are reflected in the text, and meet the standards suggested in the reviews.
+All the simulations have been rerun, data has been re-analyzed, and the results are all presented from the new simulations.
+\\
 % 1.
@@
 
-%4a. The cold system temperature was set to 210K in the resimulated systems. This much larger separation of temperatures is more in line with the reviewer's concerns regarding thermostatting. 
 4.
+The cold system temperature was set to 210K in the resimulated systems.
+This much larger separation of temperatures is more in line with the reviewer's concerns regarding temperature separation.
+Furthermore, included with this revised submission are plots of the temperature and the conserved quantity for each timestep for both low and high temperature systems.
+As is evident, the temperature increases until reaching the setpoint temperature soon after the start of the simulations.
+The plateau region of the temperature extends for the majority of the simulations.
+Lastly, it is clear that the separation of temperatures persists for the entire simulation time, confirming that the systems are representing temperatures that are distinct.
+The authors believe that this is evidence that the parameter set chosen for simulation of these small water systems are sufficient to recreate the expected temperature conditions, even when simulating in the NVE ensemble.
+\\
+%4.
 Following recent literatures on simulations of liquid water using potentials derived from density functional theory (albeit some with different codes as well as different methods), one firm conclusions can be made is that water near ambient conditions are more glassy than liquid.
@@
 
-5. “Nose” should be “Nosé”. In addition, the coupling frequency, when not chosen appropriately can cause energy drifts, was not specified.
+5. The temperature coupling constant was added to the text, and the typo was revised.
+\\
+%5. “Nose” should be “Nosé”. In addition, the coupling frequency, when not chosen appropriately can cause energy drifts, was not specified.
 
-6. The text has been updated to reflect the equilibration period of 1 ps.
+6. The text has been updated to reflect the correct equilibration period of 1 ps.
+\\
 %6. The authors indicate the system was equilibrated for “1 ns”. This just does not seem correct unless the equilibration was done using a cheaper potential other than DFT.
@@
 7.
-From the RDF, it can be seen that both hot and cold simulations converges to 1.0 as distance increases.
+Both RDFs converge to 1.0 because of the normalization routine employed in the calculations.
+This is typical of RDF calculations - normalized to unity - and the authors prefer this normalization as it aids in comparison to the other RDFs published for similar systems (see Baer et al.
+cited in the text)
+	\\
+%7.
+From the RDF, it can be seen that both hot and cold simulations converge to 1.0 as distance increases.
 But if the sulfur dioxide gas molecule is at the liquid-vapor interface, shouldnʼt this value approach something more like 0.5 instead of 1.0?
@@
 The goal of the project was to monitor specific bonding behaviors of SO2 using previously established methods, not as a study of the results from various constraints or changes to simulation parameters.
+The authors believe that additional discussion on this topic would distract from the narrative presented.
+\\
 %8. Has the authors explore the ramifications of including dispersion correction in their calculation which might play a large in such a system?
 
-Reviewer: 2 
------------
-
-Recommendation: This paper is probably publishable, but should be reviewed again in revised form before it is accepted.
-
-Reviewer: 3
------------
+\textbf{Reviewer: 3} \\
 
 1.
-Because the re-simulated systems were done using B3LYP functional a benchmark comparison is no longer necessary.
+Because the re-simulated systems were done using B3LYP functional, the authors believe a benchmark comparison is no longer necessary.
 Small cluster studies of so2 with several waters have been performed using high-accuracy basis and functionals previously, and we do not set out to reproduce that work here, but rather to build on it.
 The authors believe that the current work with the new simulation parameters achieves a level of accuracy suggested by the reviewers, but that lengthy discussion of this fact is not requisite in this particular simulation study.
+\\
 %1.	The predominant stabilization interaction between and SO2 molecule and a water molecule is due to dipoles.
@@
 
-2.	The larger concentration of unbound SO2 at a lower temperature is counter-intuitive.
+2.
+After resimulating the systems the resulting relative populations of unbound SO2 has changed.
+The more intuitive stronger binding of SO2 at the cold temperature results from the higher accuracy of the functional and basis sets used in the new simulations.
+The new results are also more in line with the author's experimental results from the previous publications cited in the text.
+\\
+%2.
+The larger concentration of unbound SO2 at a lower temperature is counter-intuitive.
 (Fig.
@@
 Thus, the authors have made the conscious decision to avoid lengthy discussion of the surrounding water's structure.
+\\
 %3.	The structure of the surrounding water molecules is never discussed.
 
-4.	It is stated that Quickstep is used. Assume this implies that the Car-Parrinello method is used, but it is not stated in the paper.
+4. Reference to the Car-Parrinello method has been added to the text.
+\\
+%4.	It is stated that Quickstep is used. Assume this implies that the Car-Parrinello method is used, but it is not stated in the paper.
 
-5.	The system includes 36 water molecules in a box. The effect of this restriction is important to discuss.
+5.
+The new systems were simulated with a smaller number of water molecules and are referred to as a small water cluster.
+The effects of using few water are not discussed in detail as the number of waters used is representative of many small-cluster MD and DFT studies.
+The authors believe that the number of waters is in line with several previous studies of molecules bound to small water clusters, and further discussion is not warranted as part of the narrative being presented.
+\\
+%5.	The system includes 36 water molecules in a box. The effect of this restriction is important to discuss.
 
@@
 Also, the CP2K simulation package does not allow for this type of calculation.
-The authors do not believe this is detrmiental to the accuracy of the calculations.
+The authors do not believe this is detrmiental to the accuracy of the calculations and further discussion is not pertinent to this study.
+\\
 %6.	Was k-point integration used in the calculation? If so how many. If only gamma-point was used, again there needs to be a discussion on the effectiveness of the calculation.
 
-7. In our new simulations, the timestep has been reduced to 0.5 fs. The text has been updated to reflect this.
+7. In the new simulations, the timestep has been reduced to 0.5 fs. The text has been updated to reflect this.
+\\
 %7.	The time-step of 1fs appears to be too large for a Car-Parrinello simulation.
@@
 
-% 8.	With respect to previous work on molecular computations using graphs, I would like to bring to the author’s attention the following work by O.
-Sinanoglu, which to my knowledge is one of the first such attempts: Oktay Sinanoglu J.
-Am.
-Chem.
-Soc., 1975, 97 (9), pp 2309–2320.
 8.
@@
 A citation to the work has been added to the text.
+\\
+% 8.	With respect to previous work on molecular computations using graphs, I would like to bring to the author’s attention the following work by O.
+Sinanoglu, which to my knowledge is one of the first such attempts: Oktay Sinanoglu J.
+Am.
+Chem.
+Soc., 1975, 97 (9), pp 2309–2320.
+
+\end{document}

### 2012-03-28 resubmission/Article.tex
New RDF images
Finishing up with the resubmission draft, and the responses to the reviewers. Added plots of the temp and conserved quantities for two representative plots at cold/hot to submit with the response to reviews.

@@
 
-To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A graph search is a recursive algorithm of queuing nodes and all neighboring nodes while performing a specified procedure on each visited node.
+To detect cyclic structures in a graph a depth-first or breadth-first search (DFS and BFS, respectively) may be used.\cite{Knuth1997, Cormen 2001} A graph search is a recursive algorithm of queueing nodes and all neighboring nodes while performing a specified procedure on each visited node.
 This is easily performed on adjacency list or connectivity matrix data structures, iterating through nodes (i.e.
@@
 In graph search terminology, all nodes are colored during graph traversal to distinguish unvisited nodes (white), queued nodes (gray), and visited nodes (black).
-Performing a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queuing adjacent neighbors of a node.
+Performing a BFS on a graph, cyclic structures are detected any time a ``gray target'' is encountered when queueing adjacent neighbors of a node.
 A benefit of BFS is the ability to determine the smallest cyclic structure containing a given node.
@@
 Type \Rmnum{3} is formed by a \suldiox~with the ``SO'' bonding coordination, which is the most dominant bonding coordination encountered.
-Each hydrate structure is labeled with its corresponding bonding coordination type using the nomenclature described in the text.}
+Each hydrate structure is labelled with its corresponding bonding coordination type using the nomenclature described in the text.}
 		\label{fig:cyclic-structures}
@@
 
-On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package, which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions.
+On-the-fly ab initio molecular dynamics simulations were performed with the QUICKSTEP package (Car-Parrinello method of MD), which is an implementation of the Gaussian plane wave method using the Kohn-Sham formulation of density functional theory (DFT).\cite{VandeVondele2005} The Kohn-Sham orbitals are expanded using a linear combination of atom-centered Gaussian-type orbital functions.
 The electronic charge density was described using an auxiliary basis set of plane waves.
-Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian TZV2P basis set with a cutoff of 280 Ry, the exchange-correlation functional of Becke, Lee, Yang, and Parr (B3LYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 0.5 fs was used, with a Nos\'{e}-Hoover thermostat set at 210K and 300K for the ``cold'' and ``hot'' simulations, respectively.
+Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian TZV2P basis set with a cutoff of 280 Ry, the exchange-correlation functional of Becke, Lee, Yang, and Parr (B3LYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 0.5 fs was used, with a Nos\'{e}-Hoover thermostat set at 210K and 300K for the ``cold'' and ``hot'' simulations, respectively, using a coupling constant of 1000 fs.
 %These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
 
-Initially, 20 equilibrated boxes of side-lengths 8.545\angs, with 24 randomly packed water molecules were used.
+Initially, 20 equilibrated boxes of side-lengths 8.9545\angs, with 24 randomly packed water molecules were used.
 Ten of the boxes were used for each of the cold and hot simulations.
@@
 At both temperatures the \suldiox~spends much of the time unbound without any bonding to the surface waters.
-The cluster of the three most populated coordinations is comprised of the ``S'', ``SO'', and ``SOO''.
+The cluster of the three most populated coordinations is comprised of the ``S'', ``SO'', and ``SOO'' bonding types.
 This grouping represents the coordinations with a single bond through the sulfur, and up to two interactions through the \suldiox-oxygens.
-In this coordination group the ``SO'' is the most populous, but in no way dominates the bonding types as the difference in populations is less than 5\% within each temperature.
+In this coordination group the ``SO'' is the most populous, but in no way dominates the bonding types as the difference in populations is less than 2\% and 5\% within the cold and hot simulations, respectively.
 The second and third most populated bonding coordinations (excluding the unbound state) are the ``S'' and ``SOO'', however their distributions differ between temperatures.
-In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally.
-The hotter temperature simulation shifts the distribution such that the ``S'' occurs approximately 4\% less frequently than in the cold, and the ``SOO'' occurs nearly 8\% less often.
-The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``S'', respectively.
+In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally (18.8\% and 18.1\%, respectively).
+The hotter temperature simulation shifts the distribution such that the ``S'' occurs 4.6\% less frequently than in the cold, and the ``SOO'' occurs 8.6\% less often.
+The distribution of bonded coordinations (those excluding the ``unbound'' state) in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``S'', respectively.
 
@@
 	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from this distribution of coordinations.
-Clearly the hot \suldiox~spends more time than the cold \suldiox, 45\% versus 23\%, respectively, completely unbound from the surface waters.
+Clearly the hot \suldiox~spends more time than the cold \suldiox, 45.8\% versus 22.8\%, respectively, completely unbound from the surface waters.
 This does not necessarily imply a complete desorption into the gas phase, but likely only a brief sojourn away from the surface waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis.
-The unbound coordination data is also supported by our recent spectroscopic experiments that resulted in increased \suldiox~adsorption at colder temperatures.\cite{Ota2011} Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
+The unbound coordination data is also supported by our recent spectroscopic experiments that resulted in increased \suldiox~surface binding at colder temperatures.\cite{Ota2011} Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures.
 Baer et al.
@@
 	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (e.g.
-``O'', ``OO'', etc.) represent the least frequently formed.
+``O'', ``OO'', etc.) form infrequently (approximately 6\% and 8\% of the total hot and cold bonding coordinations, respectively).
 A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters.
@@
 % Baer's assymetric oxygen binding
-Baer et al performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens.
+Baer et al.
+performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens.
 They concluded that there is asymmetric hydrogen bonding through the \suldiox-oxygens, with one oxygen binding more often than the other.
@@
 From the data it is clear that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones.
-Adding a third oxygen atom is also unfavorable as the ``OOO'', ``SOOO'', and ``SSOOO'' together represent less than 3\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
+Adding a third oxygen atom is also unfavorable as the ``OOO'', ``SOOO'', and ``SSOOO'' combined represent less than 4\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.
 
 A picture can now be formed of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study.
-Often the \suldiox~will remain unbound from the interface, nearly twice as often when the temperature is hot.
-The \suldiox~will mostly have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens.
+Often the \suldiox~will remain unbound from the interface, more than twice as often when the temperature is hot.
+The \suldiox~will form at least one interaction with neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens.
 The \suldiox-oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant bonding coordination will be the ``SO'' bonding arrangement.
@@
 Conversely, the hot temperature system is much more often unbound (nearly half of the simulated time) than when the temperature is cold.
-As shown, the cold \suldiox~spends nearly half as much time unbound as the hot \suldiox, with most of the unbound population in the cold simulations shifting to coordinations with single-sulfur bonds.
+A decrease in temperature causes a clear change to the surface binding behavior of \suldiox.
+At colder temperatures a \suldiox~molecule will bind more strongly to a water surface than at higher temperatures.
 
@@
 Figure \ref{fig:rdf} shows the radial distribution functions (RDF) of \suldiox~to water atoms for both cold (blue) and hot (red) temperatures.
-The S-O$_{H_2O}$ RDFs are nearly equal except for a slightly taller first peak in the cold system.
-Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF indicates that since bonding occurs frequently through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface waters.
-In the hot systems, the bonding coordinations with two oxygen bonds (e.g.
-``SOO'', etc.) occur more frequently than in the cold system.
-This additional bonding through a second oxygen interaction may slightly shift neighboring hydrating waters away from the \suldiox-sulfur towards the oxygen end of the molecule.
-This suggests that the cold \suldiox~bonds closer to the surface waters through its sulfur atom (as shown in the S-O$_{H_2O}$ RDF) and favors the more ``sulfur-centric'' bonding coordinations (e.g.
-``S'', ``SO'').
-The increased temperature of the hot system allows the \suldiox~to bond more extensively through its oxygens to the ``SOO'' coordination.
-The greater interactions through the oxygens, and higher bonding coordination, may pull the \suldiox~further into the water interface and then allow for increased bonding through the sulfur, up to double-sulfur coordinations (e.g.
-``SS'', ``SSO'', etc.).
+The S-O$_{H_2O}$ RDFs are overall very similar.
+A difference between the cold and hot simulations appears in the first peak of both RDFs.
+The cold RDFs have a slightly taller and narrower first peak than the hot counterpart, and the peak is located at a slightly shorter distance.
+Along with the slightly larger population of the cold ``S'' and ``SO'' coordinations in the cold surface, the RDF indicates that since bonding occurs frequently through the sulfur, the cold \suldiox-sulfur interacts more closely with the surface water oxygens than in the hot waters.
+This suggests that the cold \suldiox~favors the more ``sulfur-centric'' bonding coordinations (e.g.
+``S'', ``SO'', ``SOO'').
+The broader first peak of the hot RDFs likely indicates that the sulfur atoms have a slightly larger array of bonding environments to water oxygens.
+Similarly, the \suldiox-oxygen displays the same difference in the first peak on the left plot of Figure \ref{fig:rdf}.
 
@@
 		\includegraphics[scale=1.0]{rdf-small.png}
-		\caption{Radial distribution functions (RDF) show the subtle change in how the hydrating waters around \suldiox~reposition as the temperature is increased from cold (blue) to hot (green).
-Shown are the two correlations between sulfur and water oxygens (left) and the \suldiox-oxygen and water hydrogens (right).}
+		\caption{Radial distribution functions (RDF) show the subtle change in how the hydrating waters around \suldiox~reposition as the temperature is increased from cold (blue) to hot (red).
+Shown are the two correlations between sulfur and water oxygens (right) and the \suldiox-oxygen and water hydrogens (left).}
 		\label{fig:rdf}
@@
 Thicker lines correspond to a greater number of transitions between coordinations.
-The number of times a given transition occurred is labeled to the right of the corresponding line.}
+The number of times a given transition occurred is labelled to the right of the corresponding line.}
 		\label{fig:coordination-transitions}
@@
 
-As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} (left) than transitions to less populated coordinations.
+As expected for the more populated coordinations, there are generally more transitions between larger nodes in Figure \ref{fig:coordination-transitions} than transitions to less populated coordinations.
 Insights to the bonding process are made clearer from these graphs.
 In the cold system graph of Figure \ref{fig:coordination-transitions}, we find that the majority of transitions are between the unbound and ``O'' nodes, and between the single-sulfur coordinations (e.g.
-``S-SO'', ``SO-SOO'', etc.) The number of transitions within this group of bonding coordinations is an order of magnitude larger than any other transition.
-This indicates that while the \suldiox~is bound in any of the most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other coordinations.
-Clearly, the \suldiox~is rather active and constantly forming and breaking bonds through its oxygens.
-
+``S-SO'', ``SO-SOO'', etc.) The number of transitions within these groups of bonding coordinations is an order of magnitude larger than any other transition.
+This indicates that while the \suldiox~is bound in any of the most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other coordinations, typically those with a single sulfur interaction.
+Clearly, the \suldiox-sulfur bond to water oxygens is stronger than its oxygen bonds to water-hydrogens.
+The \suldiox-oxygens are actively and constantly forming and breaking bonds to neighboring waters.
 In the graph of the hot system in Figure \ref{fig:coordination-transitions} (right), the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size.
-One very surprising result is in the transition from ``SOO-SOOO''.
+%One very surprising result is in the transition from ``SOO-SOOO''.
 This transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations.
@@
 From the unbound coordination, the \suldiox~can bind to waters either through the sulfur or an oxygen to enter the ``S'' or ``O'' coordinations, respectively.
-At both temperatures we find that the coordination transition in Figure \ref{fig:coordination-transitions} from unbound to ``O'' occurs almost three times more than the transition from unbound to ``S''.
+At both temperatures we find that the coordination transition in Figure \ref{fig:coordination-transitions} from unbound to ``O'' occurs over three times more than the transition from unbound to ``S''.
 We find two possibilities that may explain this difference.
@@
 Otherwise, the ``O'' coordination may be the first step in adsorption of an unbound \suldiox, where a subsequent addition of an \suldiox-sulfur interaction to a water oxygen would lead to a transition to the most frequent coordination, ``SO''.
-In the latter case, any adsorption of \suldiox~proceeds through an oxygen binding, accounting for the increased unbound-``O'' transition frequency.
+In the latter case, any adsorption of \suldiox~proceeds through an oxygen binding, accounting for the increased unbound-``O'' transition frequency, but because it is likely a less-stable coordination, its population is less than the equivalent ``S'' or ``SO'' configuration.
 
@@
 Both cold (blue) and hot (red) data are plotted.
-The data of Figure \ref{fig:coordination-lifespans} show that most coordination configurations last a very brief time, with the majority forming for under 0.5 ps.
-The three most populous coordinations, ``S'', ``SO'', ``SOO'' (as determined from Figure \ref{fig:bonding-coordinations} by percentage) in both temperatures have coordination lifetimes of up to 1 ps, in some instances lasting up to 1.5 ps.
+The data of Figure \ref{fig:coordination-lifespans} show that most coordination configurations last a very brief time, with the majority forming for under 1 ps.
+The three most populous coordinations, ``S'', ``SO'', ``SOO'' (as determined from Figure \ref{fig:bonding-coordinations} by percentage) in both temperatures have coordination lifetimes of just over 2 ps, in some instances lasting up to 5-6 ps.
 The brevity of lifespans overall speaks to the dynamic nature of the \suldiox~surface binding.
@@
 The ``unbound'' configuration stands out as an anomaly amongst the lifespans of the other bonding coordinations.
-The few lifespans above 1.5 ps, up to 3.25 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that it also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''.
+The few lifespans above 3 ps, up to 10 ps long, suggest a \suldiox~that not only unbinds from the water surface, but that it also recedes far enough to avoid a quick rebinding and coordination change to ``S'' or ``O''.
 Those data points of a few long lived unbound species indicate times when the \suldiox~is far from the water, residing in the gas phase until the necessary water rearrangement occurs and it is drawn back to the surface to rebind.
@@
 Returning to the transition plots of Figure \ref{fig:coordination-transitions}, the behavior of the unbound transition to both ``S'' and ``O'' coordination can now be better characterized.
-Figure \ref{fig:coordination-lifespans} shows that the ``O'' coordinated lifetimes are shorter than the ``S'' coordinated ones.
+Figure \ref{fig:coordination-lifespans} shows that the ``O'' coordinated lifetimes are slightly shorter than the ``S'' coordinated ones.
 Figure \ref{fig:coordination-transitions} shows that the unbound-``O'' transition occurs almost three times more than the transition to the ``S'' bonding coordination.
@@
 This may be because the ``S'' coordination is more stable than the ``O''.
-Once in the ``S'' coordination the \suldiox~does not quickly break the interaction from its sulfur to water, but rather remains for up to 1 ps in the ``S'' coordination before (most likely) forming an oxygen bond to make the ``SO'' coordination.
+Once in the ``S'' coordination the \suldiox~does not quickly break the interaction from its sulfur to water, but rather remains for up to 2 ps in the ``S'' coordination before (most likely) forming an oxygen bond to make the ``SO'' coordination.
 We have outlined the likely behavior of \suldiox~as it transitions from the gas phase in an unbound coordination to binding to an aqueous interface by forming bonds and interactions with surface waters.
@@
 The plot in Figure \ref{fig:cyclic-breakdown} shows the distribution of how often the various types of cyclic hydrates were encountered at both cold (blue) and hot (red) temperatures.
-Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4).
-The two tallest data points, left-most in the plot, show the overall time spent in all types of cyclic structures.
+Each data point shows a percentage of the MD trajectories in which the \suldiox~was a member of a cyclic structure, for different numbers of cyclic waters (up to 4, ``Quadruple'').
+The two tallest data points labelled ``Cyclic'', left-most in the plot, show the overall time spent in all types of cyclic structures.
 Clearly the hot \suldiox~spends more time in a cyclic structure than at the cold temperature.
-It is remarkable that the time spent in a cyclic structure shows a 15\% difference (42\% cold, 57\% hot).
-The hot \suldiox~spends well over half of the simulated time bound as one of the hydrate cycles, and the cold \suldiox~spends just under half of the time as such.
+The time spent in a cyclic structure shows a 2.5\% difference (15.5\% cold, 18.0\% hot) between the temperatures.
 Thus, in addition to having earlier found the most likely bonding coordination during the simulated life of \suldiox, we have also found that the hydrates of the \suldiox~form cyclic structures for much of the time while bound to the water surface.
@@
 Now we look at the different types of cyclic hydrates, distinguished by the number of waters involved in the bonding structure.
-In Figure \ref{fig:cyclic-breakdown} the single and double water cycles are the least frequently encountered structures, accounting for less than 10\% of both cold and hot temperature simulation times.
+In Figure \ref{fig:cyclic-breakdown} the single water cycles are the least frequently encountered structures, accounting for less than 1\% of both cold and hot temperature simulation times.
 Formation of the single-water type is likely energetically unfavorable because of the proximity of the water to the \suldiox~required to form the bonding cycle.
 The double-water structure was one of two types of clusters proposed in a previous computational work as a candidate structure contributing to the overall IR spectrum of surface-bound \suldiox.\cite{Baer2010} In those static and geometry-optimized cluster calculations, lacking the extended water structure or bonding from waters external to the hydrate, both the double and triple types appear equally likely to form.
-However, the MD simulations here have introduced many waters into a dynamic environment allowing for extended bonding networks, and the results show clearly that the double-water cyclic hydrate is formed much less often (less than 5\% at both temperatures) than the triple-water form.
+However, the MD simulations here have introduced many waters into a dynamic environment allowing for extended bonding networks, and the results show clearly that the double-water cyclic hydrate is formed much less often (less than 3\% at both temperatures) than the triple-water form.
 
-The results for triple and quadruple-water structures show that larger hydrate cycles are favored at higher temperatures.
-Although the higher number of hydrating waters ($>$ 4) are not shown, those contribute minimally to the overall distribution.
+The quadruple cycle types are similar in nature to the double type cycles.
+Likely the quadruple water cycles are not as favorable as the triple type because the size and number of bonds involved increases the possibility that some part of the cycle will break as another forms.
+Coordinating four waters to meet the cyclic hydrate criteria define here is likely what keeps the four-water cycles from forming often.
+The higher number of hydrating waters ($>$ 4) are not shown as those contribute minimally to the overall distribution.
 The majority of the cyclic hydrates are formed with three waters in the triple-water type.
-This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modeled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that, of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}).
+This hydrate type matches the bonding structure inferred from our previous experiments, and also one of the cluster types modelled by others.\cite{Tarbuck2005,Tarbuck2006,Baer2010} It was further found that, of the triple-type hydrate cycles, the waters contributing to the cycles can be arranged in two ways that preserve the hydrogen-bonding between the molecules (described earlier and shown in Figure \ref{fig:type-3-varieties}).
 The triple-water type cycle results were broken-down into contributions from the type-A and type-B triple-water structures.
 The two triple cycle structures are depicted on the right side of Figure \ref{fig:cyclic-breakdown}, along with the plots of their contributions to the overall distribution.
-It is notable that each temperature has a different dominant type of triple-water cyclic structure.
-The cold system forms more of the type-B, and the hotter system forms primarily type-A.
+It is notable that at both temperatures the dominant type of triple-water cyclic structure is in the ``Triple-A'' form.
+This form of the triple cycle corresponds to the resulting geometry from the calculations of Baer et al.
+and involved an equal contribution from each water in the cycle.\cite{Baer2010}
 
@@
 To answer the question of cyclic lifespan, a method was devised to define the lifetime of a cycle.
-For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript.
+For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript, and the lifespan data was collected as was done for the bonding coordinations.
 %A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively.
@@
 contiguous spans of time spent with a fully formed cyclic hydrate) was determined.
-The distribution of cyclic lifespans is shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations.
+The lifespans are shown in Figure \ref{fig:cycle-lifespans} for the cold (blue) and hot (red) simulations.
 The most frequent lifespan for both temperature data sets is less than 50 fs.
@@
 The nature of the water surface, and the very dynamic extended hydrogen bonding network, keeps many of the structures from lasting much longer than a few fs.
-The difference between hot and cold systems in the $<50$ fs population is less than 3\%, with this trend extending to the longer lifespans as well.
+There is little difference between hot and cold systems in the $<50$ fs population, with this trend extending to the longer lifespans as well.
 The inset of figure \ref{fig:cycle-lifespans} shows an expanded view of the region above 0.2 ps.
@@
 Because the majority of cyclic structures last only briefly ($<1$ ps), the active switching of \suldiox~bonding coordinations appears to break the cyclic structure.
-Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 54\% and 70\% of the bonding coordinations in the cold and hot system, respectively.
-Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (approximately 2.5\% above the cold).
+Figure \ref{fig:bonding-coordinations} shows that the sum of coordinations that necessarily break cyclic structures account for approximately 51\% and 67\% of the bonding coordinations in the cold and hot system, respectively.
+Consequently, the distribution of Figure \ref{fig:cyclic-breakdown} also indicates that there are more cycles formed in the hot system (nearly 2.5\% above the cold).
 Of the most encountered bonding coordinations the hot system has a much higher proportion of unbound coordinations, and a relative decrease of the ``SOO'' coordination.
-The increased temperature appears to cause the \suldiox~to bond in a way that is more conducive to the formation of cyclic structures.
-This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
+The increased temperature appears to cause the \suldiox~to bond in a way that is slightly more conducive to the formation of cyclic structures.
+%This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
 
@@
 
-This study is one of very few temperature studies looking at the microscopic nature of interfacial gas molecules on water.
-By changing the temperature, it was found that a hotter water system leads to longer \suldiox~binding to the water surface.
+This is one of very few temperature studies looking at the nature of microscopic behaviors of interfacial gas molecules on water.
+By changing the temperature, it was found that a colder water system leads to greater \suldiox~binding to the water surface.
 The distribution of bonding coordinations was greatly affected by a temperature change, shifting populations of bonding configurations because of the altered \suldiox~and \wat~behavior.
-At the higher temperature, \suldiox~forms more frequent bonds to interfacial waters through the sulfur and oxygen atoms.
+At the lower temperature, \suldiox~forms more frequent bonds to interfacial waters through the sulfur and oxygen atoms.
 Overall, we have determined that the \suldiox~hydrate interactions are transient, binding and unbinding to water molecules rapidly in very dynamic bonding coordinations.
@@
 The formation of cyclic hydrate structures was probed and it was found that these hydrated bonding ring species form during much of a simulated trajectory.
-Temperature increases the occurrence of cyclic structures, and also shifts the distribution of the specific types of cycles being formed.
+Higher temperatures slightly increase the occurrence of cyclic structures, and also shift the distribution of the specific types of cycles being formed.
 Two types of cyclic tri-hydrates were discovered during the course of simulations.
-The cycle lifetimes were found to be mostly short-lived, with a majority lasting less than 1 ps before breaking and reforming due to the dynamic bonding and motion of the surface waters and \suldiox~molecules.
-Temperature did not have a very dramatic effect on the cyclic lifespans, but higher temperatures did lead to \suldiox~bonding coordinations that are more likely to form into cyclic hydrates.
+The cycle lifetimes were found to be mostly short-lived, with a majority lasting only less than 50 fs before breaking and reforming due to the dynamic bonding and motion of the surface waters and \suldiox~molecules.
+Temperature did not have a very dramatic effect on the cyclic lifespans, but lower temperatures did lead to \suldiox~bonding coordinations that are more likely to form into cyclic hydrates.
 

### 2012-03-29 resubmission/Article.tex
Finalized changes and revisions for resubmission. Only need to make a TOC graphic.

@@
 We simulate a dynamic water surface of a small cluster, complete with all the extended hydrogen-bonding interactions that capture the variability of the \suldiox~hydrate structures, and the behavior of water's surface molecules.
-The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to small cluster DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a larger water cluster are incorporated.
+The quantum MD technique described herein allows more accurate and realistic simulation than our previous classical MD.\cite{Shamay2011} It is also superior to static DFT studies because it does not assume geometry optimized configurations, and the extended interactions of a larger water cluster are incorporated.
 In our previous classical MD study we determined net orientational behavior of \suldiox~binding to a water surface, and the orientation of the waters as they respond to the presence of an adsorbing gas.
 Understanding the orientational behavior of molecules in the aqueous interfacial region during adsorption was a necessary first step to understanding the specific details of gas-binding and surface behavior.
-
-%Unlike small cluster studies that use DFT calculations on geometry optimized structures to find likely geometries in a static vacuum setting, The quantum MD technique allows us to simulate the process much more realistically, and with greater accuracy than our previous classical MD work could allow.\cite{Shamay2011} This computational study greatly enhances the picture developed in our previous work on \suldiox.
 
@@
 Energies and forces from on-the-fly simulation sampling of the Born-Oppenheimer surface were calculated for each MD step using the Gaussian TZV2P basis set with a cutoff of 280 Ry, the exchange-correlation functional of Becke, Lee, Yang, and Parr (B3LYP),\cite{LEE1988} and the atomic pseudo-potentials of the Goedecker, Teter, and Hutter type.\cite{Goedecker1996} A simulation timestep of 0.5 fs was used, with a Nos\'{e}-Hoover thermostat set at 210K and 300K for the ``cold'' and ``hot'' simulations, respectively, using a coupling constant of 1000 fs.
-%These computational parameters were verified to yield a reasonable description of bulk room temperature water when simulating a neat-water system.
 
-Initially, 20 equilibrated boxes of side-lengths 8.9545\angs, with 24 randomly packed water molecules were used.
+Initially, 20 equilibrated cubes of side-lengths 8.9545\angs, with 24 randomly packed water molecules were used.
 Ten of the boxes were used for each of the cold and hot simulations.
@@
 At both temperatures the \suldiox~spends much of the time unbound without any bonding to the surface waters.
-The cluster of the three most populated coordinations is comprised of the ``S'', ``SO'', and ``SOO'' bonding types.
+The cluster of the three most populated bonded coordinations is comprised of the ``S'', ``SO'', and ``SOO'' bonding types.
 This grouping represents the coordinations with a single bond through the sulfur, and up to two interactions through the \suldiox-oxygens.
@@
 
-% Baer's assymetric oxygen binding
 Baer et al.
@@
 At colder temperatures a \suldiox~molecule will bind more strongly to a water surface than at higher temperatures.
-
-%On the cold water surface, the ``S'' and ``SO'' are more populated than for the hot system.
-The relative decrease of these coordinations are matched in the hot surface by an increase of the ``SOO'' configuration.
-This speaks to a dramatic difference in the surface behavior of \suldiox~at the two temperatures.
-The cold \suldiox~spends nearly equal time in the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''.
-Thus, the addition or removal of a bond through the \suldiox-oxygen to a neighboring coordination (e.g.
-addition of an oxygen bond from ``SO'' to ``SOO'', or removal of the bond from ``SO'' to ``S'') is equally probable, as long as the sulfur interaction with the water oxygen does not break.
 
@@
 In the graph of the hot system in Figure \ref{fig:coordination-transitions} (right), the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size.
-%One very surprising result is in the transition from ``SOO-SOOO''.
-This transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations.
-This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox-oxygen.
-As noted earlier, it is likely that ``SOO'' coordinated \suldiox, asymmetrically binding twice through a single oxygen, is being pulled further into the water interface.
-It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox-oxygen.
-Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox~returns to the ``SOO'' coordination.
 
@@
 For each MD trajectory, coordinate data was analyzed to determine if a cyclic hydrate structure was formed as described earlier in this manuscript, and the lifespan data was collected as was done for the bonding coordinations.
-%A timeline was then produced where each timestep was given a value of 1 or 0 depending on whether a \suldiox~hydrate bonding cycle was present or not, respectively.
-This resulted in a time-function, $C(t)$, similar in nature to a time-varying digital signal.
-
-%In an electro-mechanical system, a mechanical switch often outputs a noisy signal, full of transients or ``signal bounce'' before settling to a final value.
-This will appear as a rapid on-off cycling of the signal, and the problem is one of great concern in signal processing.
-Many mechanical, electrical, and software solutions have been devised to suppress the transients, or ``debounce'' the signal.
-Analogously, the hydrate bonding cycles formed in the simulated system often undergo a period of time during formation, or before dissolution, where the $C(t)$ function bounces before settling into a final value.
-The bounce in the function manifests itself in our statistics as a series of rapid switches between cycle presence and cycle absence.
-Physically, the bond lengths within the cycle (or the cycle-to-be) are fluctuating back and forth across our bond-length criteria.
-It is thus an artifact of our algorithmic determination of the presence or absence of a bond.
-Removing this artifact will allow us to examine the longer-time bonding behavior and this is accomplished as follows.
-
-%We wish to find a function based on $C(t)$ that eliminates the brief oscillations during transitions.
-The resulting function, $f(t)$, will only contain information of whether a hydrate bonding cycle is present or absent, and none of the noisy oscillations of the transitions between the two states.
-$f(t)$ may then be used to calculate statistics about the cycle lifespans.
-A representative portion of a cycle time-function, $C(t)$, is plotted for one of the simulated trajectories shown as the dashed black line in Figure \ref{fig:debouncing}.
-At the far left of the plot the function is in the ``no cycle'' state indicating that a bonding cycle has not been detected, and then switches ``on'' as a cycle is found later in time.
-The cyclic structure is very dynamic, constantly moving and distorting, so any of the bonds forming the cyclic bonding structure are liable to break and reform quickly.
-This bouncing between states is manifested in Figure \ref{fig:debouncing} as a series of sharp spikes in the $C(t)$ function lasting less than 10 fs each.
-$C(t)$ was smoothed using a moving gaussian window function with a 10 fs width, having the effect of disregarding cycle breaks or formations of less than 20 fs duration.
-The resulting smoothed function, $C_s(t)$, was then cutoff with the following criteria:
-
-%\[
-%  f(t) = \left\{ 
-%  \begin{array}{l l}
-%    0 & \quad C_s(t) < 0.2\\
-%    1 & \quad C_s(t) \geq 0.2\\
-%  \end{array} \right.
-%\]
-
-%where $f(t)$ is the debounced time-function that represents the lifespans of cycles.
-Figure \ref{fig:debouncing} shows the original time-function of cycle formation and breaking, $C(t)$ (dashed black), the smoothed function, $C_s(t)$ (red), and the final debounced function, $f(t)$ (green).
-
-%\begin{figure}[h!]
-%	\begin{center}
-%		\includegraphics[scale=1.0]{cycle-debouncing-small.png}
-%		\caption{The dynamic environment of the water surface causes rapid breaking and formation of bonds to the \suldiox, and to waters involved in the cyclic hydrating structures.
-A debouncing procedure was developed (described in the text) to account for the very brief ($<$ 20 fs) breaks to the cyclic structure.
-A time function, $C(t)$ (dashed black), shows if the bonding distance criteria are met for all members of the cyclic hydrate structure.
-A gaussian smoothing window was then used to produce $C_s(t)$ (red), which was then used with a cutoff criteria to give the debounced cycle formation time function, $f(t)$ (green).}
-%		\label{fig:debouncing}
-%	\end{center}
-%\end{figure}
 
@@
 The increased temperature appears to cause the \suldiox~to bond in a way that is slightly more conducive to the formation of cyclic structures.
-%This may account for the slightly increased populations of the longer cycle lifespans in the hot system (greater than 4 ps in some cases) compared to the cold system in the cycle lifespan distribution of Figure \ref{fig:cycle-lifespans}.
 
